---
title: "Animate"
output:
  html_document:
    toc: true
    theme: united
---


```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
if(!exists("newId")){
  source('Util.R')
}
```

## Animation Overview

Animatinon can be broke into two basic types:

- discrete: animation in which an attribute value changes instaneously at a given point in time.
- continous: animation in which an attribute value changes or evolves over a period of time.

Discrete Animation is accomplished by using
the **set** element,  Discrete Animation is discussed further 
in the section [Discrete Animation](discrete-animation)

Continous animition is accomplished using one of three elemens: 

- **animate**: transitions most attributes (or a CSS property) from one value to another over a period of time.
- **animateTransform**: transitions *transform attribute* from one value to another over a period of time.
- **animateMotion**: moves a graphical object according to a specified path description

Another way of putting this is

- if *no* *movement* is involved: use **animate**
- if *movement* is involved:
    + if the movement is defined by simple coordinates: use **animate**
    + if the movement is defined by the *transfrom attribute* (such as rotation): use **animateTransform**
    + if the movement is to be specified by a path, use **animateMotion***
    

Continous Animations using the **animate** element are discussed further 
in the section [Continous Animation](#continous-animation)  Continous transistioning of the
*transform attribute* is discussed in the section 
[Continous Movement via the Transform Attribute](#continous-movement-via-the-transform-ttribute)
Continous Movement by path specificaton is discussed in the section
[Continous Movement along a Path](continous-movement-along-a-path).

For any animation, whether discrete or continuous, 
that animation begins by being triggered by an event. That event might be a button press, or page loading,
or completion of some animation. Events are discussed in [Animation Events](#animation-events)

For ease of presentation, all animation examples outside of sections [Animation Events](#animaton-events) section  and  
[One Line Teasers](#one-line-teasers),
are triggered using a button at the bottom right hand corner of the svg.
These buttons are constructed by invoking either the toggle or playBar function. 
The code for both functions are provided in the utilities section.
However, in practice one may want to trigger an events with other customes functions. 


For better understanding of we begin by next exploring is [Animation Events](#animaton-events)

**Note**: We do not discuss - **animateColor**, which specifies a color transformation over time. 
This is a deprecated element, since animate can perform the same task . We strongly recommend 
using **animate** in place of **animateColor**.


## Animation Events
<a name="animatonEvents"></a> 



This section explains triggering animation events. 
By examining this section, one can discover how to create custom event triggers.

For simplicity, the animations here will be restricted to discreted animations, thus the **set** element will be used.


### Onload
<a name="Onload"></a> 

The simplest animation triggering is to trigger the animation  upon loading the document (web page).  
This was illustrated in the short teasers section, however I repeat the code here.  

```
# not run
svgR( wh=c(600, 100), 
  line( xy1=c(0,50), xy2=c(600,50), 
    stroke.width=20, stroke.dasharray=8, stroke.dashoffset=16, stroke="lightblue",   
    animate(attributeName="stroke.dashoffset", from=16, to=0 , dur=0.5,   repeatCount="indefinite")
  )
)
```


As second illustration, consider the problem of filling a circle with an orange color 10 seconds after loading,. We do this by simply setting the **begin** attribute to 10. For the circle to revert back to green, we can provide ***either***

-  a **dur** attribute to indicate the duration (how long the orange is to remain)
-  a **stop** attribute to indicate when to stop ( when the orange is to terminate)

This is an examle of the **set** animation which is discussed in the [Set Animation Section](#setAnimation)

setAnimation


```{r, echo=T, results="asis"}
WH<-c(800,300)
txt<-c(
  "All circles will turn orange after 10 seconds",
  "The first will go back to green after 1 second",
  "The second will go back to green after 2 seconds",
  "The third will remain orange"
)
labs<-c("orange 1 sec","orange 2 sec","forever orange")
cxy<-rbind(WH[1]*c(.3,.5,.7), WH[2]*c(.7,.7,.7))
svgR( wh=WH,
      text("RELOAD BROWSER:", xy=c(0,40), font.size=25),
      lapply(1:4, function(i){ text(txt[i], xy=c(250,i*20+20), font.size=16)}),
      circle(id="setCir1e1", r=70, fill='lightgreen', cxy=cxy[,1],
          set( attributeName="fill", to="orange", begin=10, dur=1 )),
      circle(id="setCir1e1", r=70, fill='lightgreen', cxy=cxy[,2],
          set( attributeName="fill", to="orange", begin=10, end=12 )),
      circle(id="setCir1e1", r=70, fill='lightgreen', cxy=cxy[,3],
          set( attributeName="fill", to="orange", begin=10 )),
      lapply(1:3, function(i)text( labs[i], cxy=cxy[,i], font.size=16 ))
)
```

However, for performance reasons, it is unwise to have too many animations are running concurrently. For that reason, all animations in this section are initiated using the mouse or chaining. 

### MouseOver

```{r, echo=T, results="asis"}
WH<-c(800,200)
svgR( wh=WH,
    text("Circle turns orange on mouse over", xy=c(20,20), font.size=20),
    circle( cxy=WH/2, r=70, stroke="black", fill="lightgreen",
      set( attributeName="fill", 
           to="orange", begin="mouseover", end="mouseout") 
    )
)
```

### MouseOut

```{r, echo=T, results="asis"}
WH<-c(800,200)
svgR( wh=WH,
    text("Circle turns orange on mouse out", xy=c(20,20), font.size=20),
         circle( cxy=WH/2, r=70, stroke="black", fill="lightgreen",
      set( attributeName="fill", 
           to="orange", begin="mouseout", end="mouseover") 
    )
)
```

### MouseClick

```{r, echo=T, results="asis"}
WH<-c(800,200)
svgR( wh=WH,
    text("Circle turns orange for 1 second on mouse click", xy=c(20,20), font.size=20),
    circle( cxy=WH/2, r=70, stroke="black", fill="lightgreen",
      set( attributeName="fill", 
           to="orange", begin="click", dur=1) 
    )
)
```



### A Simple Mouse Button


```{r, echo=T, results="asis"}
WH<-c(800,200)
svgR( wh=WH,
  text("Circle turns orange for 1 second on button press", xy=c(20,20), font.size=20),
  g( id='greenButton',
    rect( cxy=c(60,WH[2]/2), wh=c(100,20) , stroke="black", fill="green", 
    set(  attributeName="fill", 
           to="lightgreen", begin="greenButton.click", dur=.25) 
    ),
    text('Press Me', cxy=c(60,WH[2]/2), stroke="white")
  ),
    #
    circle( cxy=WH/2, r=70, stroke="black", fill="lightgreen",
      set( attributeName="fill", 
           to="orange", begin="greenButton.click", dur=1) 
    )
)
```





<a name="chainingEvents"></a> 

### Chaining Events 

```{r, echo=T, results="asis"}
WH<-c(800,200)
svgR( wh=WH,
   circle( cxy=c(.1,.5)*WH, r=20, stroke="black", fill="lightgreen",
      set( id = 'circEvent0', attributeName="fill", 
           to="orange", begin="click", dur=.1) 
    ),
   circle( cxy=c(.3,.5)*WH, r=20, stroke="black", fill="lightgreen",
      set( id = 'circEvent1', attributeName="r", 
           to="50", begin="circEvent0.end", dur=.1) 
    ),
   circle( cxy=c(.5,.5)*WH, r=20, stroke="black", fill="lightgreen",
      set( id = 'circEvent2', attributeName="r", 
           to="50", begin="circEvent1.end", dur=.1) 
    ),
   circle( cxy=c(.7,.5)*WH, r=20, stroke="black", fill="lightgreen",
      set( id = 'circEvent3', attributeName="r", 
           to="50", begin="circEvent2.end", dur=.1) 
    ),
    text("Click left most circle to start chain reaction", xy=c(20,20), font.size=20
    )
     

)
```



```{r, echo=T, results="asis"}
WH<-c(800,200)
svgR( wh=WH,
    text("Click circle to start chain reaction", xy=c(20,20), font.size=20),
    circle( cxy=c(.1,.5)*WH, r=20, stroke="black", fill="lightgreen",
      set( id = 'eventchain0', attributeName="fill", 
           to="orange", begin="click", dur=.05) 
    ),
    lapply( 1:20, function(i){
    polygon(
      points=c(0,20, 20,0, 0,-20),
      transform=paste0("translate(", 80+i*30,",100)"),
      fill="lightblue",stroke="black",
      set( id=paste0("eventchain",i),attributeName="fill", 
           to="red",  dur=.05,
           begin=paste0("eventchain",i-1,".end")
      )      
    )})

)
```

### Java Script

#### Alert on Button Press

```{r, echo=T, results="asis"}
WH<-c(800,200)
rectid<-newId()
svgR( wh=WH,
  script(
paste0(
'function showColor() {
                  alert("Color of the Rectangle is: "+
                  document.getElementById("',rectid,'").getAttributeNS(null,"fill"));
               }
'      
)),
  text("Alert on button press", xy=c(20,20), font.size=20),
  g( 
    circle(id=rectid, cxy=c(60,WH[2]/2), r=60 , stroke="black", fill="brown"),
    text('Press Me', cxy=c(60,WH[2]/2), stroke="white", fill='white', font.size=20),
    onClick="showColor()"
  )
)
```

#### Animation Start-Stop on Button Press

```{r, echo=T, results="asis"}
WH<-c(800,200)
goStopCirAni<-newId()
goStopButton<-newId()
textStart<-newId()
textStop<-newId()
scriptGS<-paste0(
'var gsOn=false; ',
'function goStop() {
  var animation=document.getElementById("',goStopCirAni,'");
  var button=document.getElementById("',goStopButton,'");
  var textStart=document.getElementById("',textStart,'");
  var textStop=document.getElementById("',textStop,'");
  if(gsOn){
    gsOn=false;
    animation.endElement();
    button.setAttribute("fill","green");
    textStart.setAttribute("visibility","visible");
    textStop.setAttribute("visibility","hidden");
  } else{
    gsOn=true;
    animation.beginElement();
    button.setAttribute("fill","red")
    textStart.setAttribute("visibility","hidden");
    textStop.setAttribute("visibility","visible");
  }
};
')
svgR( wh=WH,
  script(scriptGS),
  text("Controling Animation by a Single Button Press", xy=c(20,20), font.size=20),
  g( # button
    rect( id=goStopButton, cxy=c(.2,.5)*WH, wh=c(140,60), rxy=c(30,30), 
          stroke="black", fill="green"),
    rect(  cxy=c(.2,.5)*WH, wh=c(130,50), rxy=c(30,30), 
          stroke="black", fill="white", opacity=.3),
    text('Start', id=textStart,
         cxy=c(.2,.5)*WH, fill="white", font.size=24, visibility='visible'),
    text('Stop',  id=textStop,
         cxy=c(.2,.5)*WH, fill="white", font.size=24, visibility='hidden'),
    onClick="goStop()"
  ),
  g(
      rect(cxy=WH/2, wh=WH/6, fill='lightblue', stroke='blue', stroke.width=3),
      animateTransform( id=goStopCirAni, attributeName="transform", 
                        type="rotate", from=c(0,WH/2), to=c(360,WH/2), 
                        dur=1, begin='indefinite', repeatCount="indefinite") 
  )
)
```

#### Animation Start-Stop on Button Press

```{r, echo=T, results="asis"}
WH<-c(800,200)
startButton<-newId()
stopButton<-newId()

button %<c-% function( ...){
  args<-list(...)
  stopifnot("id" %in% names(args) & "text" %in% names(args))
  defaults<-list( cxy=c(100,100), wh=c(140,60), rxy=c(30,30), font.size=24, text.fill='white', fill='green')
  args<-c(args, defaults[sapply(args[names(defaults)], is.null)])
  indx<-c("id","wh","cxy", "rxy", "font.size","fill","text", "text.fill") 
  sargs<-args[indx]
  args[indx]<-NULL
  g( id=sargs$id,
    rect(  cxy=sargs$cxy, wh=sargs$wh, rxy=sargs$rxy, 
          stroke="black", fill=sargs$fill),
    rect(  cxy=sargs$cxy, wh=sargs$wh-c(10,10), rxy=sargs$rxy,
          stroke="black", fill="white", opacity=.3),
    text(sargs$text,
         cxy=sargs$cxy, fill=sargs$text.fill, font.size=sargs$font.size),
    args
  )
}

svgR( wh=WH,
  script(scriptGS),
  text("Controling Animation by a Single Button Press", xy=c(20,20), font.size=20),
  button(id=startButton, text='start', fill='green',
      set(attributeName="visibility", to="hidden",  
          begin=paste0(startButton,".click"),
          end  =paste0(stopButton,".click" ))
  ),
  button(id=stopButton, text='stop', fill='red', visibility='hidden',
       set(attributeName="visibility", to="visible", 
           begin=paste0(startButton,".click" ),
           end  =paste0(stopButton,".click" ))
  ),
  rect(cxy=WH/2, wh=WH/6, fill='lightblue', stroke='blue', stroke.width=3,
      animateTransform( attributeName="transform", 
                        type="rotate", from=c(0,WH/2), to=c(360,WH/2), 
                        dur=1, 
                        begin=paste0(startButton,".click" ),
                        end  =paste0(stopButton,".click" ),
                        repeatCount="indefinite"
                        )
  )
)
```

