---
title: "Animate Animate"
output:
  html_document:
    toc: true
    theme: united
---


```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```


Recall, animation can be broken up as discrete or continous.

- Discrete animation is when an object's attributes are changed instaneously at a given point in time.
    + Discrete animation accomplished using the **set** element
- Continous animation is when an object attribute is changed gradually from one value to another over a period
of time. Most attributes can be animated using the Animate Element, the 2 exceptions being *motion along a path*
and the *transform attribute*. In this section we describe use of the **animate** element by presenting numerous
examples.

<a name="animateAnimation"></a> 

## Animate Element

The animate element is considered the general work-horse for continuous animation since it encompasses
everything except transforms and motions along a path. 




## Stroke Attributes

### Stroke Width
```{r, echo=T, results='asis'}
WH<-c(800,200)
svgR( 
    playBar(wh=WH),
    circle( cxy=WH/2, r=40, stroke="black", fill="yellow",
      animate(attributeName='stroke.width', from=0, to=40, 
                    begin=playButton.click(), dur=2, fill="remove") 
    )
)
```

### Stroke dashoffset
```{r, echo=T, results="asis"}
WH<-c(800,200)
soffset<-40*7
svgR( 
    playBar(wh=WH),
    ellipse( cxy=WH/2, rxy=c(40,40), stroke="blue", fill="none",
             stroke.width=10, stroke.dasharray=soffset,
             stroke.dashoffset=soffset,
      animate(attributeName='stroke.dashoffset', from=soffset, to=0, 
                    begin=playButton.click(), dur=.5, fill="remove") 
    )
)
```



## The Fill Attribute

### Fill color
```{r, echo=T, results="asis"}
WH<-c(800,200)
svgR( 
    playBar(wh=WH),
    ellipse( cxy=WH/2, rxy=c(40,40), stroke="black", fill="red",
      animate(attributeName='fill', from='red', to='yellow', 
                    begin=playButton.click(), dur=2, fill="remove") 
    )
)
```



### Animating Linear Gradient Stop Colors

```{r, echo=TRUE, results="asis"}
WH=c(450,300)
svgR(
    playBar(wh=WH),
    defs(
      linearGradient(id="myLinGradAnimate1",
        xy1=c(.1,.1), xy2=c(.9,.9),             
       stop(offset=0,stop.color='green'), #stop is the stop element
       stop(offset=0.1,stop.color='black',
            animate( attributeName='stop.color',
              from='black',to='yellow',
              begin=playButton.click(), dur=3) 
       ),
       stop(offset=0.9,stop.color='red')
      )
    ),
    rect( cxy=c(200,150), 
          fill= "url(#myLinGradAnimate1)", stroke="blue",
          wh=c(200,200)
    )
)
```

### Linear Gradient 


```{r, echo=TRUE, results="asis"}
WH=c(800,300)
svgR(
    playBar(wh=WH),
    defs(
      linearGradient(id="myLinGradAnimate2",
      colors=c('red','white','blue'),
      xy1=c(.5,0), xy2=c(0,1),             
        animate( attributeName='x2',
                from=0,to=1,
                begin=playButton.click(), dur=3
        )
      )
    ),
    rect( xy=c(20,20), wh=WH-c(40,80), 
          fill= "url(#myLinGradAnimate2)", stroke="blue"
    )
)
```

### Animating Linear Gradient Offsets

```{r, echo=TRUE, results="asis"}
WH=c(800,300)
colors=c('blue','white','red')
svgR(
    playBar(wh=WH),
    defs(
      linearGradient(id="myLinGradAnimate3",
        xy1=c(0,0), xy2=c(1,0),   
        lapply( 1:3, function(i){
          stop(offset=0, stop.color=colors[i], #stop is the stop element
            animate(attributeName='offset',
              from=.1*(i),to=.7+.1*(i), dur=3,
              begin=playButton.click())        
          )
        })
      )
    ),
    rect( xy=c(20,20), wh=WH-c(40,80), 
          fill= "url(#myLinGradAnimate3)", stroke="blue"
    )
)
```

### Animating Linear Gradient Offset Beneath a Mask

```{r, echo=TRUE, results="asis"}
WH=c(800,300)
colors=c('darkblue','white','black')
svgR(
    playBar(wh=WH),
    defs(
      linearGradient(id="myLinGradAnimate4",
        xy1=c(0,0), xy2=c(1,0),   
        lapply( 1:3, function(i){
          stop(offset=0, stop.color=colors[i], #stop is the stop element
            animate(attributeName='offset',
              from=.1*(i),to=.7+.1*(i), dur=.4, repeatCount=5,
              begin=playButton.click())        
          )
        })
      )
    ),
    text( "svgR",  cxy=c(.5,.4)*(WH-c(0,40)), font.size=150, 
          fill= "url(#myLinGradAnimate4)", stroke="blue"
    )
)
```

### Animating the Radial Gradient Radius (r)

```{r, echo=TRUE, results="asis"}
WH=c(800,300)
svgR(
    playBar(wh=WH),
    defs(
      radialGradient(id="myRadGradAnimate1",            
       cxy=c(.5,.5), fxy=c(.5,.5), r=.5,
       colors=c('red','white','green'),
             animate( attributeName='r',
               from=.5,to=.1,
               begin=playButton.click(), dur=3
             ) 
      )
    ),
    rect( xy=c(20,20), wh=WH-c(40,80), 
          fill= "url(#myRadGradAnimate1)", stroke="blue",
          wh=c(200,200)
    )
)
```

### Animating the Radial Gradient Focus (fxy)
```{r, echo=TRUE, results="asis"}
WH=c(800,300)
rgid<-autoId()
rgUrl<-sprintf('url(#%s)',rgid)
svgR(
    playBar(wh=WH),
    defs(
      radialGradient(id=rgid,            
       cxy=c(.2,.2), fxy=c(.5,.5), r=.5,
       colors=c('red','white','green'),
             animate( attributeName='cxy',
               from=c(.2,.2),to=c(.8,.8),
               begin=playButton.click(), dur=3
             ) 
      )
    ),
    rect( xy=c(20,20), wh=WH-c(40,80), 
          fill= rgUrl, stroke="blue",
          wh=c(200,200)
    )
)
```



### Animating the Radial Gradient Center (cxy)

```{r, echo=TRUE, results="asis", fig.cap="radial gradient focus animation", fig.align='center'}
WH=c(800,300)
rgid<-autoId()
rgUrl<-sprintf('url(#%s)',rgid)
svgR(
    playBar(wh=WH),
    defs(
      radialGradient(id=rgid,            
       cxy=c(.5,.5), fxy=c(.2,.2), r=.5,
       colors=c('red','white','green'),
             animate( attributeName='fxy',
               from=c(.2,.2),to=c(.8,.8),
               begin=playButton.click(), dur=3
             ) 
      )
    ),
    rect( xy=c(20,20), wh=WH-c(40,80), 
          fill= rgUrl, stroke="blue",
          wh=c(200,200)
    )
)
```




### Animating the Radial Gradient Stops

```{r, echo=TRUE, results="asis"}
WH=c(800,300)
colors=c('red','white','green')
rgid<-autoId()
rgUrl<-sprintf('url(#%s)',rgid)
svgR(
    playBar(wh=WH),
    defs(
      radialGradient(id=rgid,           
        lapply( 1:3, function(i){
          stop(offset=0, stop.color=colors[i], #stop is the stop element
            animate(attributeName='offset',
              from=.1*(i),to=.7+.1*(i), dur=3,
              begin=playButton.click())        
          )
        })
    )),
    rect( xy=c(20,20), wh=WH-c(40,80), 
          fill= rgUrl, stroke="blue",
          wh=c(200,200)
    )
)
```



## Coordinates Movement and Size Changes


Recall continous animition can be divide  into animation that 

- involves movement: 
- animation that which does not.

Continous animation that does not involve movement is implement with the animate element.

The movement specication is one of the following 3:

1. specification of beginning and ending coordinate values (xy or cxy) of an object to be changed (this section)
2. specification of begining and ending transform attribute values of an object to be changed 
3. specification of path along which an object is to move. (

Here we describe the first. See the sections 
[Continous Movement via the Transform Attribute](#continous-movement-via-the-transform-ttribute)
and 
[Continous Movement along a Path](continous-movement-along-a-path) for 2 and 3.


### Radius
```{r, echo=T, results="asis"}
WH<-c(800,200)
svgR( 
    playBar(wh=WH),
    ellipse( cxy=WH/2, rxy=c(40,40), stroke="black", fill="blue",
      animate(attributeName='rx', from=10, to=300, 
                    begin=playButton.click(), dur=2, fill="remove") 
    )
)
```

### Coordinate: cx
```{r, echo=T, results="asis"}
WH<-c(800,200)
svgR( 
    playBar(wh=WH),
    circle( cxy=c(40,100), r=30, stroke="black", fill="lightgreen",
      animate(attributeName='cx', from='40', to='760', 
                    begin=playButton.click(), dur=1, fill="remove") 
    )
)
```
### Coordinates: cxy
```{r, echo=T, results="asis"}
library(svgR)
WH<-c(800,300)
pt<-c(30,30)
svgR( 
    playBar(wh=WH),
    circle( cxy=pt, r=20, stroke="black", fill="lightgreen",
      animate(attributeName='cxy', from=pt, to=WH-pt, 
                    begin=playButton.click(), dur=2, fill="remove") 
    )
)
```

### Circle changing size, fill color and location
```{r, echo=T, results="asis"}
WH<-c(800,200)
svgR( 
    playBar(wh=WH),
    circle( cxy=c(40,100), r=30, stroke="black", fill="lightgreen",
      animate(attributeName='cx', from='40', to='760', 
                    begin=playButton.click(), dur=1, fill="freeze"), 
      animate(attributeName='r', from='30', to='1', 
                    begin=playButton.click(), dur=1, fill="freeze"), 
      animate(attributeName='fill', from='red', to='red', 
                    begin=playButton.click(), dur=1, fill="freeze") 
    )
)
```

## Text 

### Vertically Scrolling Text
```{r, echo=T, results="asis"}
WH<-c(800,200)
txt<-c(
'An ounce of algebra is worth of a ton of verbal argument.',' -JBS Haldane',
'Data sets do not give up their secrets easily.', 'They must be tortured to confess.', '-Jeff Hopper, Bell Labs',
'79.48% of all statistics are made up on the spot.','-John Paulos, Prof of Math',
'All models are wrong','but some are useful.', '-George E. P. Box','In God we trust.', 'All others must bring data.',
'-W. Edwards Deming','I guess I think of lotteries as a tax', 'on the mathematically challenged',
'― Roger Jones','Statisticians are like artists','they have a bad habit',' of falling in love with their models.',
'-- George Box')
dx<-30
WH2<- c(WH[1]-2*dx, dx)
svgR( 
    playBar(wh=WH),
    text("Stat Quotes", xy=c(30,30), font.size=30, stroke='blue'),
    svg(xy=c(dx,60), wh=WH2, viewBox=c(0,0, WH2),
      lapply(1:length(txt), function(i){
        text(txt[i], xy=c(0,i*dx-3), stroke='purple', stroke.width=2, font.size=dx-5)
      }),
      animate(attributeName='viewBox', from=c(0,0, WH2), to=c(0,dx*length(txt), WH2), 
          begin=playButton.click(), dur=1*length(txt), fill="remove")
    )
)
```

### Horizontally Scrolling Text
```{r, echo=T, results="asis"}
WH<-c(800,200)
txt<-c(
'Pure mathematics is, in its way, the poetry of logical ideas.  ~Albert Einstein',
'Mathematics is no more computation than typing is literature.—John Allen Paulos',
'A mathematician is a device for turning coffee into theorems. ~Paul Erdos', 
#'Anyone who cannot cope with mathematics is not fully human.',' At best he is a tolerable subhuman who has learned to wear shoes, bathe,',' and not make messes in the house. ~Robert Heinlein, Time Enough for Love',
"If there is a God, he's a great mathematician.  ~Paul Dirac",
'The laws of nature are but the mathematical thoughts of God.  ~Euclid', 
'The purpose of computing is insight, not numbers!—R. W. Hamming',
'Truth is ever to be found in the simplicity, and not in the multiplicity and confusion of things.—Newton')
dx<-30
WH2<- c(WH[1]-2*dx, dx)
svgR( 
    playBar(wh=WH),
    text("Math Quotes", xy=c(30,30), font.size=30, stroke='blue'),
    svg(xy=c(dx,60), wh=WH2, viewBox=c(0,0, WH2),
      lapply(1:length(txt), function(i){
        text(txt[i], xy=c(1.0*(i-1)*WH2[1],25), stroke='purple', stroke.width=2, font.size=dx/2)
      }),
      animate(attributeName='viewBox', from=c(0,0, WH2), to=c(WH2[1]*length(txt), 0, WH2), 
          begin=playButton.click(), dur=4*length(txt), fill="remove")
    )
)
```



## Gausian Filter Primitive Animations

### Gaussian Blur

This example illustrates a simple Gaussian Blur on the alpha channel where the
standard deviation (in both the x and y directions) is the attribute being animated.

```{r, echo=TRUE, results="asis"}
WH=c(800, 200) # window rect
svgR( 
  playBar(wh=WH),
  text('Animating Simple Gaussing Blur', xy=c(20,30), font.size=20),
  text( 'svgR', cxy=c(.5,.4)*WH, font.size=100, fill="blue",
    filter=filter( y=-5, height=100,
        feGaussianBlur(in1="SourceAlpha", stdDeviation=0, y="-",
          animate(attributeName='stdDeviation',
        from=0,to=20, dur=1, 
        begin=playButton.click())      
    ))
  )
)
```

### Gaussian Blur $\sigma_x$

This example illustrates a simple Gaussian Blur on the alpha channel where the
standard deviation in the x direction is the attribute being animated.


```{r, echo=TRUE, results="asis"}
WH=c(800, 200) # window rect
svgR( 
  playBar(wh=WH),
  text('horizontal blurring repeated twice', xy=c(20,30), font.size=20),
  text( 'svgR', cxy=c(.5,.3)*WH, font.size=100, fill="blue",
       filter=filter( y=-5, height=100,
          feGaussianBlur(in1="SourceGraphic", stdDeviation=c(0,0), y="-",
            animate(attributeName='stdDeviation',
          from=c(0,0),to=c(20,0), dur=1, repeatCount=2,
          begin=playButton.click() )      
      ))
  )
)
```

### Depth Illusion

This example illustrates the illusion of depth, by animating both the x-position of the text and the x postion of the Guaussian blur is animated. The back and forth motion is achieved by chaining the animations.

```{r, echo=T, results="asis"}
WH=c(800, 240) # window rect
dur=.5
svgR(
  playBar(wh=WH),
  text("animate with chaining both text and shadow", xy=c(20,30), font.size=20),
  text( 'svgR', xy=c(200,150), font.size=150,  fill="lightblue", stroke="darkblue",
    filter = filter( xy=c(-10,-10), wh=c(800,120),
      feBlend( x=-10, width=800, 
        in1="SourceGraphic", 
        in2=feGaussianBlur( stdDeviation=3, 
          in1=feOffset( dxy=c(10,10), in1="SourceAlpha",
            animate(id='animateShadow1', attributeName='dx', from=10, to=-10, dur=dur,
              begin=playButton.click(), fill='freeze'
            ),
            animate(attributeName='dx', from=-10, to=10, dur=dur,
              begin="animateShadow1.end+0.5", fill='freeze'
            )
          ) 
        )
      )
    ),
    animate( attributeName="x", from=200, to=200+10, dur=dur,
               begin=playButton.click(), fill='freeze'
            ),
    animate(attributeName="x", from=200+10, to=200, dur=dur,
               begin="animateShadow1.end+0.5", fill='freeze'
            )
    )
)
```

### Pattern


```{r, echo=T, results="asis"}
WH=c(600, 600) # window rect
svgR( 
  playBar(wh=WH),
  circle( cxy=WH/2, r=.3*WH[2], fill=
      pattern( xy=c(0,0), wh=c(40,40), patternUnits="userSpaceOnUse",
        rect(cxy=c(20,20), wh=c(30,30), fill='blue',
          animate( attributeName="fill", from='red', to='green', dur=.5,
               begin=playButton.click(), repeatCount=3
            )                  
          )
         )
  )
)
```

### Radius

```{r, echo=T, results="asis"}
WH=c(600, 600) # window rect
svgR( 
  playBar(wh=WH),
  circle( cxy=WH/2, r=.3*WH[2], stroke='red', stroke.width=5,
      fill=
      pattern( xy=c(10,10), wh=c(40,40), patternUnits="userSpaceOnUse",
        circle(cxy=c(10,10), r=5, fill='blue',
          animate( attributeName="r", from=5, to=20, dur=.5,
               begin=playButton.click(), repeatCount=8
            )                  
          )
         )
  )
)
```

### Radius

```{r, echo=T, results="asis"}
WH=c(800, 400) # window rect
svgR( 
  playBar(wh=WH),
  text('svgR', cxy=WH/2, font.size=250, stroke.width=2, stroke='blue', fill=
      pattern( xy=c(10,10), wh=c(40,40), patternUnits="userSpaceOnUse",
        circle(cxy=c(10,10), r=5, fill='blue',
          animate( attributeName="r", from=5, to=30, dur=.5,
               begin=playButton.click(), repeatCount=8
            )                  
          )
         )
  )
)
```

## Color Rotation

###  hueRotate 1

```{r, echo=T, results="asis"}
WH=c(800, 400) # window rect
svgR( 
  playBar(wh=WH),
  rect(cxy=WH/2-c(0,15), wh=WH-c(40,100), 
    fill=linearGradient( xy1=c(0,0), xy2=c(1,1),
      colors= c('red','yellow','darkblue','brown','orange','green','pink','blue')),
      filter= filter( 
      feColorMatrix(type="hueRotate", values=0,
          animate( attributeName="values", from=0, to=360, dur=1,
              begin=playButton.click(), repeatCount=5
          )                  
      )  
    )
  )
)
```

### hueRotate 2

```{r, echo=T, results="asis"}
WH<-c(800, 400) # window rect
colors<-c('red','yellow','darkblue','orange','green','blue')
svgR( 
  playBar(wh=WH),
  rect(cxy=WH/2-c(0,15), wh=WH-c(40,100), 
    fill=radialGradient( 
      colors= colors),
      filter= filter( 
      feColorMatrix(type="hueRotate", values=0,
          animate( attributeName="values", from=0, to=360, dur=1,
              begin=playButton.click(), repeatCount=5
          )                  
      )  
    )
  )
)
```


## Shape Shifting

### Morphology

Here we animate using feMorphology.

```{r, echo=T, results="asis"}
WH=c(800, 220) # window rect
svgR( 
  playBar(wh=WH), 
  text( 'feMorphology', cxy=WH/2-c(0,15), font.size=100, stroke="red", fill="none",
        filter=filter(
          feMorphology(radius=1, operator='dilate',
             animate( attributeName="radius", from=1, to=10, dur=2,
              begin=playButton.click(), repeatCount=5
             )
          )
        )
  )
)
```


This illustration uses both feTurbulence and feDisplacementMap to create a dissolve effect.

### feTurbulence and feDisplacementMap

This illustration uses both feTurbulence and feDisplacementMap to create a dissolve effect.

```{r, echo=T, results="asis"}
library(svgR)
WH=c(800, 400) # window rect
svgR( 
      rect(xy=c(0,0), wh=WH, fill='lightblue'),
      playBar(wh=WH), 
      symbol(
        rect( id="feTurbulence2", xy=c(0,0), wh=WH, 
          filter=filter( 
            feMerge(
              feMergeNode(in1=feTurbulence(baseFrequency=0.005, numOctaves=1 )),
              feMergeNode(in1=feTurbulence(baseFrequency=c(.001,.01), numOctaves=2 )),              
              feMergeNode(in1=feTurbulence(baseFrequency=0.1, numOctaves=2 ))
          )
        )
      )),
      g(
      text("svgR", cxy=WH/2-c(0,40), stroke='black', font.size=180, stroke.width=2, fill='none'),
      text("feDisplacementMap", cxy=WH/2+c(0,80), stroke='red', font.size=50, stroke.width=2, fill='none'),
           filter=
             filter( xy=c(0,0), filterUnits='userSpaceOnUse',
                feDisplacementMap(
                  in1="SourceGraphic",
                  in2=feImage(xlink.href='#feTurbulence2'),
                  scale=150,
                  yChannelSelector="R",
                  xChannelSelector="B",
                  animate(id='animateFilter2', attributeName='scale', from=150, to=1,
                    begin=playButton.click(), dur=3, fill='freeze'),
                  animate(id='animateFilter2', attributeName='scale', from=1, to=150,
                    begin='animateFilter2.end + 1', dur=2, fill='freeze')
                )
          )
      )
)
```

### Warping Text

```{r, echo=T, results="asis"}
library(svgR)
WH=c(800, 400) # window rect
x<-c(.3,.4,.5,.6)
dx<-.1
y<-.3
sc<-50
col=c("black","gold","red")
XWH<-WH-c(0,100)
svgR( 
      playBar(wh=WH), 
      symbol(
        rect( id="dispGradAni36", xy=c(0,0)*WH,  wh=XWH, fill='blue',
                filter=filter(
                  feTurbulence(baseFrequency=.005, numOctaves=3, seed=100),
                  feComponentTransfer(
                    feFuncR(type="linear", slope=3, intercept=-1),
                    feFuncG(type="linear", slope=2, intercept=-1),
                    feFuncB(type="linear", slope=4, intercept=-1),
                    feFuncA(type="linear", slope=0, intercept=1)
                  ),
                  feColorMatrix(type="saturate")
                )
        )
      ),
      text("A Data Flow", xy=c(30,30), font.size=25, fill="blue"),
      g(
        lapply(1:30, function(i)text(cxy=runif(2)*.7*WH+c(50,50), 
                                     fill=rrgb(), "Flowing Data", font.size=sample(10:30,1)
        )),
      filter=
        filter( xy=c(0,0), filterUnits='userSpaceOnUse',
          feDisplacementMap(
            in1="SourceGraphic",
            in2=feImage(xlink.href='#dispGradAni36'),
            scale=50,
            yChannelSelector="R",
            xChannelSelector="B",
            animate(id='animateFilter5', attributeName='scale', from=sc, to=1,
              begin=playButton.click(), dur=4, fill='freeze'),
            animate(id='animateFilter6', attributeName='scale', from=1, to=sc,
              begin='animateFilter5.end + 3', dur=4, fill='freeze')
          )
        )
      )
)
```

## Controling the Animation Rates

### Using keyTimes and a Linear calcMode 

*KeyTimes* is a vector of strictly increasing numeric  values starting with 0 and ending with 1.
Each *keyTime* value represents a point in time scaled by the duration, *dur*. 

*Values* is vector (or list) describing the value of the attribute at each keyTime.
With *calcMode* = linear, the values between each keyTime interval is interpolated linearly. 

In the example below, the values are the x coordinate of the center of the
circle, *cx*, so the circle appears to move at a 
constant velocity as the *cx* varies linearly within both intervals
(The keyTime intervals are $[0 , 0.95]$ and $[ 0.95, 1]$, and represent in seconds the 
$[0, 2.85]$  and
$[ 2.85, 3]$ )

```{r, echo=T, results="asis"}
WH<-c(800,150)
d<-c("m", WH*c(0,.0), "l", WH*c(1,0))
aniMotion<-autoId()
svgR( 
    playBar(wh=WH),
    rect(xy=c(0,0),wh=WH, fill="#AAFFAA"),
    rect(xy=c(0,0),wh=WH*c(0.2,1), fill="#FFAAAA"),  
    text("Using animate for attribute='cx' with calcMode='linear'", xy=c(30,30), font.size=25, fill="blue"),
    text("Slow", xy=c(30,WH[2]-30), font.size=20, fill="red"),
    text("Fast", xy=c(WH[1]/2,WH[2]-30), font.size=20, fill="green"),
        
    circle( cxy=c(0,.5)*WH, r=10, stroke="black", fill="yellow",
      animate( attributeName='cx', values=c(40, 144, 760),
                    keyTimes=c(0,0.95,1),  
                    begin=playButton.click(), dur=5)
    )
)
```

The rate changes for calcMode Linear are step functions.

**Note:** The default *calcMode* for **animate** is *linear* so we can omit it. 
(Unlike **animateMotion**, where the default *calcMode* is *paced* )


## Using keySplines with animate for a Gradual Rate Change

###  Fast  at the beginning and Slow at the end
```{r, echo=T, results="asis"}
  WH<-c(800,200)
  svgR( 
      playBar(wh=WH),
      rect(xy=c(0,0),wh=WH, fill=linearGradient(x12=c(0,1),colors=c("lightgreen","pink"))),
      text("Using animate for attribute='cx' with calcMode='spline'", xy=c(30,30), font.size=25, fill="blue"),
      circle( cxy=c(40,100), r=30, stroke="black", fill="yellow",
        animate(attributeName='cx', values=c(40,760), calcMode="spline",
              keyTimes=c(0,1), keySplines="0 0.75 0.25 1",
                      begin=playButton.click(), dur=5, fill="remove") 
      )
  )
```
### Slow  at the beginning and Fast at the end
```{r, echo=T, results="asis"}
  WH<-c(800,200)
  svgR( 
      playBar(wh=WH),
      rect(xy=c(0,0),wh=WH, fill=linearGradient(x12=c(0,1),colors=c("pink","lightgreen"))),
      text("Using animate for attribute='cx' with calcMode='spline'", xy=c(30,30), font.size=25, fill="blue"),
      circle( cxy=c(40,100), r=30, stroke="black", fill="yellow",
        animate(attributeName='cx', values=c(40,760), calcMode="spline",
              keyTimes=c(0, 1), keySplines=".75 0 1 .25",
                      begin=playButton.click(), dur=5, fill="remove") 
      )
  )
```

###  Slow  at the beginning and at the end
```{r, echo=T, results="asis"}
  WH<-c(800,200)
  svgR( 
      playBar(wh=WH),
      rect(xy=c(0,0),wh=WH, fill=linearGradient(x12=c(0,1),colors=c("pink","lightgreen","pink"))),
      text("Using animate for attribute='cx' with calcMode='spline'", xy=c(30,30), font.size=25, fill="blue"),
      circle( cxy=c(40,100), r=30, stroke="black", fill="yellow",
        animate(attributeName='cx', values=c(40,760), calcMode="spline",
              keyTimes=c(0, 1), keySplines="0 0.5 0.5 1",
                      begin=playButton.click(), dur=5, fill="remove") 
      )
  )
```
###  Fast  at the beginning and at the end
```{r, echo=T, results="asis"}
  WH<-c(800,200)
  svgR( 
      playBar(wh=WH),
      rect(xy=c(0,0),wh=WH, fill=linearGradient(x12=c(0,1),colors=c("lightgreen","pink","lightgreen"))),
      text("Using animate for attribute='cx' with calcMode='spline'", xy=c(30,30), font.size=25, fill="blue"),
      circle( cxy=c(40,100), r=30, stroke="black", fill="yellow",
        animate(attributeName='cx', values=c(40,760), calcMode="spline",
              keyTimes=c(0, 1), 
              keySplines=c(0,.5,1,.5),
                      begin=playButton.click(), dur=5, fill="remove") 
      )
  )
```


###  Two slow downs
```{r, echo=T, results="asis"}
  WH<-c(800,200)
  svgR( 
      playBar(wh=WH),
      rect(xy=c(0,0),wh=WH, fill=linearGradient(x12=c(0,1),colors=c("lightgreen","pink","lightgreen", "pink","lightgreen"))),
      text("Using animate for attribute='cx' with calcMode='spline'", xy=c(30,30), font.size=25, fill="blue"),
      circle( cxy=c(40,100), r=30, stroke="black", fill="yellow",
        animate(attributeName='cx', values=c(40,(40+760)/2, 760), calcMode="spline",
              keyTimes=c(0, .5, 1), 
              keySplines=rep(c(0,.5,1,.5),2),
                      begin=playButton.click(), dur=5, fill="remove") 
      )
  )
```

## Animating Path-Defined Shapes

### Spike and Twist

```{r, echo=T, results="asis"}
WH<-c(800,300)
M<-30; K<-4
N<-M*K
IM<-seq.int(K,N,K)
theta<-seq(0,2*pi, length.out=N+1)
theta<-theta[-1]
pts <-rbind(cos(theta),sin(theta))
r0<-matrix(80,2,N)
r1<-matrix(20,2,N)
r1[,IM]<-140
pts0<-r0*pts + WH/2
pts1<-r1*pts + WH/2
phi<-pi/4
m2<-matrix(c(cos(phi),sin(phi),-sin(phi),cos(phi)),2,2)
pts2<-pts
pts2[,IM]<-m2%*%pts[,IM]
pts2<-r0*pts2+WH/2
as.path<-function(ptsX){
  c("M",ptsX[,1],"C",ptsX[,2:(dim(ptsX)[2]) ],"Z")
}
goodFill<-'#DDDDFF'
badFill<-'#CC0000'
svgR(
  playBar(wh=WH),
  path(fill=goodFill, stroke='blue', d=as.path(pts0),
    animate( attributeName='d', 
      values=list(as.path(pts0),as.path(pts1),as.path(pts2)),
      begin=playButton.click(), dur=5, fill="remove"
    ),
    animate( attributeName='fill',
      values=c(goodFill,badFill,goodFill),
      begin=playButton.click(), dur=5, fill="remove"
    )
  )
)
```

### Path Animation

```{r, echo=T, results="asis"}
WH<-c(800,300)
M<-30; K<-4
N<-M*K
IM<-seq.int(K,N,K)
theta<-seq(0,2*pi, length.out=N+1)
theta<-theta[-1]
pts <-rbind(cos(theta),sin(theta))
r0<-matrix(90,2,N)
r1<-matrix(70,2,N)
r1[,IM]<-140
pts0<-r0*pts + WH/2
pts1<-r1*pts + WH/2
as.path<-function(ptsX){
  c("M",ptsX[,1],"L",ptsX[,2:(dim(ptsX)[2]) ],"Z")
}
svgR(
  playBar(wh=WH),
  path(fill='red', stroke='blue', d=as.path(pts0),
    animate( attributeName='d', 
      values=list(as.path(pts0),as.path(pts1)),
      begin=playButton.click(), dur=2, fill="remove"
    )
  ), 
  lapply(1:N, function(i){
    circle(cxy=pts0[,i], r=5, fill='green')
  }),
  lapply(1:N, function(i){
    circle(cxy=pts1[,i], r=5, fill='blue')
  })
)
```


### Stab

```{r, echo=T, results="asis"}
WH<-c(800,300)
N<-8
xMarg<-50
NX<-160
X<-seq(0,N*pi, length.out=NX)
sx<-(WH[1]-2*xMarg)/(N*pi)
x<-X*sx+xMarg
wave<-function(phase){
  sy<- - WH[2]/6
  y<-sin(X+phase)
  y<-y*sy+WH[2]/2
  rbind(x,y)
}
as.path<-function(ptsX){
  c("M",ptsX[,1],"C",ptsX[,1:(dim(ptsX)[2]) ])
}
values=lapply(seq(0,4*pi,length.out=24), function(alpha){
  as.path(wave(alpha))
})
svgR(
  playBar(wh=WH),
  text("sin wave", xy=c(20,40), fill='blue', font.size=30),
  path(
    fill='none', stroke='blue', 
    d=as.path( rbind(x, rep(WH[2]/2,NX))),
    animate( attributeName='d', 
      values=values,
      begin=playButton.click(), dur=10, fill="remove"
    )
  )
)
```

### Sin Wave

```{r, echo=T, results="asis"}
WH<-c(800,300)
N<-8
xMarg<-50
NX<-160
X<-seq(0,N*pi, length.out=NX)
sx<-(WH[1]-2*xMarg)/(N*pi)
x<-X*sx+xMarg
wave<-function(alpha){
  sy<- - WH[2]/6
  y<-sin(alpha)*sin(X)
  y<-y*sy+WH[2]/2
  rbind(x,y)
}
as.path<-function(ptsX){
  c("M",ptsX[,1],"C",ptsX[,1:(dim(ptsX)[2]) ],"Z")
}

values=lapply(seq(0,8*pi,length.out=24), function(alpha){
  as.path(wave(alpha))
})
svgR(
  playBar(wh=WH),
  text("standing wave", xy=c(20,40), fill='blue', font.size=30),
  path(
    fill='red', stroke='blue', 
    d=as.path( rbind(x, rep(WH[2]/2,NX))),
    animate( attributeName='d', 
      values=values,
      begin=playButton.click(), dur=5, fill="remove"
    )
  )
)
```


### Standing Wave

```{r, echo=T, results="asis"}
WH<-c(800,300)
pts<-c(1,1)

N<-8
xMarg<-50
NX<-160
X<-seq(0,N*pi, length.out=NX)
sx<-(WH[1]-2*xMarg)/(N*pi)
x<-X*sx+xMarg
wave<-function(alpha){
  sy<- - WH[2]/6
  y<-sin(alpha)*sin(X)
  y<-y*sy+WH[2]/2
  rbind(x,y)
}
as.path<-function(ptsX){
  c("M",ptsX[,1],"C",ptsX[,1:(dim(ptsX)[2]) ],"Z")
}

values=lapply(seq(0,8*pi,length.out=24), function(alpha){
  as.path(wave(alpha))
})
svgR(
  playBar(wh=WH),
  text("standing wave", xy=c(20,40), fill='blue', font.size=30),
  path(
    fill='red', stroke='blue', 
    d=as.path( rbind(x, rep(WH[2]/2,NX))),
    animate( attributeName='d', 
      values=values,
      begin=playButton.click(), dur=5, fill="remove"
    )
  )
)
```

## Animating Light Positions

### Diffuse PointLight Positioning Using x, y, z Attribute Specification

Here we rotate the light source about the center of the circle, starting slightly above 0 to $\pi/2$ using Diffuse + pointlight
Specifying both x and y animation positions seperately.

```{r, echo=T, results="asis"}
WH<-c(800,200)
rho<-90
r=100
theta<-seq(0,2*pi,length.out=20)
x=rho*cos(theta)+ WH[1]/2
y=rho*sin(theta)+ WH[2]/2
z=rep(10,length(theta))

svgR( 
  playBar(wh=WH),
  svg(wh=WH,
    circle( cxy=WH/2, r= r, fill="green", stroke="black" , 
      filter=filter(
        feComposite( 
          operator="arithmetic", k1234=c(0,1,1,0),
          in1="SourceGraphic",
          in2= feDiffuseLighting( lighting.color="white", diffuseConstant=1.2,
            in1="SourceGraphic",
            fePointLight( xyz=c(WH/2,10), #lightPos(pi/2),
              animate( attributeName='x', values=x, dur=5, begin=playButton.click(),fill="remove"),
              animate( attributeName='y', values=y, dur=5, begin=playButton.click(),fill="remove")
            ) 
          )
        )
      )
    )
  )
)
```


### Diffuse PointLight Postioning Using the  xyz Attribute Specification
Again we rotate the light source about the center of the circle, 
starting slightly above 0 to $\pi/2$ using Diffuse + pointlight. 

This time we use the combo attribute *xyz* when specifying the positioning of the light.

```{r, echo=T, results="asis"}
WH<-c(800,200)
rho<-90
r=100
rho<-r
lightPos<-mapply( 
function(theta, phi){
      c(
      x=rho*cos(phi)*cos(theta)+ WH[1]/2,
      y=rho*cos(phi)*sin(theta)+ WH[2]/2,
      z=rho*sin(phi)/10
      )
},
seq(0,2*pi,length.out=20), 
seq(pi/2,pi/8,length.out=20) 
)
svgR( 
  playBar(wh=WH),
  svg(wh=WH,
    circle( cxy=WH/2, r= r, fill="green", stroke="black" , 
      filter=filter(
        feComposite( 
          operator="arithmetic", k1234=c(0,1,1,0),
          in1="SourceGraphic",
          in2= feDiffuseLighting( lighting.color="white", diffuseConstant=1.2,
            in1="SourceGraphic",
            fePointLight( xyz=c(WH/2,10), #lightPos(pi/2),
              animate(attributeName='xyz',
              values=lightPos,
              dur=5, 
              begin=playButton.click(),fill="remove"
              )
            ) 
          )
        )
      )
    )
  )
)
```



