---
title: "AnimateTransform"
output:
  html_document:
    toc: true
    theme: united
---


```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
  source('Util.R')
```

<a name="animateTransform" /> 

## AnimateTransform


AnimateTransform is used to animate the "transform" attribute.
Thus animate transform applies to shapes (and groups) and can animate
any combination of the following:

- translate
- scale
- rotate
- skewX 
- skewY

We start by showing some basic operations:

### AnimateTransform Translate
```{r, echo=T, results="asis"}
WH<-c(800,200)
dur=5
svgR( 
  playBar(wh=WH), 
  graphPaper(wh=WH, dxy=c(40,40), labels=TRUE),
    text( "Animate Transform:  type=translate", xy=c(30,40), fill='darkblue', font.size=20),
    rect( xy=c(80,80), wh=c(40,40), fill='lightblue', stroke='blue', opacity=0.5,
      animateTransform(attributeName="transform", type='translate',
              from=c(0,0), to=c(600,0), 
              begin=playButton.click(), dur=dur
      )
    )    
)
```

## AnimateTransform Rotate

```{r, echo=T, results="asis"}
WH<-c(800,250)
svgR( 
  playBar(wh=WH), 
  graphPaper(wh=WH, dxy=c(20,20), labels=TRUE),
  text( "Animate Transform Rotation", xy=c(25,25), fill='blue', font.size=20),
  polygon( points=c( c(.5,.5), c(.55,.6), c(.45,.6))*(WH+c(0,-20)),
    fill='#80FFFF',stroke='red',
    animateTransform(attributeName="transform", type='rotate',
            from=c(0,WH/2), to=c(360,WH/2), 
            begin=playButton.click(), dur=2, 
            repeatCount=5 
    )
   )
)

```



```{r, echo=T, results="asis"}
WH<-c(800,150)
N<-12
fop<-.07*(2 + (1:N)*(8-2)/N) 
dur=1.25
rc=5
svgR( 
  playBar(wh=WH), 
  text( "I'm busy", xy=c(300,80), fill='red', font.size=60, visibility='hidden',
    set(attributeName="visibility", to="visible",begin=playButton.click(), dur=dur*rc)
  ),
  g(
    lapply(1:N, function(i){
      rect( xy=c(150,50),wh=c(3,10), fill='blue', fill.opacity=fop[i],
          transform=list( rotate=c(i*360/N, 150, 50), translate=c(0,10))
      )
    }),
    animateTransform(attributeName="transform", type='rotate',
            from=c(0,150,50), to=c(360,150,50), 
            begin=playButton.click(), dur=dur, 
            repeatCount=rc 
    )
  )
)
```


## AnimateTransform Scale

#### Animated Scaling With the Origin as the Fixed Point

Scaling can be animated using **animateTransform**, however there are a few
 caveats.

A simple **animateTransform** with *type=scales* will scale about the origin, 
so items not at the origin appear to be moved.

```{r, echo=T, results="asis"}
WH<-c(800,250)
dur=5
rect.wh=c(50,50)
svgR( 
  playBar(wh=WH), 
  graphPaper(wh=WH, dxy=c(20,20), labels=TRUE),
  svg( wh=WH, xy=c(0,0),
    text( "Scaling, with the origin as the fixed point", xy=c(30,30), fill='red', font.size=50, visibility='hidden'),
    rect( xy=c(100,50),wh=rect.wh, fill='lightblue', stroke='blue', opacity=.6,
      animateTransform(attributeName="transform", type='scale',
              from=c(1,1), to=c(2,2), 
              begin=playButton.click(), dur=dur
      )
    )    
  )

)
```
**Norw** here, the *graphPaper* was outside of the scope of *animateTransform*, so it was not scaled


To scale  with a given fixed point *pt* we need to be a slightly more clever.

#### Animate Transform Scaling With a Given Fixed Point.


In this example we 
take a rectangle centered at the middle of the *viewPort* (WH/2) and use 
**animateTransform**  with *type=scale* to scale it, while keeping the rectangle center as a fixed point. 

We get this by 2 nested applications of **animateTransform**. 

```{r, echo=T, results="asis"}
WH<-c(800,250)
dur=5
pt<-WH/2
alpha<-c(3,2)
rect.wh=c(50,50)
svgR( 
  playBar(wh=WH), 
  graphPaper(wh=WH, dxy=c(20,20), labels=TRUE),
  svg( wh=WH, xy=c(0,0),
    text( "Transform ", xy=c(30,30), fill='red', font.size=50, visibility='hidden'),
    g(
      rect( xy=pt-rect.wh/2, wh=rect.wh, fill='lightblue', stroke='blue', opacity=.6,
        animateTransform(attributeName="transform", type='scale', 
              from=c(1,1), to=alpha, 
              begin=playButton.click(), dur=dur
        )
      ),
      animateTransform(attributeName="transform", type='translate', additive='sum',
        from=c(0,0), to=(c(1,1)-alpha)*pt, 
        begin=playButton.click(), dur=dur
      )     
    )
  )
)
```

This works, but the to calculation requires too much forethought.

An easier approach is to use an svg element.

### Animate Transform Scaling With a Given Fixed Point Using svg.


This example, like the last example, 
takes a rectangle centered at the middle of the *viewPort*  and employs 
animateTransform to scale it, keeping the rectangle center fixed. 

However,
instead of using two applications of **animatedTransform**, we 
 wrap the rectangle to be animated inside an **svg** element. 
In particular, we set the viewBox  of that **svg** element to 
have an origin at  at the given point -*pt*, while
we keeping the same width and height.
We use the same viewPort for this **svg** as it's container (xy=c(0,0),wh=WH).

The effect is that the c(p,0) point inside this **svg** is wilt appears at *pt*.
Thus, 
centering the rectangle at c(0,0) gives ust the desired result.




```{r, echo=T, results="asis"}
WH<-c(800,250)
dur=5
pt<-WH/2
rect.wh=c(50,50)
alpha<-c(3,2)
svgR( 
  playBar(wh=WH), 
  graphPaper(wh=WH, dxy=c(20,20), labels=TRUE),
  svg( wh=WH, xy=c(0,0), 
    text( "Transform ", xy=c(30,30), fill='red', font.size=50, visibility='hidden'),
    svg( viewBox=c(-pt, WH),
    # all contained coordinates are relative to xy=pt
      circle(cxy=c(0,0), r=200, fill='red', opacity=.3),
      rect( cxy=c(0,0), wh=rect.wh, fill='lightgreen', stroke='blue', opacity=.6,
        animateTransform(attributeName="transform", type='scale', 
              from=c(1,1), to=alpha, 
              begin=playButton.click(), dur=dur
        )
      ) 
    )   
  )
)

```







```{r, echo=T, results="asis"}
WH<-c(800,250)
dur=5
pt<-WH/2
alpha<-c(3,2)
svgR( 
  playBar(wh=WH), 
  graphPaper(wh=WH, dxy=c(40,40), labels=TRUE),
  svg( wh=WH, xy=c(0,0),
    text( "Transform ", xy=c(30,30), fill='red', font.size=50, visibility='hidden'),
    g(
      rect( cxy=pt, wh=c(50,50), fill='lightblue', stroke='blue', opacity=.6,
        animateTransform(attributeName="transform", type='scale', 
              from=c(1,1), to=alpha, 
              begin=playButton.click(), dur=dur
        )
      ),
      animateTransform(attributeName="transform", type='translate', additive='sum',
        from=c(0,0), to=(c(1,1)-alpha)*pt, 
        begin=playButton.click(), dur=dur
      )     
    )
  )
)
```

## AnimateTransform Skew

### Skewing Along the X Axis with the Origin as the Fixed Point

Here we take a rectangle located at c(0,0) and use **animateTransform** *type='skewX'*
to skew it.

```{r, echo=T, results="asis"}
WH<-c(800,250)
svgR( 
  playBar(wh=WH), 
  graphPaper(wh=WH, dxy=c(20,20), labels=TRUE),
  text( "Animate Transform skewX", xy=c(400,25), fill='blue', font.size=20),
  rect( xy=c(0,0), wh=c(100,100), fill='lightblue', stroke='blue', opacity=.6,
    animateTransform(attributeName="transform", type='skewX',
            from=0, to=45, 
            begin=playButton.click(), dur=5 
            
    )
   )
)
```

### Skewing Along the Y Axis with the Origin as the Fixed Point

Here we take a rectangle located at c(0,0) and use **animateTransform** *type='skewY'*
to skew it.

```{r, echo=T, results="asis"}
WH<-c(800,250)
svgR( 
  playBar(wh=WH), 
  graphPaper(wh=WH, dxy=c(20,20), labels=TRUE),
  text( "Animate Transform skewY", xy=c(400,25), fill='blue', font.size=20),
  rect( xy=c(0,0), wh=c(100,100), fill='lightblue', stroke='blue', opacity=.6,
    animateTransform(attributeName="transform", type='skewY',
            from=0, to=45, 
            begin=playButton.click(), dur=5 
            
    )
   )
)
```

### Skewing Along the X Axis with the Origin with a Given Fixed Point

In this example we 
take a rectangle centered at the middle of the *viewPort* (WH/2) and use 
**animateTransform**  with *type=skewX* to skew it, 
while keeping the rectangle center as a fixed point. 

We employ, the same trick as before, (which should be standard by now)
We wrap the rectangle to be animated inside an **svg** element. 
We then set the viewBox  of that **svg** element to 
have an origin at the point -*pt*, 

The net effect is that the c(p,0) point inside this **svg** is wilt appears at *pt*.
Thus, 
centering the rectangle at c(0,0) gives us the desired result.



```{r, echo=T, results="asis"}
WH<-c(800,260)
rect.wh<-c(100,100)
pt<-WH/2
dur=5
svgR( 
  playBar(wh=WH), 
  graphPaper(wh=WH, dxy=c(20,20), labels=TRUE),
  text( "Animate Transform skewX", xy=c(400,25), fill='blue', font.size=20),
    svg( viewBox=c(-pt, WH),
    # all contained coordinates are relative to xy=pt
      rect( cxy=c(0,0), wh=rect.wh, fill='lightgreen', stroke='blue', opacity=.6,
        animateTransform(attributeName="transform", type='skewX', 
              from=0, to=45, 
              begin=playButton.click(), dur=dur
        )
      ) 
    )  
)
```


### Chaining Transformations



