---
title: "Animate Set"
output:
  html_document:
    toc: true
    theme: united
---


```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
  source('Util.R')
```

 <a name="setAnimation"></a> 



### Overview 
The examples in the [Chaining Events section](#chainingEvents) use the set command to achieve animation. The set command  can change a graphics attribute, but that change is instaneous. The set command is discrete, not continous. 

Attributes that are to be changed are specified by **attributeName**, and the value to change to is given by **values**  Each **set** element can have only one attribute value specification. Thus for multiple attributes, multiple  **set** elements are required, although they can still use the same event to begin, and hence can run concurrently. A [see Animate Onload][Onload]

Specifying a duration  (**dur** attribute) within a **set** element,  determines how long the new value persists. Specifying an **end** attribute within a **set** element determines a time when the new value expires.

If a **dur** or **end** is not specified, then the new attribute value persists indefinitely. In all examples in this section, an **end** or a **dur** will be specified.

## Fill

The fill attribute can be reset using the set command.

In this example we toggle the fill color of some text.

```{r, echo=T, results="asis"}
WH<-c(800,200)
svgR( wh=WH,
   toggleBar(WH),
   text( 'svgR', xy=c(40,120),  font.size=100, stroke="black", fill="lightgreen",
      set(attributeName="fill", to="green",
          begin=toggleButton.click('left'), 
          end=toggleButton.click('right')
          )
    )
)
```


## Visibility & Opacity 

The *Opacity* attributes specifies a level of transparncy, with 1 being opaque, and 0 being essentially
invisible, and .5 being mid-way between. In contrast, the *visibility* attribute
is either visible or hidden. 

### Set  Visibility
Our first example toggles visibility
```{r, echo=T, results="asis"}
WH<-c(800,200)
svgR( wh=WH,
    toggleBar(WH),
   text( 'svgR', xy=c(40,120),  font.size=100, stroke="black", fill="lightgreen",
      set(attributeName="visibility", to="hidden",
          begin=toggleButton.click('left'), 
          end=toggleButton.click('right')
          )
    )
)
```


### Set Opacity
This second example toggles Opacitiy between opaque and semi-transparent.
```{r, echo=T, results="asis"}
WH<-c(800,200)
svgR( wh=WH,
    toggleBar(WH),
   text( dataScience, cxy=WH/2-c(0,10),  font.size=100, stroke="black", fill="lightgreen"),
   ellipse(cxy=WH/2-c(0,30), rxy=WH/2-c(0,40), fill='lightblue',
           stroke='blue',
      set(attributeName="opacity", to=.5,
          begin=toggleButton.click('left'), 
          end=toggleButton.click('right')
          )
    )
)
```

##  Positioning

### Set  Position x

```{r, echo=T, results="asis"}
WH<-c(800,200)
svgR( wh=WH,
   toggleBar(WH),
   text( 'svgR', xy=c(50,120),  font.size=100, stroke="black", fill="lightgreen",
      set(attributeName="x", to="550",
          begin=toggleButton.click('left'), 
          end=toggleButton.click('right')
          )
    )
)
```

### Set  Position xy

```{r, echo=T, results="asis"}
# WH<-c(800,200)
# svgR( wh=WH,
#    toggleBar(WH),
#    text( 'svgR', xy=c(50,120),  font.size=50, stroke="black", fill="lightgreen",
#       set(attributeName="xy", to=c(200,50),
#           begin=toggleButton.click('left'), 
#           end=toggleButton.click('right')
#           )
#     )
#)
```

### Set  Dimensions wh 

```{r, echo=T, results="asis"}
# WH<-c(800,200)
# svgR( wh=WH,
#     toggleBar(WH),
#    rect(  xy=c(50,50),  wh=c(600,50), stroke="black", fill="lightgreen",
#       set(attributeName="wh", to=c(200,100),
#           begin=toggleButton.click('left'), 
#           end=toggleButton.click('right')
#           )
#     )
# )
```

### Set ViewBox (Zooming)

```{r, echo=T, results="asis"}
WH<-c(800,300)
svgR( wh=WH,
    toggleBar(WH),
    svg(xy=c(0,0),wh=c(200,200), viewBox=c(0,0,40,40),
      fill='lightgreen',
      rect(xy=c(1,1), wh=c(2,2), fill='red', stroke='black'),
      set(attributeName="viewBox", to=paste(0,0,4,4),
            begin=toggleButton.click('left'), 
            end=toggleButton.click('right')
      )
    )
)
```

## Stroke Related 

### Set  Stroke  

```{r, echo=T, results="asis"}
WH<-c(800,200)
svgR( wh=WH,
    toggleBar(WH),
   line(  xy1=c(40,120), xy2=c(700,20), stroke.width=10 ,
         stroke="black", fill="lightgreen",
      set(attributeName="stroke", to="red",
          begin=toggleButton.click('left'), 
          end=toggleButton.click('right')
          )
    )
)
```



```{r, echo=T, results="asis"}
WH<-c(800,200)
attr<-c("stroke","stroke.width","stroke.dasharray", "stroke.linecap")
val<-c("red",20,50,"round")
svgR( wh=WH,
    toggleBar(WH),
   line(  xy1=c(40,120), xy2=c(700,20), stroke.width=10 ,
         stroke="black", stroke.dasharray=100,
      lapply(1:4, function(i){
        set(attributeName=attr[i], to=val[i],
            begin=toggleButton.click('left'), 
            end=toggleButton.click('right')
            )
      })
   )
)
```

## Text Setting

### Set Font Size 


In this example letters change size sequentially. This effect is achieved by
using set on each tspan element and triggering them using an offset to the button click.

```{r, echo=T, results="asis"}
WH<-c(800,160)
svgR( playBar(wh=WH),
      text( xy=c(40,100), lapply(1:26, function(i){
        tspan( LETTERS[i] , font.size=30, fill=rrgb(),
            set(attributeName='font.size', from=20,to=50,
                dur=.1, begin=paste0(playButton.click(),' + ',.1*i))
        )}
      ))
)
```

### Set  Font Family
```{r, echo=T, results="asis"}
WH<-c(800,200)
svgR( wh=WH,
   toggleBar(WH),
   text( 'svgR', xy=c(40,100),  font.size=100,  fill="lightgreen", stroke='black',
      set(attributeName="font.family", to="fantasy",
          begin=toggleButton.click('left'), 
          end=toggleButton.click('right')
          )
    )
)
```
