---
title: "Case Study of Lighting"
output:
  html_document:
    toc: true
    theme: united
---


```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
if(!exists("newId")){
  source('Util.R')
}
```



### A case study in combining Light Sources

#### Step 1

Gaussian blur to create depth
```{r, echo=T, results='asis'}
library(svgR)
WH=c(800,400)
R<-.4*WH[2]
svgR(wh=WH+c(0,30),
    circle( cxy=WH/2, r= R,
      filter=filter(
        feGaussianBlur( in1="SourceAlpha", stdDeviation=R/2)
      )
    ),
    text("Step 1: A Gaussian blur", cxy=c(.5,1)*WH-c(0,15), stroke='black')
)
```

### Step 2

Clip to circle

```{r, echo=T, results='asis'}
library(svgR)
WH=c(800,400)
R<-.4*WH[2]
svgR(wh=WH+c(0,30),
    circle( cxy=WH/2, r= R,
      filter=filter(
        feGaussianBlur( in1="SourceAlpha", stdDeviation=R/2)
      ),
      clip.path=clipPath(circle(cxy=WH/2, r=R))
    ),
    text("Step 2: Add clip path", cxy=c(.5,1)*WH-c(0,15), stroke='black')
)
```

#### Step 3 Add diffuse colored light.


```{r, echo=T, results='asis'}
library(svgR)
WH=c(800,400)
R<-.4*WH[2]
svgR(wh=WH,
    circle( cxy=WH/2, r= R,  
      filter=filter( 
        feDiffuseLighting( lighting.color="#44BBFF",  
          diffuseConstant=1,
          surfaceScale=R,
          in1=feGaussianBlur(in1="SourceAlpha", stdDeviation=.5*R),
          feDistantLight( elevation= 60, azimuth=90 )
        )
      ),
      clip.path=clipPath(circle(cxy=WH/2,r=R))
    ),
    text("Step 3: Add diffuse colored light", cxy=c(.5,1)*WH-c(0,15), stroke='black')
  )
```


#### Step 4 Reset filter viewPort

```{r, echo=T, results='asis'}
library(svgR)
WH=c(800,400)
R<-.4*WH[2]
svgR(wh=WH,
    circle( cxy=WH/2, r= R,  
      filter=filter( 
        xy=-.4*c(1,1), wh=1.8*c(1,1), #in percentage
        feDiffuseLighting( lighting.color="#44BBFF",  
          diffuseConstant=1,
          surfaceScale=R,
          in1=feGaussianBlur(in1="SourceAlpha", stdDeviation=.5*R),
          feDistantLight( elevation= 60, azimuth=90 )
        )
      ),
      clip.path=clipPath(circle(cxy=WH/2,r=R))
    ),
    text("Step 4: Add diffuse colored light", cxy=c(.5,1)*WH-c(0,15), stroke='black')
  )
```


#### Step 4
But this is too grainy! so lets smooth it a little :)

```{r, echo=T, results='asis'}
library(svgR)
WH=c(800,400)
R<-.4*WH[2]
svgR(wh=WH,
    circle( cxy=WH/2, r= R,  
      filter=filter( 
        xy=-.4*c(1,1), wh=1.8*c(1,1), #in percentage
        feGaussianBlur(
          stdDeviation=5,
          in1=feDiffuseLighting( lighting.color="#44BBFF",  
            diffuseConstant=1,
            surfaceScale=R,
            in1=feGaussianBlur(in1="SourceAlpha", stdDeviation=.5*R),
            feDistantLight( elevation= 60, azimuth=90 )
          )        
        )
      ),
      clip.path=clipPath(circle(cxy=WH/2,r=R))
    ),
    text("Step 4: Smooth with Gaussian Blur", cxy=c(.5,1)*WH-c(0,15), stroke='black')
  )
```


#### Step 5

Add a spot light with white light and  combine with feComposite

```{r, echo=T, results='asis'}
library(svgR)
WH=c(800,400)
R<-.4*WH[2]
phi<-pi/4
spot.xyz<- c(WH/2,0)+ 2*R*c(0,-sin(phi),cos(phi))
spot.ptsAt<-c(WH/2,0)+ R*c(0,-sin(phi),cos(phi))
svgR(wh=WH,
    circle( cxy=WH/2, r= R,  stroke="none",  fill = "#000068",
      filter=filter( 
        xy=-.4*c(1,1), wh=1.8*c(1,1), #in percentage
        feGaussianBlur(
          stdDeviation=5,
          in1=feComposite( operator="arithmetic", k1234=c(0,1,1,0),
            in1=feDiffuseLighting( lighting.color="#44BBFF",  
                diffuseConstant=1,
                surfaceScale=R,
                in1=feGaussianBlur(in1="SourceAlpha", stdDeviation=.5*R),
                feDistantLight( elevation= 60, azimuth=90 )
            ),
            in2= feDiffuseLighting( lighting.color="#white",  
                diffuseConstant=.8,
                surfaceScale=R,
                in1=feGaussianBlur(in1="SourceAlpha", stdDeviation=.5*R),
                feSpotLight( xyz=spot.xyz, pointsAtXYZ=spot.ptsAt, limitingConeAngle=30 )
            )
          )
        ) 
      ),
      clip.path=clipPath(circle(cxy=WH/2,r=R))
    ),
    text("Step 5: A Spot Light", cxy=c(.5,1)*WH-c(0,15), stroke='black')
  )
```

#### Step 6 Add a shadow


```{r, echo=T, results='asis'}
library(svgR)
WH=c(800,400)
R<-.4*WH[2]
phi<-pi/4
spot.xyz<- c(WH/2,0)+ 2*R*c(0,-sin(phi),cos(phi))
spot.ptsAt<-c(WH/2,0)+ R*c(0,-sin(phi),cos(phi))
RXY<-c(1,.2)*R
svgR(wh=WH,
    ellipse(cxy=WH/2+c(0,R), rxy=RXY, opacity=.5,
      filter=filter( feGaussianBlur(in1="SourceAlpha", stdDeviation=5))
    ),
    circle( cxy=WH/2, r= R,  stroke="none",  fill = "#000068",
      filter=filter( 
        xy=-.4*c(1,1), wh=1.8*c(1,1), #in percentage
        feGaussianBlur(
          stdDeviation=5,
          in1=feComposite( operator="arithmetic", k1234=c(0,1,1,0),
            in1=feDiffuseLighting( lighting.color="#44BBFF",  
                diffuseConstant=1,
                surfaceScale=R,
                in1=feGaussianBlur(in1="SourceAlpha", stdDeviation=.5*R),
                feDistantLight( elevation= 60, azimuth=90 )
            ),
            in2= feDiffuseLighting( lighting.color="#white",  
                diffuseConstant=.8,
                surfaceScale=R,
                in1=feGaussianBlur(in1="SourceAlpha", stdDeviation=.5*R),
                feSpotLight( xyz=spot.xyz, pointsAtXYZ=spot.ptsAt, limitingConeAngle=30 )
            )
          )
        ) 
      ),
      clip.path=clipPath(circle(cxy=WH/2,r=R))
    ),
    text("Step 6: Shadow", cxy=c(.5,1)*WH-c(0,15), stroke='black')
  )
```

#### Step 7 Add a Final Touch


```{r, echo=T, results='asis'}
library(svgR)
WH=c(800,400)
R<-.4*WH[2]
phi<-pi/4
spot.xyz<- c(WH/2,0)+ 2*R*c(0,-sin(phi),cos(phi))
spot.ptsAt<-c(WH/2,0)+ R*c(0,-sin(phi),cos(phi))
RXY<-c(1,.2)*R
svgR(wh=WH,
    ellipse(cxy=WH/2+c(0,R), rxy=RXY, fill="#888888", 
      filter=filter( feGaussianBlur(in1="SourceGraphic", stdDeviation=5))
    ),
    circle( cxy=WH/2, r= R,  stroke="none",  fill = "#000068",
      filter=filter( 
        xy=-.4*c(1,1), wh=1.8*c(1,1), #in percentage
        feGaussianBlur(
          stdDeviation=5,
          in1=feComposite( operator="arithmetic", k1234=c(0,1,1,0),
            in1=feDiffuseLighting( lighting.color="#44BBFF",  
                diffuseConstant=1,
                surfaceScale=R,
                in1=feGaussianBlur(in1="SourceAlpha", stdDeviation=.5*R),
                feDistantLight( elevation= 60, azimuth=90 )
            ),
            in2= feDiffuseLighting( lighting.color="#white",  
                diffuseConstant=.8,
                surfaceScale=R,
                in1=feGaussianBlur(in1="SourceAlpha", stdDeviation=.5*R),
                feSpotLight( xyz=spot.xyz, pointsAtXYZ=spot.ptsAt, limitingConeAngle=30 )
            )
          )
        ) 
      ),
      clip.path=clipPath(circle(cxy=WH/2,r=R))
    ),
    text("R",cxy=WH/2, stroke="white", font.size=200, fill="white")
  )
```


#### An Alternate Approach

One thing you might notice in the proceding sections
was that we only used light filtering.
The original shape was not even filled! 

In practice, we get better results by first filling the shape.
In this we 

- first combine the SourceImage with the distantlight
- then combine that result with the spotlight.
  
The following an example, where
the shape is filled with a darkblue, the spotlight is white, and the distant light
is lightgreen.

```{r, echo=T, results='asis'}
library(svgR)
WH=c(800,300)
R<-.4*WH[2]
phi<-pi/4
spot.xyz<- c(WH/2,0)+ 2*R*c(0,-sin(phi),cos(phi))
spot.ptsAt<-c(WH/2,0)+ R*c(0,-sin(phi),cos(phi))
svgR(wh=WH,
    ellipse(cxy=WH/2+c(0,R), rxy=RXY, fill="#888888", 
      filter=filter( feGaussianBlur(in1="SourceGraphic", stdDeviation=5))
    ),
    circle( cxy=WH/2, r= R,  stroke="none",  fill = "darkblue",
      filter=filter(
        xy=-.4*c(1,1), wh=1.8*c(1,1), #in percentage
        feComposite( operator="arithmetic", k1234=c(0,1,1,0),
          in1=feGaussianBlur(stdDeviation=2,
                in1=feDiffuseLighting( lighting.color="white",  
                  diffuseConstant=.75,
                  surfaceScale=R,
                  in1=feGaussianBlur( stdDeviation=.5*R),
                  feSpotLight( 
                    xyz=spot.xyz, 
                    pointsAtXYZ=spot.ptsAt, 
                    limitingConeAngle=30 
                  )
                )
          ),
          in2=feComposite( operator="arithmetic", k1234=c(0,1,1,0),
            in2=feGaussianBlur(stdDeviation=10,
              in1=feDiffuseLighting( lighting.color="lightgreen",  
                diffuseConstant=1,
                surfaceScale=R,
                in1=feGaussianBlur( stdDeviation=.5*R),
                feDistantLight( elevation= 20, azimuth=90 )
              )
            ),
            in2="SourceGraphic"
          )
        )
      ),
    clip.path=clipPath(circle(cxy=WH/2,r=R))
    ),
    text("R",cxy=WH/2, stroke="white", font.size=R, fill="white")
  )
```



