---
title: "ClipPath"
output:
  html_document:
    toc: true
    theme: united
---


```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
if(!exists("newId")){
  source('Util.R')
}
```

With a defs clipPath works seamlessly, but
```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 100) # window rect
svgR( wh=WH,
    defs(
    clipPath(
      id="clip0",
      rect(cxy=c(0,50), wh=c(600,50))
    )    
    ),
    g(  clip.path="url(#clip0)",
      circle( cxy=c(50,50), r=30, fill='darkblue'),
      circle( cxy=c(150,50), r=50, fill='red'  )
    )
    
)
```


```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 100) # window rect
svgR( wh=WH,
    g(
      clip.path=clipPath( rect(cxy=c(0,50), wh=c(600,50))),
      circle( cxy=c(50,50), r=30, fill='darkblue'),
      circle( cxy=c(150,50), r=50, fill='red'  )
    )
)
```


**Note** removing the defs, we see that it works in firefox, chromium, but not in the RStudio browser on Ubuntu.
```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 100) # window rect
svgR( wh=WH,
    #defs(
    clipPath(
      id="clip1",
      rect(cxy=c(0,50), wh=c(600,50))
    #)    
    ),
    g(  clip.path="url(#clip1)",
      circle( cxy=c(50,50), r=30, fill='green'),
      circle( cxy=c(150,50), r=50, fill='orange')
    )
)
```
