---
title: "Fill"
output:
  html_document:
    toc: true
    theme: united
---


```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```


*Fill'er up*

At this point, we have seen several shapes, with the fill attribute specified either as 

- "none"
- a color, such as 'red', or '#FF0000' or "rgb(255,0,0)"

However, we can also fill with

- a linear gradient
- a radial gradient
- a pattern


## Fill with a Linear Gradient

As  first example, we take an ellipse to fill with a linear gradient.
```{r, echo=T, results="asis"}
colors=c('blue','green','white','orange','red')
svgR( 
  wh=c(600, 200), 
  ellipse(cxy=c(250,100), rxy=c(200,80), 
          fill=linearGradient( xy1=c(0,0), xy2=c(1,0), colors=colors)     
  )
)
```

The above example invokes the svgR quick approach with fill=lineargradient(...), but this is not standard svg. Standard svg requires a reference to a gradient to be assigned to the fill attribute, i.e. fill="url(#id)" and usually the
gradient definition is placed inside a defs element, as illustrated in the following.


```{r, echo=T, results="asis"}
colors<-c('green','white','orange')
svgR( 
   wh=c(600, 200), 
  defs(
    linearGradient(id='grad1', xy1=c(0,0), xy2=c(1,0), colors=colors)
  ),
  ellipse(cxy=c(250,100), rxy=c(200,80), fill="url( #grad1 )") 
)
```

In the second approach, we can reuse the gradient definition
multiple times referencing the url for each use

```{r, echo=T, results="asis"}
color<-c('green','white','orange')
svgR( 
   wh=c(600, 200), 
  defs(
    linearGradient(id='grad1', xy1=c(0,0), xy2=c(1,0),colors=colors)
  ),
  ellipse(cxy=c(50,100), rxy=c(20,80), fill="url( #grad1 )"),
  ellipse(cxy=c(100,100), rxy=c(20,80), fill="url( #grad1 )"),   
  ellipse(cxy=c(150,100), rxy=c(20,80), fill="url( #grad1 )"), 
  ellipse(cxy=c(200,100), rxy=c(20,80), fill="url( #grad1 )"),
  ellipse(cxy=c(250,100), rxy=c(20,80), fill="url( #grad1 )") 
)
```


However, if we explicity assign an id to the fill=lineargrad(id=...), for the first ellipse, we can also refer to that gradient in the subsequent ellipses

```{r, echo=T, results="asis"}
colors<-c('blue','green','white','orange','red')
svgR( 
  wh=c(600, 200), 
  ellipse(cxy=c(50,100), rxy=c(20,80), 
          fill=linearGradient( id="grad2", xy1=c(0,0), xy2=c(1,0), colors=colors)), 
  ellipse(cxy=c(100,100), rxy=c(20,80), fill="url( #grad2 )"),
  ellipse(cxy=c(150,100), rxy=c(20,80), fill="url( #grad2 )"),
  ellipse(cxy=c(200,100), rxy=c(20,80), fill="url( #grad2 )"),
  ellipse(cxy=c(250,100), rxy=c(20,80), fill="url( #grad2 )") 
)
```

Of course, an elegant approach is to put the ellipses in a group and iterate
```{r, echo=T, results="asis"}
colors<-c('blue','green','yellow','white','orange','red','purple')
svgR( 
  wh=c(600, 200),
  g(
    lapply(seq(50,550,by=50), function(x)ellipse(cxy=c(x,100), rxy=c(20,80))),
    fill=linearGradient( xy1=c(0,0), xy2=c(1,0), colors=colors)
  )
)
```

## Filling with a radial gradient

Filling radial gradient is essentially the same as with a linear gradient

```{r, echo=T, results="asis"}
colors=c('blue','green','white','orange','red')
svgR( 
  wh=c(600, 200), 
  ellipse(cxy=c(250,100), rxy=c(200,80), 
          fill=radialGradient(
            colors=colors)     
  )
)
```

##### Radial Gradient fill with center offset
```{r, echo=T, results="asis"}
colors=c('white', 'green', 'blue')
WH=c(600, 200)
S=190 # Size of square
svgR( 
  wh=WH, 
  rect(cxy=WH/2, wh=c(190,190), stroke='black', stroke.width=3,
      fill=radialGradient( 
        cxy=c(.5,.5), 
        r=.5,
        fxy=c(.6,.6)+c(0,0),
        colors=colors)     
  )
)
```

Here the attributes of the radialGradient are:

- cxy The center of the outer circle
- r The radius of the outer circe
- fxy The center of the gradient focus
- colors, an array of gradient colors

All the above units are in relative to the bounding box, which in this case is the dimensions of the rectangle.

Alternativly, we can set the units via
filterUnits="userSpaceOnUse",

##### Radial Gradient fill with center offset
```{r, echo=T, results="asis"}
colors=c('white', 'green', 'blue')
WH=c(600, 200)
S=190 # Size of square
svgR( 
  wh=WH, 
  rect(cxy=WH/2, wh=c(S,S), stroke='black', stroke.width=3,
    fill=radialGradient( gradientUnits="userSpaceOnUse",
            cxy=WH/2, 
            r=.5*S,
            fxy=WH/2+.1*S,
            colors=colors)     
  )
)
```

## Fill with Patterns

We can also fill with a pattern (Although, depending on the your system, you may need to refresh after scrolling)

```{r, echo=T, results="asis"}
WH=c(600, 200) # window rect
svgR( wh=WH,
  ellipse( cxy=c(250,100), rxy=c(200,80), 
    fill=pattern( xy=c(10,10), wh=c(40,40), patternUnits="userSpaceOnUse",
                  circle(cxy=c(10,10), r=10, fill='blue')
         )
  )
)
```


Or using defs for the fill

```{r, echo=T, results="asis"}
WH=c(600, 200) # window rect
patternId<-autoId()
patternUrl<-sprintf("url(#%s)",patternId)
svgR( wh=WH,
  defs(
    pattern(
        id=patternId, xy=c(10,10), wh=c(40,40), patternUnits="userSpaceOnUse" ,
        circle(cxy=c(10,10), r=10, fill='red')
    )
  ),
  ellipse( cxy=c(250,100), rxy=c(200,80), fill=patternUrl)

)
```

And we can compose pattern with gradients:

```{r, echo=T, results="asis"}
WH=c(600, 200) # window rect
colors=c('white','yellow','orange','green', 'blue','black')
svgR( wh=WH,
  ellipse( cxy=c(250,100), rxy=c(200,80), 
    fill=pattern( xy=c(10,10), wh=c(40,40), patternUnits="userSpaceOnUse",
                  rect(cxy=c(20,20), wh=c(40,40), 
                       fill=radialGradient(  colors=colors)
                  )
         )
  )
)
```

## Filling text

We can also fill text with gradients and patterns

```{r, echo=T, results="asis"}
WH=c(600, 300) # window rect
colors1=c('white','yellow','orange','green', 'blue','black')
colors2=c('blue','green','orange','yellow','orange', 'green','blue')
svgR( wh=WH,
  text( cxy=c(150,100), font.size=300, "R", stroke='green',
        fill=pattern( xy=c(0,0), wh=c(20,20), patternUnits="userSpaceOnUse",
                  text("R", cxy=c(10,10), stroke='red')
        )
  )
)
```

```{r, echo=T, results="asis"}
WH=c(600, 300) # window rect
svgR( wh=WH,
  text( cxy=c(150,100), font.size=300, "R",
        fill=radialGradient( colors=colors1)
  ),
  text( cxy=c(420,50), font.size=60, "Programming",
        fill=linearGradient(colors=colors2)
  )
)
```




