---
title: "Filter Combining Elements"
output:
  html_document:
    toc: true
    theme: united
---


```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
  source('Util.R')
```

Almost always, more than one filter component is required by the filter, and those components need to be combined. There are
three elements that can be used to combine filter components.

## feComposite (operator=over)


```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
    circle(cxy=WH/2, r=30, stroke='black', fill='red' ,
      filter=filter(  wh=c('200%','100%'), xy=c(0,0), 
        feComposite(  
          operator="over", 
          in1="SourceGraphic",
          in2=feOffset(  dxy=c(20, 0), 
             in1=feColorMatrix( type="hueRotate", values=240, "SourceGraphic")
          )
       )
     )
  )
)
```

### feComposite comparison (operator= over, in, out, atop, xor) 

Compariong the operator options over, in, out, atop, xor for feComposite:

#### feComposite comparison with opacity=1

```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
operators=c("over","in","out","atop","xor")
svgR( wh=WH,
      lapply( 1:5, function(i){
        x=100+(i-1)*150
        g( 
          circle(cxy=c(x,40), r=30, stroke='black', fill='red',
            filter=filter(  wh=c('200%','100%'), xy=c(0,0), 
              feComposite(  
                operator=operators[i], 
                in1="SourceGraphic",
                in2=feOffset(  dxy=c(20, 0), 
                   in1=feColorMatrix( type="hueRotate", values=240, "SourceGraphic")
                )
             )
           )
          ),
          text( operators[i], xy=c(x,82), font.size=15 )
        )
    })
)
```


Here we compare feComposite for an source with opacity=50%

#### feComposite comparison with opacity=.5

```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
operators=c("over","in","out","atop","xor")
svgR( wh=WH,
      lapply( 1:5, function(i){
        x=100+(i-1)*150
        g( 
          circle(cxy=c(x,40), r=30, stroke='black', fill='red', fill.opacity=0.5,
            filter=filter(  wh=c('200%','100%'), xy=c(0,0), 
              feComposite(  
                operator=operators[i], 
                in1="SourceGraphic",
                in2=feOffset(  dxy=c(20, 0), 
                   in1=feColorMatrix( type="hueRotate", values=240, "SourceGraphic")
                )
             )
           )
          ),
          text( operators[i], xy=c(x,82), font.size=15 )
        )
    })
)
```

### feComposite (operator=arithmetic)

The arithmetic operator combines the inputs by 
summing the individual inputs and their product, and then adds a constant.

The result is given by
$$ result= \kappa_1 in_1 in_2 + \kappa_2 in_1  + \kappa_3 in_2 + \kappa_4$$


Thus 
- k1234=c(0,1,1,0) sums the two components $in_1$ and $in_2$.
- k1234=c(0,0,0,1) produces white (essentially ignoring the inputs)
- k1234=c(0,1,0,0) produces the $in_1$ ( ignoring the $in_2$)
- k1234=c(0,0,1,0) produces the $in_2$ ( ignoring the $in_1$)

The following example illustrates various combinations of values

```{r, echo=T, results="asis"}
WH=c(800, 800) # window rect
WH2=WH/4
kappaXX<-expand.grid( c(0,1),c(0,2),0:1,c(0,-.5))
pos<-expand.grid(0:3,0:3)
svgR( wh=WH,
      lapply( 1:16, function(i){
        xy=pos[i,]*WH2
        kap<-kappaXX[i,]
        svg( xy=xy, wh=WH2,
          #rect(xy=c(5,5), wh=WH2-c(10,10), fill='black'),
          #circle(cxy=WH2/2, r=WH2[1]/8, stroke='black', fill='red'),
          circle(cxy=WH2/2-c(10,0), r=WH2[1]/8, stroke='black', fill='red', #opacity=.5,
            filter=filter(  wh=c("200%","100%"), xy=c(0,0), 
              feComposite(  
                operator= "arithmetic", 
                k1234=as.numeric(kap),
                in1="SourceGraphic",
                in2=feOffset(  dxy=c(20, 0), 
                   in1=feColorMatrix( type="hueRotate", values=140, "SourceGraphic")
                )
              )
            )
          ),
          text( paste(kap, collapse=", "), xy=c(.1,.9)* WH2, font.size=15 , fill='black')
        )
    })
)->doc
cat("'",as.character(doc),"'")
```





## feBlend

Next we consider feBlend

```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
modes=c("normal","multiply","screen","darken","lighten")
svgR( wh=WH,
      lapply( 1:5, function(i){
        x=100+(i-1)*150
        g( 
          circle(cxy=c(x,40), r=30, stroke='black', fill='red', 
            filter=filter(  wh=c('200%','100%'), xy=c(0,0), 
              feBlend(  
                mode=modes[i], 
                in1="SourceGraphic",
                in2=feOffset(  dxy=c(20, 0), 
                   in1=feColorMatrix( type="hueRotate", values=240, "SourceGraphic")
                )
             )
           )
          ),
          text( modes[i], xy=c(x,82), font.size=15 )
        )
    })
)
```


### feMerge

Finally feMerge

```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
    circle(cxy=WH/2, r=30, stroke='black', fill='red' ,
      filter=filter(  wh=c('200%','100%'), xy=c(0,0), 
        feMerge(  
          feMergeNode(in1="SourceGraphic"),
          feMergeNode(in1=feOffset(  dxy=c(20, 0), 
             in1=feColorMatrix( type="hueRotate", values=240, in1="SourceGraphic")
          ))
       )
     )
  )
)
```

####feMerge using defs
```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
  defs( 
    filter(  id="myMergeFilter", wh=c('200%','100%'), xy=c(0,0), 
     feColorMatrix( in1='SourceGraphic', type="hueRotate", values=240,  result='myColor'),
     feOffset(dxy=c(20, 0), in1="myColor", result="offed" ),
     feMerge(  
        feMergeNode(in1="SourceGraphic"),
        feMergeNode(in1= "offed")
     )
    )
  ),
  circle(cxy=WH/2, r=50, stroke='black', fill='red' , 
    filter="url(#myMergeFilter)"
  )
)
```

