---
title: "DropShadows"
output:
  html_document:
    toc: true
    theme: united
---


```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```


Drop shadows of a figure can be made by applying 
alpha channel of an image with a small offset and then combining it with the original image.  Specifically we 

1. Apply an offset to the original source using feOffset
2. Take the result of 1 and apply feGaussianBlur to the Alpha channel of the original source to 
create the shadow portion.
3. Combine using **feComposite** the original source with the result of 2

#### feBlend

```{r, echo=T, results="asis"}
WH=c(800, 120) # window rect
svgR( wh=WH,
  text( 'svgR', xy=c(50,70), font.size=50,  fill="lightblue", stroke="darkblue",
    filter = filter( xy=c(-10,-10), wh=c(200,60),
              feBlend( x=-10, width=200, 
                in1="SourceGraphic", 
                in2=feGaussianBlur( stdDeviation=3, in1=feOffset( dxy=c(3,3), in1="SourceAlpha") )
              )
      )   
  )
)
```

**Note:** Here, for feBlend, in1 is placed on top, in2 is placed below.

Alternatively, we can place the filter definition in a defs statement and refer to the filter by its id. 

```{r, echo=T, results="asis"}
WH=c(600, 120) # window rect
svgR( wh=WH,
  defs( 
    filter(
      id="blendFilter2", xy=c(-10,-10), wh=c(200,100),
      feOffset( in1="SourceAlpha", dxy=c(3,3), result="offset2"), 
      feGaussianBlur(in1="offset2", stdDeviation=3, result="blur2"),
      feBlend( in1="SourceGraphic",  in2="blur2", x=-10, width=200)
    )         
  ),
  text( 'svgR', xy=c(50,70), font.size=50,  fill="lightblue", stroke="darkblue",
           filter="url(#blendFilter2)")
  
)
```

Now suppose that we want to replace the black shadow with a different colored shadow, for example 'red'. Recall that our
shadow came about by applying a feGaussianBlur to a SourceAlpha. But SourceAlpha gives us black.
So we need to replace SourceAlpha with something else. One way is to apply a red flood to the source graphic. To limit the flood to SourceGraphic, we will feComposite the flood with the SourceGraphic.

The following demonstrates our approach for obtaining the red text using the flood filter:
```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(800, 120) # window rect
svgR( wh=WH,
  text( 'svgR', xy=c(50,70), font.size=50,  fill="lightblue", stroke="darkblue",
    filter = filter(
      feComposite(
        operator="in",
        in1=feFlood(flood.color='red', flood.opacity=1),
        in2="SourceGraphic"
      )
    )
  )
)
```

Now we replace "SourceAlpha" with the feComposite in our original drop shadow to get

```{r, echo=T, results="asis"}
WH=c(800, 120) # window rect
svgR( wh=WH,
  text( 'svgR', xy=c(50,70), font.size=50,  fill="lightblue", stroke='darkblue',
    filter = filter(
      feBlend( x=-10, width=200,
        in1="SourceGraphic", 
        in2=feGaussianBlur( stdDeviation=3,
          in1=feOffset( dxy=c(3,3), 
              in1=feComposite( operator="in",
                  in1=feFlood(flood.color='red', flood.opacity=1), 
                  in2="SourceGraphic"
              )
          )
        )
      )
    )
  )
)
```

#### Some Fun with feComposite
Let's return to our shadow example




## Fun with Drop Shadows

Lets have some fun with drop shadows and feComposite.

First lets change the color again.
```{r, echo=T, results="asis"}
WH=c(800, 120) # window rect
svgR( wh=WH,
  text( 'svgR', xy=c(50,70), font.size=50,  fill="orange", stroke='brown',
    filter = filter(
      feBlend( x=-10, width=200,
        in1="SourceGraphic", 
        in2=feGaussianBlur( stdDeviation=3,
          in1=feOffset( dxy=c(3,3), 
              in1=feComposite( operator="in",
                  in1=feFlood(flood.color='blue', flood.opacity=1), 
                  in2="SourceGraphic"
              )
          )
        )
      )
    )
  )
)
```

Or stick with blue

```{r, echo=T, results="asis"}
WH=c(800, 120) # window rect
svgR( wh=WH,
  text( 'svgR', xy=c(50,70), font.size=50,  fill="lightblue", stroke='blue',
    filter = filter(
      feBlend( x=-10, width=200,
        in1="SourceGraphic", 
        in2=feGaussianBlur( stdDeviation=3,
          in1=feOffset( dxy=c(3,3), 
              in1=feComposite( operator="in",
                  in1=feFlood(flood.color='blue', flood.opacity=1), 
                  in2="SourceGraphic"
              )
          )
        )
      )
    )
  )
)
```

Or play with the in operator of feComposite

```{r, echo=T, results="asis"}
WH=c(800, 120) # window rect
operators=c("over","in","out","atop","xor")
svgR( wh=WH,
  lapply( 1:5, function(i){
    x=20+(i-1)*150
    g( 
      text( 'svgR', xy=c(x,55), font.size=50,  fill="lightblue", stroke='black',
        filter = filter(
          feComposite( x=x-10, width=200, 
            operator=operators[i],
            in1="SourceGraphic", 
            in2=feGaussianBlur( stdDeviation=5,
              in1=feOffset( dxy=c(3,3), in1="SourceAlpha")
            )
          )
        )
      ),
      text( operators[i], xy=c(x,92), font.size=15 )
    )
  })
)
```

Next we modify the alpha channel

```{r, echo=T, results="asis"}
WH=c(800, 120) # window rect
operators=c("over","in","out","atop","xor")
svgR( wh=WH,
  lapply( 1:5, function(i){
    x=20+(i-1)*150
    g( 
      text( 'svgR', xy=c(x,55), font.size=50,  fill="blue", stroke='black',
        filter = filter(
          feComposite( x=x-10, width=200, 
            operator=operators[i],
            in1="SourceGraphic", 
            in2=feComponentTransfer(
             feFuncA( type='table', 
                       tableValues=c(0, .5, 0, .5, 0, .5, 0, .5) 
              ),
              in1=feGaussianBlur( stdDeviation=10, in1="SourceAlpha")
            )
          )
        )
      ),
      text( operators[i], xy=c(x,92), font.size=15 )
    )
  })
)
```

## feComponentTransfer

Colors can also be remapped using feComponentTransfer and the components feFuncR, feFuncG, feFuncB, feFuncA. The feFuncX's correspond to the 
Red, Green, Blue, and Alpha channels. Each channel has a value between between 0 and 1 which can be reassinged using the component transfers. 
There are four types of color component transfers:

- **table**: Reassigns the colors according to a piece-wise continous linear function that is specified by a "table"" values of points. The points are assumend to form a uniform partion of [0,1] so only the values need to be specified. (Using linear interpolation of the values between the points) 
- **linear** Similar to table, but with only 2 points. The function is specifed using the slope and intercept
- **discrete** Reassigns the colors according to a step function that is specified by a "table"" of points. Again the points are assumed to be a uniform partition of [0,1], so only the values need be specified in the "table"." (Between points, the output is the value of the nearest point.
- **gamma** The value is given by
$C_{out} = \alpha \times C_{in}^\gamma + \beta$

####  feComponentTransfer: Table
We illustrate this with a simple example that interchanges the colors blue and red of a color spectrum.

```{r, echo=T, results="asis"}
WH=c(800, 200) # window rect
colors=c('red','blue')
svgR( wh=WH, stroke="black",
  rect(wh=c(.6,.3)*WH, xy=c(.2,0)*WH, 
       fill=linearGradient(colors=colors)
  ),
  rect(wh=c(.6,.3)*WH, xy=c(.2,.3)*WH, 
       fill=linearGradient(colors=colors),
       filter=filter( 
          feComponentTransfer( 
              feFuncR( type="table", tableValues="1 0"),
              feFuncB( type="table", tableValues="1 0 ")
          )
        )
  ),
  text('original', xy=c(20,1.5*WH[2])),
  text('transfered', xy=c(20,4.5*WH[2])),
  text('feComponentTransfer using type="table"', cxy=c(.5,.8)*WH)
)
```


#### feComponentTransfer: Linear

```{r, echo=T, results="asis"}
WH=c(800, 200) # window rect
colors=c('red','blue')
svgR( wh=WH, stroke="black",
  rect(wh=c(.6,.3)*WH, xy=c(.2,0)*WH, 
       fill=linearGradient(colors=colors)
  ),
  rect(wh=c(.6,.3)*WH, xy=c(.2,.3)*WH, 
       fill=linearGradient(colors=colors),
       filter=filter( 
          feComponentTransfer( 
              feFuncR( type="linear", intercept=1, slope=-1),
              feFuncB( type="linear", intercept=1, slope=-1)
          )
        )
  ),
  text('original', xy=c(20,1.5*WH[2])),
  text('transfered', xy=c(20,4.5*WH[2])),
  text('feComponentTransfer using type="linear"', cxy=c(.5,.8)*WH)
)
```


#### feComponentTransfer: Discrete

```{r, echo=T, results="asis"}
WH=c(800, 200) # window rect
colors=c('red','blue')
svgR( wh=WH, stroke="black",
  rect(wh=c(.6,.3)*WH, xy=c(.2,0)*WH, 
       fill=linearGradient(colors=colors)
  ),
  rect(wh=c(.6,.3)*WH, xy=c(.2,.3)*WH, 
       fill=linearGradient(colors=colors),
       filter=filter( 
          feComponentTransfer( 
              feFuncR( type="discrete", tableValues="1 0"),
              feFuncB( type="discrete", tableValues="1 0 ")
          )
        )
  ),
  text('original', xy=c(20,1.5*WH[2])),
  text('transfered', xy=c(20,4.5*WH[2])),
  text('feComponentTransfer using type="discrete"', cxy=c(.5,.8)*WH)
)
```


#### feComponentTransfer: Gamma exponent=1

```{r, echo=T, results="asis"}
WH=c(800, 200) # window rect
colors=c('red','blue')
svgR( wh=WH, stroke="black",
  rect(wh=c(.6,.3)*WH, xy=c(.2,0)*WH, 
       fill=linearGradient(colors=colors)
  ),
  rect(wh=c(.6,.3)*WH, xy=c(.2,.3)*WH, 
       fill=linearGradient(colors=colors),
       filter=filter( 
          feComponentTransfer( 
              feFuncR( type="gamma", offset=1, amplitude=-1, exponent=1),
              feFuncB( type="gamma", offset=1, amplitude=-1, exponent=1)
          )
        )
  ),
  text('original', xy=c(20,1.5*WH[2])),
  text('transfered', xy=c(20,4.5*WH[2])),
  text('feComponentTransfer using type="gamma"', cxy=c(.5,.8)*WH)
)
```


#### feComponentTransfer: Gamma, exponent=10

```{r, echo=T, results="asis"}
WH=c(800, 200) # window rect
colors=c('red','blue')
svgR( wh=WH, stroke="black",
  rect(wh=c(.6,.3)*WH, xy=c(.2,0)*WH, 
       fill=linearGradient(colors=colors)
  ),
  rect(wh=c(.6,.3)*WH, xy=c(.2,.3)*WH, 
       fill=linearGradient(colors=colors),
       filter=filter( 
          feComponentTransfer( 
              feFuncR( type="gamma", offset=1, amplitude=-1, exponent=2),
              feFuncB( type="gamma", offset=1, amplitude=-1, exponent=2)
          )
        )
  ),
  text('original', xy=c(20,1.5*WH[2])),
  text('transfered', xy=c(20,4.5*WH[2])),
  text('feComponentTransfer using type="table"', cxy=c(.5,.8)*WH)
)
```


###feComponentTransfer: Inverting the grey scale

```{r, echo=T, results="asis"}
WH=c(800, 200) # window rect
colors=c('red','blue')
svgR( wh=WH, stroke="black",
  rect(wh=c(.6,.3)*WH, xy=c(.2,0)*WH, 
       fill=linearGradient(colors=colors)
  ),
  rect(wh=c(.6,.3)*WH, xy=c(.2,.3)*WH, 
       fill=linearGradient(colors=colors),
       filter=filter( 
          feComponentTransfer( 
              feFuncR( type="table", tableValues="1 0"),
              feFuncG( type="table", tableValues="1 0"),
              feFuncB( type="table", tableValues="1 0 ")
          )
        )
  ),
  text('original', xy=c(20,1.5*WH[2])),
  text('transfered', xy=c(20,4.5*WH[2])),
  text('feComponentTransfer using type="table"', cxy=c(.5,.8)*WH)
)
```









