---
title: "Filter Attribute"
output:
  html_document:
    toc: true
    theme: united
---



```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```


##Overview of Filters

###Introduction


In the previous section we explored how the *fill* attribute can be applied to text and shapes.
Another
attribute for text and shapes, is the *filter* attribute. 
*Filter *attributes  may take as  values  a **filter element** or a *reference* 
to a **filter element**. 
A **Filter element** is a container for **filter effect elements**, 
also called **filter primitives** or **fe-elements**
. The **fe-elements**  feed into each other to 
form  a directed graph with a root. Typically, the fe elements are linked together using the
*in1*, *in2* attributes. One might visualize a filter to look something like



```{r, echo=FALSE, results="asis"}
WH<-c(800,150)
svgR( wh=WH,
     text(cxy=WH/2,"ToDo!!! Filter Diagram", font.size=60, fill='red', stroke='black')
)
```


While the fill attribute is restricted to the interior of a shape, filters can effect the entire  shape.

 
**Note** Pure *SVG* allows only references, but *svgR* allows to use
the **filter element** directly as the value of a *filter attribute*.


**Todo!!!:** 

- feFunc***X*** 
- feTile



### The In Attributes of  Filter components

Most **fe-element** (filter primitive) require an input specification, which 
is preformend by assigning
values to their *in* attributes: *in1* and *in2*.


(Two notable exceptions to this are the fe-elements: **feImage** and **feFlood**.)

Values for the in *attributes* are given by the following table

Value | Description
----------------------|------------------------------------------------------------
*SourceGraphic* |represents the original element to which the filter is applied (as full RGBA image data)
*SourceAlpha* | represents just the alpha channel of the original element to which the filter is applied
*BackgroundImage* | represents the area directly underneath the filter region
*BackgroundAlpha* | represents just the alpha channel of the area underneath the filter region
*FillPaint* | represents whatever fill was applied to the element in question
*StrokePaint* | represents whatever stroke was applied to the element
a reference | Any other value should be a reference equal to the result attribute for any preceding filter primitive. In this case, the output of the referenced primitive will be used as the specified input.





### The fe-Elements (Filter primitives)


Element               |                Attributes                                  
----------------------|------------------------------------------------------------
feBlend               | in2="***second source***" <br> mode="normal"  "multiply"  "screen"  "darken"  "lighten"
feColorMatrix         | type="matrix"  "saturate"  "hueRotate"  "luminanceToAlpha" <br> values="***matrix values***"  "***saturation value (0-1)***"  "***rotate degrees***"
feComponentTransfer   | container for <br>feFuncR, <br>feFuncG, <br>feFuncB, and <br>feFuncA elements.
feFunc***X***         | type="identity"  "table"  "discrete"  "linear"  "gamma" <br> tableValues="***intervals for table, steps for discrete***" <br> slope="***linear slope***" <br> intercept="***linear intercept***" <br> amplitude="***gamma amplitude***" <br> exponent="***gamma exponent***" <br> offset="***gamma offset***" 
feComposite           | in2="***second source***" <br> operator="over"  "in"  "out"  "atop"  "xor"  "arithmetic" <br> The following attributes are used with arithmetic: <br> k1="***factor for in1,in2***" <br> k2="***factor for in1***" <br> k3="***factor for in2***" <br> k4="***additive offset***"
feConvolveMatrix      | order="***columns rows***" (default 3 by 3) <br> kernel="***values***" <br> bias="***offset value***"
feDiffuseLighting     | Container for a light source element. <br> surfaceScale="***height***" (default 1) <br> diffuseConstant="***factor***" (must be non-negative; default 1)
feDisplacementMap     | scale="***displacement factor***" (default 0) <br> xChannelSelector="R"  "G"  "B"  "A" <br> yChannelSelector="R"  "G"  "B"  "A" <br> in2="***second input***"
feFlood               | flood.fill.color="***color specification***" <br> flood.fill.opacity="***value (0-1)***"
feGaussianBlur        | stdDeviation="***blur spread***" (larger is blurrier; default 0)
feImage               | xlink:href="***image source***"
feMerge               | container for feMergeNode elements
feMergeNode           | ***in1*** = "***intermediate result***"
feMorphology          | operator="erode" "dilate" <br> radius="***x-radius** **y-radius***" <br> radius="***radius***"
feOffset              | dx="***x offset***" (default 0)<br>dy="***y offset***" (default 0)
feSpecularLighting    | Container for a light source element.<br>surfaceScale="***height***" (default 1)<br>specularConstant="***factor***" (must be non-negative; default 1)<br>specularExponent="***exponent***" (range 1-128; default 1)
feTile                | tiles the ***in1*** layer
feTurbulence          | type="turbulence"  "fractalNoise" <br> baseFrequency="***x-frequency y-frequency***" <br> baseFrequency="***frequency***" <br> numOctaves="***integer***" <br> seed="***number***"
feDistantLight        | azimuth="***degrees***" (default 0)<br>elevation="***degrees***" (default 0)
fePointLight          | x="***coordinate***" (default 0)<br>y="***coordinate***" (default 0)<br>z="***coordinate***" (default 0)
feSpotLight           | x="***coordinate***" (default 0)<br>y="***coordinate***" (default 0)<br>z="***coordinate***" (default 0)<br>pointsAtX="***coordinate***" (default 0)<br>pointsAtY="***coordinate***" (default 0)<br>pointsAtZ="***coordinate***" (default 0)<br>specularConstant="***focus control***" (default 1)<br>limitingConeAngle="***degrees***"

