---
title: "Filter Colors"
output:
  html_document:
    toc: true
    theme: united
---

```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
  source('Util.R')
```


## Color Flooding


The feFlood filter is probably the simplest filter. It is used to flood a region with a given color. The feFlood filter is often combined with other filters and rarely used alone. In this section we use feFlood alone, but later sections will use it in combination with other filters.

A simple example of feFlood is
```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
  use(
    filter=filter( feFlood(flood.color='yellow') )
  )
)
```

The filter as shown above, assigns an filter element to a filter attribute directly.  This filter=filter(...) is a shortcut introduced in svgR for programming convienience. 

The more traditional approach is to place the filter as a child of a defs element and use a reference to the filter when when assigning to an attribute.

For example

```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
  defs(
    filter(id='myFlood',
           feFlood(flood.color='yellow')
    )
  ),
  use(
    filter="url(#myFlood)"
  )
)
```

The filter=filter(...) short cut in svgR does a similar construction in the output. However, for complex filters, where filter components are reused the tradional approach might be easier.

In this document we use the shortcut approach.

Our second example of feFlood is to illustrate that a filter has on a rectangular region on which it is effective. That region given by it's xy, wh coordinates.

```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
    text('10 semi-transparent floods', xy=c(5,20), stroke="black", font.size=20),
    rect(cxy=WH/2, wh=c(1,.2)*WH, fill='blue'),
    lapply(1:10, 
      function(i){g(
           filter=filter( xy=c(i*60,10),  wh=c(30,90), feFlood(flood.color='yellow', flood.opacity='.5'))
           ) 
      }
      )
)
```

Caution: feFlood floods the filters bounding box defined by the filter xy and wh. To flood the interior of a graphics, one needs
to use one of feComposite, feBlend, feMerge.

## Rotating Hues

The **feColorMatrix** is a filter primitive
that can be used to rotate color hues


Begining withcollection of red rectangles, we apply
a hue rotation to each. The first rectangle, (left most
rectangle) is rotated by 0 (degrees) and retains it's original
color. The last rectangle is rotated by 360 (degrees) and
so keeps it's original color. But the intermediate rectangles
have their colors remapped according to a color wheel.

```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
    lapply(seq(from=0, to=360, by=16), 
      function(i){
        rect(xy=c(i*2,10), wh=c(30,32), fill='red',
           filter=filter( 
             feColorMatrix(
               type="hueRotate", values=i
             )  
           )
        )
      }
    )
)
```

This time we illustrate hue rotation applied to a rectangles filled
with a linear gradient. The left most is the original, and each step
to the right rotates that gradient by 16 degrees on the color wheel.

```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
    lapply(seq(from=0, to=360, by=16), 
      function(i){
        rect(xy=c(i*2,10), wh=c(30,96), 
          fill=linearGradient( xy1=c(0,0), xy2=c(0,1),
            colors=c('red','yellow','green','blue')),
           filter=filter( 
             feColorMatrix(
               type="hueRotate", values=i
             )  
           )
        )
      }
    )
)
```

## Color Saturation


Color saturation is a measure of preceived color intensity.
For example, light from a laser at high intensity  having 
single wave-length as it's color
is considered to be saturated. 

The filter primitive element, **feColorMatrix** 
that can be used to modify the saturation of a given
color by *desaturating*  the original image. (Remove color)
The amount of desaturation is specified by a value between
0 and 1, with 0 being no desaturation and 1 being totally desaturated.

Applying satuate to a series of red rectangles we have:

```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
    lapply(seq(from=1, to=0, by=-.05), 
      function(i){
        rect(xy=c(i*800,10), wh=c(32,32), fill="red",
           filter=filter( 
             feColorMatrix(
               type="saturate", values=i
             )  
           )
        )
      }
    )
)
```
The left-most rectanle  was given a desaturation value of 1, and the 
right-most rectangle was given value of 0.


Again, applying hue saturate to a color spectrum 
generated by a linear gradient. we have

```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
    lapply(seq(from=1, to=0, by=-.05),
      function(i){
        rect(xy=c(i*800,10), wh=c(32,96), fill=
          linearGradient( xy1=c(0,0), xy2=c(0,1),
            colors=c('red','yellow','green','blue')),
           filter=filter( 
             feColorMatrix(
               type="saturate", values=i
             ) 
           )
        )
      }
    )
)
```

## Applying  a color matrix
(!!To fix)

Applying satuate to a red rectangle
```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
    lapply(seq(from=0, to=1, by=.05), 
      function(i){
        rect(xy=c(i*800,10), wh=c(32,32), fill="red",
           filter=filter( 
             feColorMatrix(
               type="matrix", 
               values=paste(runif(25,-1,1),collapse=" ")
             )  
           )
        )
      }
    )
)
```


