---
title: "feComponentTransfer"
output:
  html_document:
    toc: true
    theme: united
---


```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```


## A Simple Example

To create textures, we often employ the **feTurbulance** filter
primitive.
The feTurlence supports two *types*:

- type='turbulence' (default)
- type ='fractal'


A simple example of the former is:

```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
      use(
        filter=filter( filterUnits="userSpaceOnUse",
          feTurbulence(baseFrequency=0.1, numOctaves=2 ) 
        )
      )
)
```

Here we have specified  

- *baseFrequency*=0.1
- *numOctaves*=2

## baseFrequency

The effect of base frequence is seen here

```{r, echo=T, results="asis"}
WH=c(800, 880) # window rect
N<-99
bfreq<-1:N * .003
svgR( wh=WH,
  text('Base frequencey Effects', xy=c(WH[1]/2, 20)),
  lapply( 1:N, function(i){
    x<-((i-1)%%11)*70
    y<-(floor((i-1)/11))*90+50
    g(
      use( filter=filter( xy=c(x,y), wh=c(60,60), filterUnits="userSpaceOnUse",
          feTurbulence(baseFrequency=bfreq[i], numOctaves=2) 
        ) ),
      text(bfreq[i], xy=c(x,y-10))

    )
  })
)
```


The base frequence can also be expressed along the x or y axis independently by seting baseFrequency=c(xFreq, yFreq)

```{r, echo=T, results="asis"}
WH=c(800, 800) # window rect
baseFeqRng=1:6
numOctRng=1:6
indx<-expand.grid(baseFeqRng, numOctRng)
xp<-function(i)(i-1)*130+50
yp<-function(j)(j-1)*100+50
svgR( wh=WH,
    text("x-baseFrequency", xy=c(WH[1]/2,10)),
    text("y-baseFrequency", xy=c(10,WH[2]/2), writing.mode='tb'),
    lapply(baseFeqRng, function(i){text(0.005*(i-1), xy=c(xp(i),35))}),
    lapply(numOctRng, function(j){text(0.005*(j-1), xy=c(20,yp(j)+50))}),
    apply(indx, 1, function(z){
      i=z[1];j=z[2]
      use(filter=filter( xy=c(xp(i),yp(j)),  wh=c(120,90), filterUnits="userSpaceOnUse",
            feTurbulence(baseFrequency=0.005*c(i-1,j-1), numOctaves=1) 
          )
        ) 
      })
)
```

## Octaves

To see the effect of baseFrequency vs numOctaves, we iterate

```{r, echo=T, results="asis"}
WH=c(800, 1200) # window rect
baseFeqRng=1:5
numOctRng=1:10
indx<-expand.grid(baseFeqRng, numOctRng)
xp<-function(i)(i-1)*140+50
yp<-function(j)(j-1)*100+50
svgR( wh=WH,
    text("baseFrequency", xy=c(WH[1]/2,10)),
    text("numOctaves", xy=c(10,WH[2]/2), writing.mode='tb'),
    lapply(baseFeqRng, function(i){text(.02*i, xy=c(xp(i),35))}),
    lapply(numOctRng, function(j){text(2*j-1, xy=c(20,yp(j)+50))}),
    apply(indx, 1, function(z){
      i=z[1];j=z[2]
      use(filter=filter( xy=c(xp(i),yp(j)),  wh=c(130,90), filterUnits="userSpaceOnUse",
            feTurbulence(baseFrequency=.02*i, numOctaves=2*j-1 )
          )
        ) 
      })
)
```

## Types of feTurbulence

The type attribute of feTurbulence can have one of two values:

- turbulence (the default)
- fractalNoise

We have seen the turblance type in action, fractalNoise is similar, 
but the appears to be softer


```{r, echo=T, results="asis"}
WH=c(800, 150) # window rect
svgR( wh=WH,
      use(
        filter=filter( xy=c(50,70),wh=c(150,60), filterUnits="userSpaceOnUse",
          feTurbulence(type='turbulence', baseFrequency="0.1", numOctaves="2" ) 
        )
      ),
            use(
        filter=filter( xy=c(250,70),wh=c(150,60), filterUnits="userSpaceOnUse",
          feTurbulence(type='fractalNoise', baseFrequency="0.1", numOctaves="2" ) 
        )
      ),
      text("turbulence", xy=c(100,65)),
      text("fractalNoise", xy=c(300,65))

)
```


## Combining feTurbulance

By using feTurbulance, feComponentTransfer and feColorMatrix we can discover some  intersting
effects.  For example, define the following filter

### feTurbulance + feComponentTransfer
```{r, echo=T, results='asis'}
funkyFilter %<c-% function(id, baseFrequency=.01, numOctaves=3, slope=c(4,4,4), seed=100){
  return(filter( id=id, 
    feTurbulence(baseFrequency=baseFrequency, numOctaves=numOctaves, seed=100),
    feComponentTransfer(
      feFuncR(type="linear", slope=slope[1], intercept=-1),
      feFuncG(type="linear", slope=slope[2], intercept=-1),
      feFuncB(type="linear", slope=slope[3], intercept=-1),
      feFuncA(type="linear", slope=0, intercept=1)
    ),
    feColorMatrix(type="saturate") 
  ))
} 
```

To see this filter function at work:
```{r, echo=T, results='asis'}
bf=3 # This is a kludge to force bf
library(svgR)
WH=c(800, 800) # window rect
N<-4
dH<-WH[2]/(N+1)
y<-0:(N-1)*dH
ww<-WH[1]-40
svgR( wh=WH,
  lapply(1:N, function(i){
    id=paste0("funky1-",i/N)
    bf<<-.02/i # Very bad kludge
    txt<-paste0('funky filter base frequency=', bf)
    g(
      text(txt, xy=c(20, y[i]+35) ),
      rect(xy=c(20,y[i]+60), wh=c(ww, dH-60 ), 
         filter = funkyFilter(id,  bf))
    )      
  })   
)
```

### feTurbulance + feComponentTransfer + feConvolveMatrix
We can sharpen this using an **feConvolveMatrix**. 

```{r, echo=T}
sharpAndFunkyFilter %<c-% function(baseFrequency=.01, numOctaves=3, slope=c(4,4,4), seed=100){
  filter( 
    feTurbulence(baseFrequency=baseFrequency, numOctaves=numOctaves, seed=100),
    feComponentTransfer(
      feFuncR(type="linear", slope=slope[1], intercept=-1),
      feFuncG(type="linear", slope=slope[2], intercept=-1),
      feFuncB(type="linear", slope=slope[3], intercept=-1),
      feFuncA(type="linear", slope=0, intercept=1)
    ),
    feColorMatrix(type="saturate"),
    feConvolveMatrix( order=3, kernelMatrix=matrix(c(1,-1,1,-1,-.1,-1,1,-1,1), 3,3))
  )
} 
```
And the sharpened result is
```{r, echo=T, results='asis'}
library(svgR)
WH=c(800, 800) # window rect
N<-4
dH<-WH[2]/(N+1)
y<-0:(N-1)*dH
ww<-WH[1]-40
svgR( wh=WH,
  lapply(1:N, function(i){
    id=paste0("funky2-",i/N)
    bf=.02/(i)
    g(
      text(paste0('sharp and funky filter base frequency=', bf), xy=c(20, y[i]+35) ),
      rect(xy=c(20,y[i]+60), wh=c(ww, dH-60 ), 
         filter = sharpAndFunkyFilter( baseFrequency = bf))
    )      
  })
    
)
```


Messing with the base frequency gives a different texture

```{r, echo=T, results='asis'}
library(svgR)
WH=c(800, 800) # window rect
N<-4
dH<-WH[2]/(N+1)
y<-0:(N-1)*dH
ww<-WH[1]-40
svgR( wh=WH,
  lapply(1:N, function(i){
    id=paste0("funky2-",i/N)
    bf=c(.1,.1/(i))
    g(
      text(paste0('sharp and funky filter base frequency=', paste(bf, collapse=",") ), xy=c(20, y[i]+35) ),
      rect(xy=c(20,y[i]+60), wh=c(ww, dH-60 ), 
         filter = sharpAndFunkyFilter( baseFrequency = bf))
    )      
  })
    
)
```


## Wood

To create a wood texture we use

- feDisplacementMap
- feTurbulence
- feFlood
- feColorMatrix

```{r, echo=T, results='asis'}
WH=c(800, 200) # window rect
m<-rbind( c(1,0,0,0,0), c(0,-1,0,0,0), c(0,0,-1,0,0), c(1,-1,1,0,0))
svgR( wh=WH, 
  rect( cxy=WH/2, wh=WH, stroke.width=2, stroke='black',
    filter = filter(
      feDisplacementMap(  scale=9, xChannelSelector="R", yChannelSelector="B",
        in1=feMerge(
          feMergeNode(in1= feFlood(xy=c(0,0), wh=WH,
          flood.color="#CCAC6C") ),
          feMergeNode(
            in1=feColorMatrix( type="matrix", values=m , 
              in1=feTurbulence( baseFrequency=c(.008,.25), numOctaves=1)
            )
          )
        ),
        in2=feTurbulence( baseFrequency=c(.04,.2), numOctaves=1) 
      )
    ),
    clip.path=clipPath(
      text("Wood", font.size=250, cxy=WH/2, font.face='bold')
    )
  )
)
```

<br><br>

## Stucco

```{r, echo=T, results='asis'}
WH=c(800, 200) # window rect
m<-rbind( c(3,0,0,0,0), c(0,0,0,0,0), c(0,0,-1,0,0), c(1,1,1,0,0))
svgR( wh=WH, 
  rect( cxy=WH/2, wh=WH, stroke.width=2, stroke='black',
    filter = filter(
      feDisplacementMap(  scale=9, xChannelSelector="R", yChannelSelector="B",
        in1=feMerge(
          feMergeNode(in1= feFlood(xy=c(0,0), wh=WH,
          flood.color="#CCDCDC"), flood.opacity=.5 ),
          feMergeNode(
            in1=feColorMatrix( type="matrix", values=m , 
              in1=feTurbulence( baseFrequency=c(.2,.3), numOctaves=1)
            )
          )
        ),
        in2=feTurbulence( baseFrequency=c(.006,.004), numOctaves=2) 
      )
    ),
    clip.path=clipPath(
      text("Stucco", font.size=250, cxy=WH/2, font.face='bold')
    )
  )
)
```


```{r, echo=T, results='asis'}
WH=c(800, 200) # window rect
m<-rbind( c(3,0,0,0,0), c(0,0,0,0,0), c(0,0,-1,0,0), c(1,1,1,0,0))
svgR( wh=WH, 
  rect( cxy=WH/2, wh=WH, stroke.width=2, stroke='black',
    filter = filter(
      feDisplacementMap(  scale=9, xChannelSelector="R", yChannelSelector="B",
        in1=feMerge(
          feMergeNode(in1= feFlood(xy=c(0,0), wh=WH,
          flood.color="#FFDCDC"), flood.opacity=.5 ),
          feMergeNode(
            in1=feColorMatrix( type="matrix", values=m , 
              in1=feTurbulence( baseFrequency=c(.5,.1), numOctaves=4)
            )
          )
        ),
        in2=feTurbulence( baseFrequency=c(.006,.004), numOctaves=2) 
      )
    ),
    clip.path=clipPath(
      text("Stucco", font.size=250, cxy=WH/2, font.face='bold')
    )
  )
)
```

## Jade
```{r, echo=T, results='asis'}
WH=c(800, 200) # window rect
m<-rbind( c(3,0,0,0,0), c(0,0,0,0,0), c(0,0,-1,0,0), c(1,1,1,0,0))
svgR( wh=WH, 
  rect( cxy=WH/2, wh=WH, stroke.width=2, stroke='black',
    filter = filter(
      feDisplacementMap(  scale=9, xChannelSelector="R", yChannelSelector="B",
        in1=feMerge(
          feMergeNode(in1= feFlood(xy=c(0,0), wh=WH,
          flood.color="#CFFCDC"), flood.opacity=.5 ),
          feMergeNode(
            in1=feColorMatrix( type="matrix", values=m , 
              in1=feTurbulence( baseFrequency=c(.1,.1), numOctaves=4)
            )
          )
        ),
        in2=feTurbulence( baseFrequency=c(.06,.004), numOctaves=2) 
      )
    ),
    clip.path=clipPath(
      text("Jade", font.size=250, cxy=WH/2, font.face='bold')
    )
  )
)
```

## Polished

```{r, echo=T, results='asis'}
WH=c(800, 300) # window rect
m<-rbind( c(3,0,0,0,0), c(0,0,0,0,0), c(0,0,-1,0,0), c(1,1,1,0,0))
svgR( wh=WH, 
  rect( cxy=WH/2, wh=WH, stroke.width=2, stroke='black',
    filter = filter(
      feDisplacementMap(  scale=9, xChannelSelector="R", yChannelSelector="B",
        in1=feMerge(
          feMergeNode(in1= feFlood(xy=c(0,0), wh=WH,
          flood.color="#CFFCDC"), flood.opacity=.5 ),
          feMergeNode(
            in1=feColorMatrix( type="matrix", values=m , 
              in1=feTurbulence( baseFrequency=c(.01,.01), numOctaves=1)
            )
          )
        ),
        in2=feTurbulence( baseFrequency=c(.006,.04), numOctaves=2) 
      )
    ),
    clip.path=clipPath(
      text("Shiny", font.size=250, cxy=WH/2, font.face='bold')
    )
  )
)
```
