---
title: "ChangingShapes"
output:
  html_document:
    toc: true
    theme: united
---


```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

### On Fire
Another approach to shape modification is to perform some
kind of ***warping** of a image shape. The  **feDisplacementMap** is filter primitive, which provides the reshaping of one image (in1) using the colors provided by a second image (in2) to dictate the amount of displacement.

Here we illustrate using the feDisplacementMap filter applied to a one of two yellow rectangles. The displacment is being supplied by the output of an **feTurbulence** filter.

```{r, echo=T, results="asis"}
library(svgR)
WH=c(800, 200) # window rect
svgR( wh=WH, 
      
      text("feDisplacementMap on fire", xy=c(20,50), stroke='black', fill='white', font.size=20),
#      text('Data Denier 4 Pres'
      rect( cxy=WH/2, wh=c(600,20), fill='yellow', 
           stroke='red', stroke.width=2,
           filter=
             filter( xy=c(0,0), filterUnits='userSpaceOnUse',
                feDisplacementMap(
                  in1="SourceGraphic",
                  in2=feTurbulence( baseFrequency=0.05, numOctaves=3), 
                  scale=30,
                  yChannelSelector="R"                )
          )
      ),
       rect( cxy=WH/2+c(0,20), wh=c(600,20), fill='yellow', stroke='none')
)
```

<br><br>





### Seville

In the last section, we applied the  to a simple rectangle
More often, the  **feDisplacementMap** is applied to images. In this example we choose
an jpg of the palace in Seville.
Here is the jpg image without any filtering.

```{r, echo=T, results="asis"}
library(svgR)
WH=c(800, 600) # window rect
svgR( wh=WH,       
      image(cxy=WH/2, wh=WH, xlink.href="./IMG_4703.JPG"),
      text("Seville", xy=c(600,100), font.size=46)
)
```
<br><br>

Next we apply a feDisplacement filter

```{r, echo=T, results="asis"}
library(svgR)
WH=c(800, 600) # window rect
svgR( wh=WH, 
      image(cxy=WH/2, wh=WH, xlink.href="./IMG_4703.JPG",
           filter=filter(
                feDisplacementMap(
                  in1="SourceGraphic",
                  in2=feTurbulence( baseFrequency=0.066666666666667, numOctaves=3), 
                  scale=15,
                  yChannelSelector="R"
                )
           )
      ),
      text("Seville", xy=c(600,100), font.size=46)
)
```

<br><br>

To apply the filter to the text we simply include both a group

```{r, echo=T, results="asis"}
library(svgR)
WH=c(800, 600) # window rect
svgR( wh=WH, 
      g(
        image(cxy=WH/2, wh=WH, xlink.href="./IMG_4703.JPG"),
        text("Seville", xy=c(600,100), font.size=46),
        filter=filter(
                feDisplacementMap(
                  in1="SourceGraphic",
                  in2=feTurbulence( baseFrequency=0.066666666666667, numOctaves=3), 
                  scale=15,
                  yChannelSelector="R"
                )
           )
      )
)
```

<br><br>

And again, but with a larger scale factor

```{r, echo=T, results="asis"}
library(svgR)
WH=c(800, 600) # window rect
svgR( wh=WH, 
      image(cxy=WH/2, wh=WH, xlink.href="./IMG_4703.JPG",
           filter=filter(
                feDisplacementMap(
                  in1="SourceGraphic",
                  in2=feTurbulence( baseFrequency=0.0066666666666667, numOctaves=3), 
                  scale=150,
                  yChannelSelector="R"
                )
           )
      ),
      text("Decay", xy=c(20,30), font.size=26)
)
```

<br><br>

### Data Corruption
<br><br>

Finally, we apply a displacement to some text.

```{r, echo=T, results="asis"}
library(svgR)
WH=c(800, 400) # window rect
svgR( wh=WH, 
      g(  lapply(1:50, function(i)text("Data", cxy=runif(2)*.8*WH, 
                          fill=rrgb(),  font.size=sample(10:30,1))),
          filter=filter( xy=c(0,0), filterUnits='userSpaceOnUse',
              feDisplacementMap(
                in1="SourceGraphic",
                in2=feTurbulence( baseFrequency=0.00666666666666667, numOctaves=3), 
                scale=50,
                yChannelSelector="R",
                xChannelSelector="B"
              )
          )
      ),
      text("Data corruption", xy=c(20,30), font.size=26)
)
```

<br><br>

