---
title: "Filter Attribute"
output:
  html_document:
    toc: true
    theme: united
---



```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
if(!exists("newId")){
  source('Util.R')
}
```

##Overview


The Filter attribute can only have a filter element or a reference to a filter element as it's value.
A Filter element is a container for fe-elements, which in this document we will refer to as .
The fe-elements together form a directed graph, with root, 
whose construction is discussed in the section [Combining Filter Primitives]

That ensemble is used to
create interesting effects by modifying existing images. 

Many filter primitives  require  input of image data 
(some require two), and these are supplied via the in1 and in2 attributes. 
Two notable exceptions are feImage and feFlood.



todo: diagrams


todo: 

- feFunc***X*** 
- feTile



### The  In Attribute of  Filter components

As we noted before, filter components may have one or two ***in*** attributes: in1, in2.
The in1 and in2 attributes can take the values SourceGraphic, SourceAlpha, BackgroundImage, BackgroundAlpha, FillPaint, StrokePaint, or a reference to another primitive's result. The behaviours for each are as follows:

- SourceGraphic represents the original element to which the filter is applied (as full RGBA image data)
- SourceAlpha represents just the alpha channel of the original element to which the filter is applied
- BackgroundImage represents the area directly underneath the filter region
- BackgroundAlpha represents just the alpha channel of the area underneath the filter region
- FillPaint represents whatever fill was applied to the element in question
- StrokePaint represents whatever stroke was applied to the element
- Any other value should be a reference equal to the result attribute for any preceding filter primitive. In this case, the output of the referenced primitive will be used as the specified input.





### Filter Components Summary

Element               |                Attributes                                  
----------------------|------------------------------------------------------------
feBlend               | in2="***second source***" <br> mode="normal"  "multiply"  "screen"  "darken"  "lighten"
feColorMatrix         | type="matrix"  "saturate"  "hueRotate"  "luminanceToAlpha" <br> values="***matrix values***"  "***saturation value (0-1)***"  "***rotate degrees***"
feComponentTransfer   | container for <br>feFuncR, <br>feFuncG, <br>feFuncB, and <br>feFuncA elements.
feFunc***X***         | type="identity"  "table"  "discrete"  "linear"  "gamma" <br> tableValues="***intervals for table, steps for discrete***" <br> slope="***linear slope***" <br> intercept="***linear intercept***" <br> amplitude="***gamma amplitude***" <br> exponent="***gamma exponent***" <br> offset="***gamma offset***" 
feComposite           | in2="***second source***" <br> operator="over"  "in"  "out"  "atop"  "xor"  "arithmetic" <br> The following attributes are used with arithmetic: <br> k1="***factor for in1,in2***" <br> k2="***factor for in1***" <br> k3="***factor for in2***" <br> k4="***additive offset***"
feConvolveMatrix      | order="***columns rows***" (default 3 by 3) <br> kernel="***values***" <br> bias="***offset value***"
feDiffuseLighting     | Container for a light source element. <br> surfaceScale="***height***" (default 1) <br> diffuseConstant="***factor***" (must be non-negative; default 1)
feDisplacementMap     | scale="***displacement factor***" (default 0) <br> xChannelSelector="R"  "G"  "B"  "A" <br> yChannelSelector="R"  "G"  "B"  "A" <br> in2="***second input***"
feFlood               | flood.fill.color="***color specification***" <br> flood.fill.opacity="***value (0-1)***"
feGaussianBlur        | stdDeviation="***blur spread***" (larger is blurrier; default 0)
feImage               | xlink:href="***image source***"
feMerge               | container for feMergeNode elements
feMergeNode           | ***in1*** = "***intermediate result***"
feMorphology          | operator="erode" "dilate" <br> radius="***x-radius** **y-radius***" <br> radius="***radius***"
feOffset              | dx="***x offset***" (default 0)<br>dy="***y offset***" (default 0)
feSpecularLighting    | Container for a light source element.<br>surfaceScale="***height***" (default 1)<br>specularConstant="***factor***" (must be non-negative; default 1)<br>specularExponent="***exponent***" (range 1-128; default 1)
feTile                | tiles the ***in1*** layer
feTurbulence          | type="turbulence"  "fractalNoise" <br> baseFrequency="***x-frequency y-frequency***" <br> baseFrequency="***frequency***" <br> numOctaves="***integer***" <br> seed="***number***"
feDistantLight        | azimuth="***degrees***" (default 0)<br>elevation="***degrees***" (default 0)
fePointLight          | x="***coordinate***" (default 0)<br>y="***coordinate***" (default 0)<br>z="***coordinate***" (default 0)
feSpotLight           | x="***coordinate***" (default 0)<br>y="***coordinate***" (default 0)<br>z="***coordinate***" (default 0)<br>pointsAtX="***coordinate***" (default 0)<br>pointsAtY="***coordinate***" (default 0)<br>pointsAtZ="***coordinate***" (default 0)<br>specularConstant="***focus control***" (default 1)<br>limitingConeAngle="***degrees***"

