---
title: "Filter Attribute"
output:
  html_document:
    toc: true
    theme: united
---



```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```


##Overview of Filters

###Introduction


In the previous section we explored how the *fill* attribute can be applied to text and shapes.
Another
attribute for text and shapes, is the *filter* attribute. 
*Filter *attributes  may take as  values  a **filter element** or a *reference* 
to a **filter element**. 
A **Filter element** is a container for **filter effect elements**, 
also called **filter primitives** or **fe-elements**
. The **fe-elements**  feed into each other to 
form  a directed graph with a root. Typically, the fe elements are linked together using the
*in1*, *in2* attributes. One might visualize a filter to look something like



```{r, echo=FALSE, results="asis"}
WH<-c(650,680)

#Defined by mouse: edit with care!
ptR<-list(
  arrows=matrix(
    c(c( 150.5,205 ),c( 150.5,150 ),c( 265.5,152 ),c( 276.5,108 ),c( 13.5,12 ),c( 451.5,204 ),c( 443.5,153 ),c( 347.5,154 ),c( 329.5,112 ),c( 26.5,9 ),c( 452.5,302 ),c( 452.5,272 ),c( 453.5,245 ),c( 21.5,22 ),c( 300.5,448 ),c( 298.5,396 ),c( 452.5,397 ),c( 453.5,344 ),c( 33.5,20 ),c( 150.5,601 ),c( 151.5,551 ),c( 279.5,549 ),c( 279.5,494 ),c( 33.5,33 ),c( 450.5,601 ),c( 451.5,549 ),c( 332.5,553 ),c( 327.5,494 )),
  2,),
  labels=matrix(
    c(c( 201.5,138 ),c( 400.5,139 ),c( 438.5,279 ),c( 372.5,385 ),c( 219.5,539 ),c( 400.5,537 )),
  2,),
  uniboxes=matrix(
    c(c( 301.5,62 ),c( 151.5,204 ),c( 452.5,203 ),c( 452.5,303 ),c( 300.5,450 ),c( 147.5,601 ),c( 453.5,603 )),
  2,)
)


unifboxes%<c-%function(x=c(), txt="", wh=c(100,30), rxy=c(20,20), ...){ 
  if(length(x)==0){
    return(NULL)
  }
  n<-ncol(x)
  if(length(txt)<n){   
    txt<-c(txt, paste0(length(txt):c))
  }
  lapply(1:n, function(i){
    xy<-x[,i]-c(.5,0)*wh
    cxy<-xy+wh/2
    g(
      rect(xy=xy, wh= wh, rxy=rxy, ...),
      if(nchar(txt[i])>0){
        text(cxy=cxy, txt[i],  ...)
      }  else {
        NULL
      }
    )
  })
} 



roundedPolyLine%<c-%function(x, r=5, ...){
  nx<-length(x)
  if(nx<=4){
    return(nx)
  }
  n<-ncol(x)
  if(n==2){
    return(path(d=c("M", x[,1], "L", x[,n]), ...))
  }

  dv<-x[,-n]-x[,-1]
  d<-apply( dv ,2, function(x)sqrt(sum(x)^2) )
  lambda<-sapply(r/d, function(x)min(x,.5))
  
  m<- n-1
  if(m==2){
    mA<-matrix(c(lambda[1],1-lambda[1]),2,1)
    mB<-matrix(c(1-lambda[2],lambda[2]),2,1)
  } else {
    mA<- rbind(diag(lambda[-m]),0) + rbind(0,diag(1-lambda[-m]))
    mB<- rbind(0,diag(lambda[-1])) + rbind(diag(1-lambda[-1]),0)
  } 
  a<-x[,-n]%*%mA
  b<-x[,-1]%*%mB   
  rL<-rep("L", n-2)
  rQ<-rep("Q", n-2)
  if(m==2){
    rr<-c(rL,a,rQ,x[,2],b)
  } else {
    rr<-rbind(rL, a, rQ, x[,-c(1,n)], b)
  }
  path(d=c("M", x[,1], rr, "L", x[,n]),  ...)
  }

roundedPolyLineArrays%<c-%function(p, ...){
  if(length(p)<4){
    return(NULL)
  }
  n<-ncol(p)
  brks<-apply(p,2, function(x){ max(x)<50})
  brkss<-(1+cumsum(brks))*(!brks)
  if(max(brkss)==0){
    return(NULL)
  }
  
  lapply(1:max(brkss),function(i){
    p_i<-p[,brkss==i]
    roundedPolyLine(p_i, ...)
  }) 
}



arrowHeadId<-autoId()
arrowHeadUrl<-paste0("url(#",arrowHeadId,")")

btxt<-c("feBlend",  "SourceGraphic", "feGaussianBlur",
    "feOffset", "feComposite", "feFlood", "SourceGraphic" )

labels<-c("in1","in2","in1", "in1", "in1","in2")
svgR(wh=WH, 
     #your custom code goes here
     defs(
       marker(id=arrowHeadId, viewBox=c(0, 0, 10, 10), refXY=c(9,5), 
              markerWidth=6, markerHeight=6, orient="auto",
              path( d=c("M", 0, 0, "L", 10, 5, "L", 0, 10, "z"),
              fill='darkgreen'
              )
            )
     ),
    #  boxes(ptR$boxes, txt=btxt, stroke="black", 
    #     fill='lightblue', rxy=c(20,20)
    #  ),
    #rect(xy=c(50,25), wh=c(500,650), stroke='black',
        #fill='lightblue'),
        #text(cxy=c(100,50),'filter',stroke='black',
        #font.size=50),
     unifboxes(ptR$uniboxes, txt=btxt, wh=c(150,40), 
         stroke="darkgreen",
        fill="white"),
     roundedPolyLineArrays(ptR$arrows, r=30, 
                          stroke.width=3, stroke='darkgreen', 
                          fill="none",
                          marker.end=arrowHeadUrl
     ),
     lapply(1:ncol(ptR$labels), function(i){
        text(cxy=ptR$labels[,i], labels[i], stroke='darkgreen')
     })
     
     
     
)
```

One distinct difference between fills and filters, is that a fill
is restricted to the interior, while filters are not.

**Note** Technically this is not a tree, since the *SourceGraphics* label appears muliple places.

**Note** Pure *SVG* allows only references, but *svgR* allows to use
the **filter element** directly as the value of a *filter attribute*.




### The In Attributes of  Filter components

Most **fe-element** (filter primitive) require an input specification, which 
is preformed by assigning
values to their *in* attributes: *in1* and *in2*.


(Two notable exceptions to this are the fe-elements: **feImage** and **feFlood**.)

Values for the in *attributes* are given by the following table

Value | Description
----------------------|------------------------------------------------------------
*SourceGraphic* |represents the original element to which the filter is applied (as full RGBA image data)
*SourceAlpha* | represents just the alpha channel of the original element to which the filter is applied
*BackgroundImage* | represents the area directly underneath the filter region
*BackgroundAlpha* | represents just the alpha channel of the area underneath the filter region
*FillPaint* | represents whatever fill was applied to the element in question
*StrokePaint* | represents whatever stroke was applied to the element
a reference | Any other value should be a reference equal to the result attribute for any preceding filter primitive. In this case, the output of the referenced primitive will be used as the specified input.





### The fe-Elements (Filter primitives)


Element               |                Attributes                                  
----------------------|------------------------------------------------------------
feBlend               | in2="***second source***" <br> mode="normal"  "multiply"  "screen"  "darken"  "lighten"
feColorMatrix         | type="matrix"  "saturate"  "hueRotate"  "luminanceToAlpha" <br> values="***matrix values***"  "***saturation value (0-1)***"  "***rotate degrees***"
feComponentTransfer   | container for <br>feFuncR, <br>feFuncG, <br>feFuncB, and <br>feFuncA elements.
feFunc***X***         | type="identity"  "table"  "discrete"  "linear"  "gamma" <br> tableValues="***intervals for table, steps for discrete***" <br> slope="***linear slope***" <br> intercept="***linear intercept***" <br> amplitude="***gamma amplitude***" <br> exponent="***gamma exponent***" <br> offset="***gamma offset***" 
feComposite           | in2="***second source***" <br> operator="over"  "in"  "out"  "atop"  "xor"  "arithmetic" <br> The following attributes are used with arithmetic: <br> k1="***factor for in1,in2***" <br> k2="***factor for in1***" <br> k3="***factor for in2***" <br> k4="***additive offset***"
feConvolveMatrix      | order="***columns rows***" (default 3 by 3) <br> kernel="***values***" <br> bias="***offset value***"
feDiffuseLighting     | Container for a light source element. <br> surfaceScale="***height***" (default 1) <br> diffuseConstant="***factor***" (must be non-negative; default 1)
feDisplacementMap     | scale="***displacement factor***" (default 0) <br> xChannelSelector="R"  "G"  "B"  "A" <br> yChannelSelector="R"  "G"  "B"  "A" <br> in2="***second input***"
feFlood               | flood.color="***color specification***" <br> flood.opacity="***value (0-1)***"
feGaussianBlur        | stdDeviation="***blur spread***" (larger is blurrier; default 0)
feImage               | xlink:href="***image source***"
feMerge               | container for feMergeNode elements
feMergeNode           | ***in1*** = "***intermediate result***"
feMorphology          | operator="erode" "dilate" <br> radius="***x-radius** **y-radius***" <br> radius="***radius***"
feOffset              | dx="***x offset***" (default 0)<br>dy="***y offset***" (default 0)
feSpecularLighting    | Container for a light source element.<br>surfaceScale="***height***" (default 1)<br>specularConstant="***factor***" (must be non-negative; default 1)<br>specularExponent="***exponent***" (range 1-128; default 1)
feTile                | tiles the ***in1*** layer
feTurbulence          | type="turbulence"  "fractalNoise" <br> baseFrequency="***x-frequency y-frequency***" <br> baseFrequency="***frequency***" <br> numOctaves="***integer***" <br> seed="***number***"
feDistantLight        | azimuth="***degrees***" (default 0)<br>elevation="***degrees***" (default 0)
fePointLight          | x="***coordinate***" (default 0)<br>y="***coordinate***" (default 0)<br>z="***coordinate***" (default 0)
feSpotLight           | x="***coordinate***" (default 0)<br>y="***coordinate***" (default 0)<br>z="***coordinate***" (default 0)<br>pointsAtX="***coordinate***" (default 0)<br>pointsAtY="***coordinate***" (default 0)<br>pointsAtZ="***coordinate***" (default 0)<br>specularConstant="***focus control***" (default 1)<br>limitingConeAngle="***degrees***"

