---
title: "filter-ShinyBevel"
output:
  html_document:
    toc: true
    theme: united
---



```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
  source('Util.R')
```

Shiny beveling is a popular example of what can be accomplished
using the right lighting.

## Shining with Lighting
We can 
shiny bevel
```{r, echo=T, results="asis"}
WH=c(800, 200) # window rect
svgR( wh=WH,
  text( "R Rocks!", cxy=c(400,100), fill="darkblue", font.size=190, font.face='bold',
    filter = filter(
      feMerge( 
         feMergeNode(in1="SourceGraphic"),
         feMergeNode(
           in1=feComposite( operator='in',
                in1=feSpecularLighting( surfaceScale=6,
                                        specularConstant=1,
                                        specularExponent=30,
                                        lighting.color="white",
                                        in1=feGaussianBlur(
                                          stdDeviation=5,
                                          in1="SourceAlpha"),
                                        fePointLight(xyz=c(40,-30,200))
                                        ),
                in2="SourceAlpha"
                )
           )
         )
    )
  )
)
```

## Wrapping a Filter
But we can also wrap this filter as a user defined filter element by

```{r, echo=T}
sbevelFilter %<c-% function(lighting.color="white", xyz=c(40,-30,200) ){
  filter(
      feMerge( 
         feMergeNode(in1="SourceGraphic"),
         feMergeNode(
           in1=feComposite( operator='in',
                in1=feSpecularLighting( surfaceScale=6,
                                        specularConstant=1,
                                        specularExponent=30,
                                        lighting.color=lighting.color,
                                        in1=feGaussianBlur(
                                          stdDeviation=5,
                                          in1="SourceAlpha"),
                                        fePointLight(xyz=xyz)
                                        ),
                in2="SourceAlpha"
                )
           )
         )
    )
}
```

and then  use as like so
```{r, echo=T, results="asis"}
WH=c(800, 200) # window rect
svgR( wh=WH,
  text( "R Rocks!", cxy=c(400,100), fill="darkblue", font.size=190, font.face='bold',
    filter = sbevelFilter()
  )
)
```

## Shing S

In this section we create several shiny letters S using the sbevelFilter

### Silver Metallic Shiny S

For a  shiny bevel letter S, with grey scale we simple 
add a circle and change the text to "S" , both with a black stroke.




```{r, echo=T, results="asis"}
WH=c(800, 250) # window rect
svgR( wh=WH,
  g( 
    text( "S", cxy=WH/2, fill="black", font.size=100, 
          font.family="serif", stroke.width=3),
    circle(cxy=WH/2, r=39, stroke.width=10, fill='none'  ),
     stroke='black', filter=sbevelFilter()
  )
)
```

### Metallic Shiny S's
For variety, we can vary color of the text stroke

```{r, echo=T, results="asis"}
WH=c(800, 250) # window rect
colors<-c("purple", "darkblue", "darkgreen", "brown", "darkorange")
svgR( wh=WH,
  lapply(1:5, function(i){
    cxy=WH*c(i*.15,.5)
    g( 
      text( "S", cxy=cxy, fill=colors[i], font.size=100, 
          font.family="serif", stroke.width=3),
      circle(cxy=cxy, r=39, stroke.width=10, fill='none'  ),
        stroke='black', filter=sbevelFilter()
    )
  })
)
```

### Metallic Shiny S's Lighting

And, we can also vary color of the light

```{r, echo=T, results="asis"}
WH=c(800, 250) # window rect
colors<-c("purple", "darkblue", "darkgreen", "brown", "darkorange")
lighting.color<-c("lightblue", "yellow", "pink", "orange", "lightgreen")

svgR( wh=WH,
  lapply(1:5, function(i){
    cxy=WH*c(i*.15,.5)
    g( 
      text( "S", cxy=cxy, fill=colors[i], font.size=100, 
          font.family="serif", stroke.width=3),
      circle(cxy=cxy, r=39, stroke.width=10, fill='none'  ),
        stroke='black', filter=sbevelFilter(lighting.color=lighting.color[i])
    )
  })
)
```
Looking at the above, we see that the position of reflection varies from S to S. This
is because the position of the S varies, while the lights 
is fixed.  (There are 5 colored lights: "lightblue", "yellow", "pink", "orange", "lightgreen")


### Metallic Shiny S's Lighting Compare

To compare each S, we probably want to require the same light relative light posistion be applied
to each letter. 
Here we maintain the same that relative position.
This causes the highlights to appear in the same position across all letters S. Now we can compare,
however, as a complete image, the effect is poor. This is because the human eye assumes a light source
at a fixed posistion and wants the reflections to vary accordingly.
 


```{r, echo=T, results="asis"}
WH=c(800, 250) # window rect
colors<-c("purple", "darkblue", "darkgreen", "brown", "darkorange")
lighting.color<-c("lightblue", "yellow", "pink", "orange", "lightgreen")

svgR( wh=WH,
  lapply(1:5, function(i){
    cxy=WH*c(i*.15,.5)
    g( 
      text( "S", cxy=cxy, fill=colors[i], font.size=100, 
          font.family="serif", stroke.width=3),
      circle(cxy=cxy, r=39, stroke.width=10, fill='none'  ),
        stroke='black', filter=sbevelFilter(lighting.color=lighting.color[i])
    )
  })
)
```



## A Logo for a User Guide

### A Metallic svgR

Of course we do use any text, with any size for the same filter.

shiny bevel
```{r, echo=T, results="asis"}
WH=c(800, 250) # window rect
svgR( wh=WH,
  g(
    text( "svgR", cxy=WH/2, fill="darkred", font.size=100, 
          font.family="san serif", stroke.width=3),
    circle(cxy=WH/2, r=99, stroke.width=10, fill='none'  ),
     stroke='black', filter=sbevelFilter()
  )
)
```

### Enhancing

Adding a little more text, and setting the lighting.color to "yellow" produces
something could be used as a logo on title page of a document:)

```{r, echo=T, results="asis"}
WH=c(800, 250) # window rect
svgR( wh=WH,     
  g(
    text( "svgR", cxy=WH*c(.25,.5), fill="darkblue", font.size=100, 
          font.family="san serif", stroke.width=3),
    text( "Users Guide", cxy=WH*c(.7,.25), fill="darkblue", font.size=60, 
          font.family="fantasy", stroke.width=3),
    circle(cxy=WH*c(.25,.5), r=99, stroke.width=10, fill='none'  ),
     stroke='black',
    filter = sbevelFilter(lighting.color="yellow")
  )
)
```
