---
title: "ChangingShapes"
output:
  html_document:
    toc: true
    theme: united
---


```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
if(!exists("newId")){
  source('Util.R')
}
```


## Changing Shapes with feMorphology

feMorphology allows thickening and thinning of graphics.
To thicken we use the dilate attribute, to thin we use the erode attribute. The radius specifies the amount, which can be different for x and y, ie. dilate=c(3,1) will dilate in the x direction, but not in the y. Also, dilate=3 is equivalent to dilate=c(3,3) 

```{r, echo=T, results="asis"}
WH=c(800, 220) # window rect
svgR( wh=WH,
  graphPaper(WH, dxy=c(40,40), labels=TRUE), 
  text( 'svgR without feMorphology', xy=c(20,80), font.size=60, stroke="red", fill="none"), 
  text( 'svgR feMorphology dilate=3', xy=c(20,160), font.size=60, stroke="red", fill="none", 
        filter=filter(
          feMorphology(radius=3, operator='dilate')
        )
  )
)
```

feMorphology is does not simply change the stroke-width, rather it smears the source graphics. This is exibited by 
changing the fill from 'none' to 'blue'.

```{r, echo=T, results="asis"}
WH=c(800, 220) # window rect
svgR( wh=WH,
  graphPaper(WH, dxy=c(40,40), labels=TRUE), 
  text( 'svgR without feMorphology', xy=c(20,80), font.size=60, stroke="red", fill="blue"), 
  text( 'svgR feMorphology dilate=3', xy=c(20,160), font.size=60, stroke="red", fill="blue", 
        filter=filter(
          feMorphology(radius=2, operator='dilate')
        )
  )
)
```
Erosion is the opposite of dilation:

```{r, echo=T, results="asis"}
WH=c(800, 220) # window rect
svgR( wh=WH,
  graphPaper(WH, dxy=c(40,40), labels=TRUE), 
  text( 'svgR without feMorphology', xy=c(20,80), font.size=60, stroke="red", fill="blue"), 
  text( 'svgR feMorphology erode=3', xy=c(20,160), font.size=60, stroke="red", fill="blue", 
        filter=filter(
          feMorphology(radius=2, operator='erode')
        )
  )
)
```

