---
title: "feComponentTransfer"
output:
  html_document:
    toc: true
    theme: united
---


```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
if(!exists("newId")){
  source('Util.R')
}
```

## Textures, Marbleing effects 

### Generating Turbulance

Interesting effects can be generate using feTurbulence. A simple example is:


A simple example is:

```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
      use(
        filter=filter(
          feTurbulence(baseFrequency="0.1", numOctaves="2" ) 
        )
      )
)
```

The effect of base frequence is seen here

```{r, echo=T, results="asis"}
WH=c(800, 880) # window rect
N<-99
bfreq<-1:N * .003
svgR( wh=WH,
  text('Base frequencey Effects', xy=c(WH[1]/2, 20)),
  lapply( 1:N, function(i){
    x<-((i-1)%%11)*70
    y<-(floor((i-1)/11))*90+50
    g(
      use( filter=filter( xy=c(x,y), wh=c(60,60),
          feTurbulence(baseFrequency=bfreq[i], numOctaves=2) 
        ) ),
      text(bfreq[i], xy=c(x,y-10))

    )
  })
)
```


```{r, echo=T, results="asis"}
WH=c(800, 880) # window rect
N<-99
bfreq<-1:N * .003
svgR( wh=WH,
  text('Base frequencey Effects', xy=c(WH[1]/2, 20)),
  lapply( 1:N, function(i){
    x<-((i-1)%%11)*70
    y<-(floor((i-1)/11))*90+50
    g(
      use( filter=filter( xy=c(x,y), wh=c(60,60),
          feTurbulence(baseFrequency=bfreq[i], numOctaves=2) 
        ) ),
      text(bfreq[i], xy=c(x,y-10))

    )
  })
)
```

The base frequence can also be expressed along the x or y axis independently by seting baseFrequency=c(xFreq, yFreq)

```{r, echo=T, results="asis"}
WH=c(800, 800) # window rect
baseFeqRng=1:6
numOctRng=1:6
indx<-expand.grid(baseFeqRng, numOctRng)
xp<-function(i)(i-1)*130+50
yp<-function(j)(j-1)*100+50
svgR( wh=WH,
    text("x-baseFrequency", xy=c(WH[1]/2,10)),
    text("y-baseFrequency", xy=c(10,WH[2]/2), writing.mode='tb'),
    lapply(baseFeqRng, function(i){text(0.005*(i-1), xy=c(xp(i),35))}),
    lapply(numOctRng, function(j){text(0.005*(j-1), xy=c(20,yp(j)+50))}),
    apply(indx, 1, function(z){
      i=z[1];j=z[2]
      use(filter=filter( xy=c(xp(i),yp(j)),  wh=c(120,90),
            feTurbulence(baseFrequency=0.005*c(i-1,j-1), numOctaves=1) 
          )
        ) 
      })
)
```


To see the effect of baseFrequency vs numOctaves, we iterate


```{r, echo=T, results="asis"}
WH=c(800, 1200) # window rect
baseFeqRng=1:5
numOctRng=1:10
indx<-expand.grid(baseFeqRng, numOctRng)
xp<-function(i)(i-1)*140+50
yp<-function(j)(j-1)*100+50
svgR( wh=WH,
    text("baseFrequency", xy=c(WH[1]/2,10)),
    text("numOctaves", xy=c(10,WH[2]/2), writing.mode='tb'),
    lapply(baseFeqRng, function(i){text(.02*i, xy=c(xp(i),35))}),
    lapply(numOctRng, function(j){text(2*j-1, xy=c(20,yp(j)+50))}),
    apply(indx, 1, function(z){
      i=z[1];j=z[2]
      use(filter=filter( xy=c(xp(i),yp(j)),  wh=c(130,90),
            feTurbulence(baseFrequency=.02*i, numOctaves=2*j-1 )
          )
        ) 
      })
)
```

##### Types of feTurbulence

The type attribute of feTurbulence can have one of two values:

- turbulence (the default)
- fractalNoise

We have seen the turblance type in action, fractalNoise is similar, but the appears to be softer


```{r, echo=T, results="asis"}
WH=c(800, 150) # window rect
svgR( wh=WH,
      use(
        filter=filter( xy=c(50,70),wh=c(150,60),
          feTurbulence(type='turbulence', baseFrequency="0.1", numOctaves="2" ) 
        )
      ),
            use(
        filter=filter( xy=c(250,70),wh=c(150,60),
          feTurbulence(type='fractalNoise', baseFrequency="0.1", numOctaves="2" ) 
        )
      ),
      text("turbulence", xy=c(100,65)),
      text("fractalNoise", xy=c(300,65))

)
```


```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
     symbol(id='svgRaves', viewbox=c(0,0,100,100),
            text('svgR', xy=c(0,30), stroke='red')),
     use(filter( xy=c(0,0),wh=c("100%","100%"),
            feTile( in1="svgRaves")
          ))
)
```


By using feTurbulance, feComponentTransfer and feColorMatrix we can discover some  effects.  For example, define the following filter

```
funkyFilter<-function(id, baseFrequency=.01, numOctaves=3, slope=c(4,4,4), seed=100){
  filter( id=id, 
    feTurbulence(baseFrequency=baseFrequency, numOctaves=numOctaves, seed=100),
    feComponentTransfer(
      feFuncR(type="linear", slope=slope[1], intercept=-1),
      feFuncG(type="linear", slope=slope[2], intercept=-1),
      feFuncB(type="linear", slope=slope[3], intercept=-1),
      feFuncA(type="linear", slope=0, intercept=1)
    ),
    feColorMatrix(type="saturate") 
  )
} 
```

To see this filter function at work:
```{r, echo=T, results='asis'}
library(svgR)
WH=c(800, 1200) # window rect
N<-4
dH<-WH[2]/(N+1)
y<-0:(N-1)*dH
ww<-WH[1]-40
svgR( wh=WH,
  lapply(1:N, function(i){
    id=paste0("funky1-",i/N)
    bf=.02/(i)
    g(
      text(paste0('funky filter base frequency=', bf), xy=c(20, y[i]+35) ),
      rect(xy=c(20,y[i]+60), wh=c(ww, dH-60 ), 
         filter = funkyFilter(id, baseFrequency = bf))
    )      
  })
    
)
```

##
```{r, echo=T, results="asis"}
library(svgR)
WH=c(800, 400) # window rect
svgR( wh=WH, 
      symbol( 
        rect( id="dispy", xy=c(0,0)*WH,  wh=WH, fill='blue',
                filter = funkyFilter("funky2", baseFrequency = .005)
        )
      ),
      g(  lapply(1:50, function(i)text("Data", cxy=runif(2)*.8*WH, 
                          fill=rrgb(),  font.size=sample(10:30,1))),
          filter=filter( xy=c(0,0), filterUnits='userSpaceOnUse',
              feDisplacementMap(
                in1="SourceGraphic",
                in2=feImage( xlink.href='#dispy'),
                scale=50,
                yChannelSelector="R",
                xChannelSelector="B"
              )
          )
      ),
      text("Data corruption", xy=c(20,30), font.size=26)
)
```





<b>r<br>





