title: "ChangingShapes"
output:
  html_document:
    toc: true
    theme: united
---


```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
  source('Util.R')
```


Another approach to shape modification is to perform some
kind of ***warping*** of a image shape. **feDisplacementMap** allows for this displacing an image according to the colors of a second image. 


Here we illustrate using a linear gradient for the second image.
```{r, echo=T, results="asis"}
library(svgR)
WH=c(800, 200) # window rect
svgR( wh=WH, 
      symbol(
        rect( id="spec", xy=c(0,0), wh=WH, 
          fill=linearGradient( 
            xy1=c("40%",0), xy2=c("60%","0%"),
            colors=rep(c("red","green"),5) 
          )
        )
      ),
      use(xlink.href="#spec"),
      text("feDisplacementMap: linear gradient", xy=c(20,20), stroke='white', fill='white', font.size=20),
      line(xy1=c(100,WH[2]/2),xy2=c(700,WH[2]/2), 
           stroke='yellow', stroke.width=20,
           filter=
             filter( xy=c(0,0), filterUnits='userSpaceOnUse',
                feDisplacementMap(
                  in1="SourceGraphic",
                  in2=feImage(xlink.href='#spec'),
                  scale=50,
                  yChannelSelector="R"
                )
          )
      )
)
```

Our second example uses a radial gradient as the second image

```{r, echo=T, results="asis"}
library(svgR)
WH=c(400, 400) # window rect
svgR( wh=WH, 
      symbol(
        rect( id="radDispl", xy=c(0,0), wh=WH, 
          fill=radialGradient( 
            xy1=c(0,0), xy2=c(1,1), 
            colors=rep(c("red","blue"),5) 
          )
        )
      ),
      use(xlink.href="#radDispl"),
       text("feDisplacementMap: radial gradient", xy=c(20,20), stroke='white', fill='white', font.size=20),
      line(xy1=c(0,150),xy2=c(400,150), 
           stroke='yellow', stroke.width=20,
           filter=
             filter( xy=c(0,0), filterUnits='userSpaceOnUse',
                feDisplacementMap(
                  in1="SourceGraphic",
                  in2=feImage(xlink.href='#radDispl'),
                  scale=50,
                  xChannelSelector="R",
                  yChannelSelector="R"
                )
          )
        ),
        line(xy1=c(0,260),xy2=c(800,260), 
         stroke='yellow', stroke.width=20,
         filter=
           filter( xy=c(0,0), filterUnits='userSpaceOnUse',
              feDisplacementMap(
                in1="SourceGraphic",
                in2=feImage(xlink.href='#radDispl'),
                scale=50,
                xChannelSelector="B",
                yChannelSelector="B"
              )
            )
        )
)
```
<br><br>






Next we try a different approach: use the text as the displacement template and the image is a simple radial gradient
```{r, echo=T, results="asis"}
library(svgR)
WH=c(800, 600) # window rect
svgR( wh=WH, 
      symbol(
        g( id="radDisp2",  
           text("svgR", fill='red', cxy=c(.65,.45)*WH, font.size=200),
            text("in Action", fill='red', cxy=c(.65,.75)*WH, font.size=100)
        )
      ),
      rect(cxy=WH/2, wh=WH, 
           fill=radialGradient( colors=c("white","black") ),
           filter=filter(
                feDisplacementMap(
                  in1="SourceGraphic",
                  in2=feImage(xlink.href='#radDisp2'),
                  scale=50,
                  xChannelSelector="B",
                  yChannelSelector="R"
                )
           )
      )
)
```


Next we try a different approach: use the text as the displacement template and the image is a simple radial gradient
```{r, echo=T, results="asis"}
library(svgR)
WH=c(800, 600) # window rect
svgR( wh=WH, 
      symbol(
        g( id="radDisp2",  
           text("svgR", fill='red', cxy=c(.65,.45)*WH, font.size=200),
            text("in Action", fill='red', cxy=c(.65,.75)*WH, font.size=100)
        )
      ),
      rect(cxy=WH/2, wh=WH, 
           fill=radialGradient( colors=c("grey","white","grey","black") ),
           filter=filter(
                feDisplacementMap(
                  in1="SourceGraphic",
                  in2=feImage(xlink.href="./IMG_4703.JPG"),
                  scale=50,
                  xChannelSelector="B",
                  yChannelSelector="R"
                )
           )
      )
)
```

Finally we apply a displayment to an image.
```{r, echo=T, results="asis"}
library(svgR)
WH=c(800, 600) # window rect
svgR( wh=WH, 
      symbol(
        g( id="radDisp3",  
           text("Seville", fill='red', cxy=c(.8,.5)*WH, font.size=100)
        )
      ),
      image(cxy=WH/2, wh=WH, xlink.href="./IMG_4703.JPG",
           filter=filter(
                feDisplacementMap(
                  in1="SourceGraphic",
                  in2=feImage(xlink.href='#radDisp3'),
                  scale=150,
                  yChannelSelector="R"
                )
           )
      )
)
```



Finally, we use a filter as the second image.
```{r, echo=T, results="asis"}
library(svgR)
WH=c(800, 600) # window rect
svgR( wh=WH, 
      symbol(
        rect( id="symFunkyRect1", xy=c(0,0), wh=WH, 
          filter=funkyFilter('funky4Dispacement1')
        )
      ),
      #use(xlink.href="#symFunkyRect1"),
      text("feDisplacementMap: linear gradient", xy=c(20,20), stroke='white', fill='white', font.size=20),
#      text('Data Denier 4 Pres'
rect( cxy=WH/2, wh=c(600,20), fill='lightblue', 
           stroke='blue', stroke.width=2,
           filter=
             filter( xy=c(0,0), filterUnits='userSpaceOnUse',
                feDisplacementMap(
                  in1="SourceGraphic",
                  in2=feImage(xlink.href='#symFunkyRect1'),
                  scale=50,
                  yChannelSelector="R",
                  yChannelSelector="B"
                )
          )
      )
)
```

## Data Corruption

```{r, echo=T, results="asis"}
library(svgR)
WH=c(800, 400) # window rect
svgR( wh=WH, 
      symbol( 
        rect( id="dispy", xy=c(0,0)*WH,  wh=WH, fill='blue',
                filter = funkyFilter("funky2", baseFrequency = .005)
        )
      ),
      g(  lapply(1:50, function(i)text("Data", cxy=runif(2)*.8*WH, 
                          fill=rrgb(),  font.size=sample(10:30,1))),
          filter=filter( xy=c(0,0), filterUnits='userSpaceOnUse',
              feDisplacementMap(
                in1="SourceGraphic",
                in2=feImage( xlink.href='#dispy'),
                scale=50,
                yChannelSelector="R",
                xChannelSelector="B"
              )
          )
      ),
      text("Data corruption", xy=c(20,30), font.size=26)
)
```
