---
title: "filter-GaussianBlur"
output:
  html_document:
    toc: true
    theme: united
---



```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```


```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
  text(xy=c(20,15), "On the left, the original shape, on the right, the shape with a gaussian blur applied"),
  text( 'svgR', xy=c(50,80), font.size=50, fill="blue"),
  text( 'svgR', xy=c(250,80), font.size=50, fill="blue",
    filter =filter( 
      y=-5, height=40, 
      feGaussianBlur(in1="SourceGraphic", stdDeviation=3)
    )
  ),
  text( 'svgR', xy=c(450,80), font.size=50, fill="blue",
    filter=filter( 
      y=-5, height=40, 
      feGaussianBlur(in1="SourceAlpha", stdDeviation=3)
    )
  ),
  text( 'svgR', xy=c(650,80), font.size=50, fill="blue",
    filter=filter( 
      y=-5, height=40, 
      feGaussianBlur(in1="SourceGraphic", stdDeviation=c(8,0))
    )
  )
)
```


Alternatively, we can place the filter inside a defs element. 

```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
  defs( 
    filter(
      id="blurFilter1", 
      y=-5, height=40,
      feGaussianBlur(in1="SourceGraphic", stdDeviation=3)
    ),          
    filter(
      id="blurFilter2", 
      y=-5, height=40,
      feGaussianBlur(in1="SourceAlpha", stdDeviation=3)
    ),           
    filter(
      id="blurFilter3", 
      y=-5, height=40,
      feGaussianBlur(in1="SourceGraphic", stdDeviation=c(8,0))
    )   
  ),
  text( 'svgR', xy=c(50,80),  font.size=50, fill="blue"),
  text( 'svgR', xy=c(250,80), font.size=50, fill="blue", filter="url(#blurFilter1)"),
  text( 'svgR', xy=c(450,80), font.size=50, fill="blue", filter="url(#blurFilter2)"),
  text( 'svgR', xy=c(650,80), font.size=50, fill="blue", filter="url(#blurFilter3)")
)
```

**Note** The last filter effect gives the appearence of motion and is achieved 
using a stdDeviation in the x direction of 8, and a stdDeviation in the y direction of 0.


### Gaussian Blur and Opacity
It is important to note that blur effect is accomplished by
adjusting the opacity, as see below. This property is extremely useful for
creating an alpha channel to be used in lighting.

```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
  rect(xy=c(0,0), wh=WH, 
    fill=pattern( xy=c(0,0), wh=c(20,20), patternUnits="userSpaceOnUse",
      text("R", cxy=c(10,10), font.size=18, stroke='red')
    )
  ), 
  text( 'DATA', cxy=WH/2, font.size=90, fill="blue",
    filter=filter( y=-5, height=40, feGaussianBlur(in1="SourceGraphic", stdDeviation=5))
  )
)
```

