---
title: "Marker"
output:
  html_document:
    toc: true
    theme: united
---



```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
if(!exists("newId")){
  source('Util.R')
}
```

```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 100) # window rect
svgR( wh=WH,
  path(d="M100,20 l50,0 l0,50 l50,0", fill="none",
    stroke='black', stroke.width=2, 
    marker.end=marker(viewBox=c(0, 0, 10, 10), refXY=c(4,4), 
      markerWidth=7, markerHeight=7, orient="auto",
      rect(xy=c(1,1),wh=c(5,5), fill='black')
    ) 
  )
)
```

Testing:
```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 100) # window rect
svgR( wh=WH,
    defs(
    marker(
      id="marker1",  viewBox=c(0, 0, 10, 10), refXY=c(1,5), 
      markerWidth=6, markerHeight=6, orient="auto",
      path( d=c("M", 0, 0, "L", 10, 5, "L", 0, 10, "z") ) 
    )    
    ),
    line(xy1=c(0,0), xy2=c(300,50), 
         stroke='black', stroke.width=2, 
         marker.end='url(#marker1)'
    )
)
```

With a defs marker works seamlessly and
```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 100) # window rect
svgR( wh=WH,
    defs(
    marker(
      id="marker1",  viewBox=c(0, 0, 10, 10), refXY=c(1,5), 
      markerWidth=6, markerHeight=6, orient="auto",
      path( d="M 0 0 L 10 5 L 0 10 z" ) 
    )    
    ),
    line(xy1=c(0,0), xy2=c(300,50), 
         stroke='black', stroke.width=2, 
         marker.end='url(#marker1)'
    )
)
```

and with the alternate (lazy-inline) approach

### marker end
```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 100) # window rect
svgR( wh=WH,
  line(xy1=c(20,50), xy2=c(300,50), 
    stroke='black', stroke.width=2, 
    marker.end=marker(viewBox=c(0, 0, 10, 10), refXY=c(1,5), 
      markerWidth=6, markerHeight=6, orient="auto",
      path( d=c("M", 0, 0, "L", 10, 5, "L", 0, 10, "z") )
    ) 
  )
)
```

### marker start

```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 100) # window rect
svgR( wh=WH,
  line(xy1=c(20,50), xy2=c(300,50), 
    stroke='black', stroke.width=2, 
    marker.start=marker(viewBox=c(0, 0, 10, 10), refXY=c(1,5), 
      markerWidth=6, markerHeight=6, orient="auto",
      path( d="M 0 0 L 10 5 L 0 10 z" )
    ) 
  )
)
```
### Marker mid

We can also place markers in middle points that define a path.

```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 100) # window rect
svgR( wh=WH,
  path(d="M100,20 l50,0 l0,50 l50,0", fill="none",
    stroke='black', stroke.width=2, 
    marker.mid=marker(viewBox=c(0, 0, 10, 10), refXY=c(4,4), 
      markerWidth=7, markerHeight=7, orient="auto",
      rect(xy=c(1,1),wh=c(5,5), fill='black')
    ) 
  )
)
```


also works when removing the defs.
```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 100) # window rect
svgR( wh=WH,
    #defs(
    marker(
      id="marker1",  viewBox=c(0, 0, 10, 10), refXY=c(1,5), 
      markerWidth=6, markerHeight=6, orient="auto",
      path( d="M 0 0 L 10 5 L 0 10 z" ) 
    #)    
    ),
    line(xy1=c(0,0), xy2=c(300,50), 
         stroke='black', stroke.width=2, 
         marker.end='url(#marker1)'
    )
)
```
