---
title: "Marker"
output:
  html_document:
    toc: true
    theme: united
---



```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

## Markers

Markers are used to decorate the ends paths and any adjoining segements.
Markers allow for a completely customizable looking arrowhead and tails

First we show some simple examples for the end, start, middle



### An Arrow Tail

```{r, echo=T, results="asis"}
WH=c(600, 100) # window rect
svgR( wh=WH,
  line(xy1=c(20,50), xy2=c(300,50), 
    stroke='black', stroke.width=2, 
    marker.start=marker(viewBox=c(0, 0, 10, 10), refXY=c(1,5), 
      markerWidth=6, markerHeight=6, orient="auto",
      path( d="M 0 0 L 10 5 L 0 10 z" )
    ) 
  )
)
```
### An Arrow Middle

We can also place markers in middle points that define a path.

```{r, echo=T, results="asis"}
WH=c(600, 100) # window rect
svgR( wh=WH,
  path(d="M100,20 l50,0 l0,50 l50,0", fill="none",
    stroke='black', stroke.width=2, 
    marker.mid=marker(viewBox=c(0, 0, 10, 10), refXY=c(4,4), 
      markerWidth=7, markerHeight=7, orient="auto",
      rect(xy=c(1,1),wh=c(5,5), fill='black')
    ) 
  )
)
```

### An Arrow Head

```{r, echo=T, results="asis"}
WH=c(600, 100) # window rect
svgR( wh=WH,
  line(xy1=c(20,50), xy2=c(300,50), 
    stroke='black', stroke.width=2, 
    marker.end=marker(viewBox=c(0, 0, 10, 10), refXY=c(1,5), 
      markerWidth=6, markerHeight=6, orient="auto",
      path( d=c("M", 0, 0, "L", 10, 5, "L", 0, 10, "z") )
    ) 
  )
)
```

Putting it together

### Tail, Middle, Head 


We can also place markers in middle points that define a path.

```{r, echo=T, results="asis"}
WH=c(600, 100) # window rect
svgR( wh=WH,
  path(d="M100,20 l50,0 l0,50 l50,0", fill="none",
    stroke='black', stroke.width=2, 
    marker.start=marker(viewBox=c(0, 0, 10, 10), refXY=c(1,5), 
      markerWidth=6, markerHeight=6, orient="auto",
      path( d="M 0 0 L 10 5 L 0 10 z" )
    ),
    marker.mid=marker(viewBox=c(0, 0, 10, 10), refXY=c(4,4), 
      markerWidth=7, markerHeight=7, orient="auto",
      rect(xy=c(1,1),wh=c(5,5), fill='black')
    ),
    marker.end=marker(viewBox=c(0, 0, 10, 10), refXY=c(1,5), 
      markerWidth=6, markerHeight=6, orient="auto",
      path( d=c("M", 0, 0, "L", 10, 5, "L", 0, 10, "z") )
    ) 
  )
)
```

### Exotic Markers

Varing the marker path allows for a more exotic look.

```{r, echo=T, results="asis"}
WH=c(600, 100) # window rect
svgR( wh=WH,
  path(d="M100,20 l50,0 l0,50 l50,0", fill="none",
    stroke='black', stroke.width=2, 
    marker.start=marker(viewBox=c(0, 0, 10, 10), refXY=c(1,5), 
      markerWidth=16, markerHeight=16, orient="auto",
      path( d=c( "M", c(0,0), "A", c(5,5), 0, 1,  1,  c(0,10),  "Z" ) )
    ),
    marker.mid=marker(viewBox=c(0, 0, 10, 10), refXY=c(4,4), 
      markerWidth=7, markerHeight=7, orient="auto",
      circle(cxy=c(5,5),r=5, fill='red')
    ),
    marker.end=marker(viewBox=c(0, 0, 20, 20), refXY=c(1,10), 
      markerWidth=16, markerHeight=20, orient="auto",
      path( fill="blue",
        d=c("M", c(0,0), 
            "A", c(20,20), 0, 0,  0,  c(10,10),
            "A", c(20,20), 0, 0,  0,  c(0,20),
            "Z") 
      )
    ) 
  )
)
```


### Multiple Arrows

When employing a marker multiple times, it may be more
efficient to include it in a defs container and then 
provide a reference to it.


```{r, echo=T, results="asis"}
WH=c(600, 100) # window rect
svgR( wh=WH,
  defs(
    marker(id="marker.start", viewBox=c(0, 0, 10, 10), refXY=c(1,5), 
      markerWidth=16, markerHeight=16, orient="auto",
      path( d=c( "M", c(0,0), "A", c(5,5), 0, 1,  1,  c(0,10),  "Z" ) )
    ),
    marker(id="marker.mid", viewBox=c(0, 0, 10, 10), refXY=c(4,4), 
      markerWidth=7, markerHeight=7, orient="auto",
      circle(cxy=c(5,5),r=5, fill='red')
    ),
    marker(id="marker.end", viewBox=c(0, 0, 20, 20), refXY=c(1,10), 
      markerWidth=16, markerHeight=20, orient="auto",
      path( fill="blue",
        d=c("M", c(0,0), 
            "A", c(20,20), 0, 0,  0,  c(10,10),
            "A", c(20,20), 0, 0,  0,  c(0,20),
            "Z") 
      )
    ) 
  ),
  path(d="M100,20 l50,0 l0,50 l50,0", fill="none",
    stroke='black', stroke.width=2, 
    marker.start='url(#marker.start)',
    marker.end='url(#marker.end)',
    marker.mid='url(#marker.mid)'
  ),
  path(d = c("M",200, 20, "C", 250, 80, 380, 0, 400, 50, "L", 300, 80),  
    stroke='green', stroke.width=2, fill="none",
    marker.start='url(#marker.start)',
    marker.end='url(#marker.end)',
    marker.mid='url(#marker.mid)'
  )
)
```




