---
title: "Mask"
output:
  html_document:
    toc: true
    theme: united
---


```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
  source('Util.R')
```

Using explicit defs

```{r, echo=T, results="asis"}

WH=c(600, 100) # window rect
svgR( wh=WH,
  defs(
    mask(
      id="mask0",
      rect(cxy= WH/2, wh=c(600,40), fill='white'),
      rect(cxy= WH/2, wh=c(600,10), fill='black')
    )
  ),
  g(
    circle( cxy=c(50,50), r=30, fill='darkblue'),
    circle( cxy=c(150,50), r=30, fill='red'),
    mask="url(#mask0)"
  )
)
```

Using implied defs
```{r, echo=T, results="asis"}
WH=c(600, 100) # window rect
svgR( wh=WH,
    g( 
      circle( cxy=c(50,50), r=30, fill='darkblue'),
      circle( cxy=c(150,50), r=30, fill='red'),
      mask=mask(
        rect(cxy= WH/2, wh=c(600,40), fill='white'),
        rect(cxy= WH/2, wh=c(600,10), fill='black')
        )
    )
)
```


```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 100) # window rect
svgR( wh=WH,
    defs(
        mask(
        id="mask1", 
        xy=c(0,0), wh=c(100,100),
        text(xy=c(0,50),'hello all you brilliant data scientists', 
             stroke='white',fill='black')
        )
    ),
    #rect(xy=c(0,0), wh=WH, fill='yellow'),
    circle( cxy=c(50,50), r=30, fill='darkblue', mask="url(#mask1)"),
    circle( cxy=c(150,50), r=30, fill='red',     mask="url(#mask1)")
)
```

```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 100) # window rect
svgR( wh=WH,
    defs(
        mask(
        id="mask2", xy=c(0,0), wh=c(100,100),
        rect( xy=c(0,0),  wh=c(50,100), fill='white', stroke='none'),
        rect( xy=c(50,0), wh=c(50,100), fill='grey', stroke='none')
        )
    ),
    #rect(xy=c(0,0), wh=WH, fill='yellow'),
    text(xy=c(0,50),'hello all you brilliant data scientists', stroke='red',fill='white'),
    circle( cxy=c(50,50), r=39, fill='darkblue', mask="url(#mask2)"
             )
)
```


```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 100) # window rect
svgR( wh=WH,   
    mask(
    id="mask2-5", xy=c(0,0), wh=c(100,100),
    rect( xy=c(0,0),  wh=c(50,100), fill='white', stroke='none'),
    rect( xy=c(50,0), wh=c(50,100), fill='grey', stroke='none')
    ),
    #rect(xy=c(0,0), wh=WH, fill='yellow'),
    text(xy=c(0,50),'hello all you brilliant data scientists', stroke='red',fill='white'),
    circle( cxy=c(50,50), r=39, fill='darkblue', mask="url(#mask2-5)"
             )
)
```


```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 100) # window rect
svgR( wh=WH,
    defs(
        mask(
        id="mask3", xy=c(0,0), wh=c(100,100),
        rect( xy=c(0,0),  wh=c(50,100), fill='rgb(190,190,130)', stroke='none'),
        rect( xy=c(50,0), wh=c(50,100), fill='grey', stroke='none')
        )
    ),
    #rect(xy=c(0,0), wh=WH, fill='yellow'),
    text(xy=c(0,50),'hello all you brilliant data scientists', stroke='red',fill='white'),
    circle( cxy=c(50,50), r=39, fill='darkblue', mask="url(#mask3)"
             )
)
```


```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 100) # window rect
svgR( wh=WH,
  mask(
    id="mask4", xy=c(0,0), wh=c(100,100),
    rect( xy=c(0,0),  wh=c(50,100), fill='rgb(190,190,130)', stroke='none'),
    rect( xy=c(50,0), wh=c(50,100), fill='grey', stroke='none')
  ),
  #rect(xy=c(0,0), wh=WH, fill='yellow'),
  text(xy=c(0,50),'hello all you brilliant data scientists', stroke='red',fill='white'),
  circle( cxy=c(50,50), r=39, fill='darkblue', mask="url(#mask4)"
           )
)
```


```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 100) # window rect
svgR( wh=WH,
  pattern(
    id="pattern2", xy=c(5,5), wh=c(10,10), patternUnits="userSpaceOnUse" ,
    circle(cxy=c(5,5), r=5, fill='white')
  ),  
  mask( id="mask5", xy=c(0,0), wh=WH,
    rect(xy=c(0,0), wh=WH, fill="black"),
    rect(xy=c(0,0), wh=WH, fill="url(#pattern2)")
  ),
  text(xy=c(0,50),'hello all you brilliant data scientists', stroke='black',font.size=30,
        mask="url(#mask5)"
       )
)
```


```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 100) # window rect
svgR( wh=WH,
  mask( id="mask5", xy=c(0,0), wh=WH,
    rect(xy=c(0,0), wh=WH, fill="black"),
    lapply(seq(1,90,5), function(i){rect(xy=c(0,i), wh=c(600,3), fill='white')})
  ),
  text(xy=c(0,50),'hello all you brilliant data scientists', stroke='black',font.size=30,
        mask="url(#mask5)"
       )
)
```

```{r, echo=T, results="asis"}
#```{r, echo=T}
WH=c(600, 200) # window rect
svgR( wh=WH,
  pattern(
      id="pattern3", xy=c(10,10), wh=c(20,20), patternUnits="userSpaceOnUse" ,
      circle(cxy=c(10,10), r=10, fill='green')
  ),
  g(
    rect(xy=c(0,0), wh=WH, fill='orange'),
    ellipse( cxy=c(55,60), rxy=c(55,25), fill="url(#pattern3)")
    )
)
```
