---
title: "Brief Intro"
output:
  html_document:
    toc: true
    theme: united
---



```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
  source('Util.R')
```

```{r, echo=FALSE}
library(stringr)
showCode %<c-% function(txt){
  indentCount<-function(s){
    if(grepl(pattern="^ +", s)){
      nchar(str_match(pattern="^ +", s))   
    } else {
      0
    }
  }
  line2span<-function(s, default='black'){
    pattern='\\\\\\w+'
    #browser()
    if(grepl(pattern,s)){
      spns<-str_split(pattern=pattern,s)[[1]]
      cols<-as.character(str_match_all(pattern=pattern,s)[[1]])
      cols<-gsub("\\\\","",cols)
      if(length(cols)<length(spns)){
        cols<-c('black',cols)
      }
    } else {
      spns<-s
      cols<-'black'
    }
    lapply(1:length(spns), function(i){
      font.size=ifelse( cols[i]=="black",12,16 )
      stroke.width=ifelse( cols[i]=="black",.2,1 )
      tspan(spns[i], stroke=cols[i], fill=cols[i], font.size=font.size, stroke.width=stroke.width)
    })
  }
  
  lines<-str_split(pattern="\\n",txt1)[[1]]  
  lapply(1:length(lines), function(i){
    line<-lines[i]
    ic<-indentCount(line)
    xy<-c(ic*10+20, i*18+10)
    text(xy=xy, line2span(line))
  })
}

WH<-c(910,100)

codeBlock  %<c-% function(wh){
  rect(xy=c(1,1), wh=wh-c(2,2), fill="#F4F4F4", stroke='#444', rxy=c(5,5), stroke.width=.3)
}
```



The svgR package provides a DSL to create SVG markup from R. It is similar to
grid graphics, in the sense that it is a low level api upon which higher level graphics
constructs can be built.

The best way to learn is by example, so here is a short example.

## A Short Example

We start by first considering the short example show below together
with the rendered result.

```{r, echo=TRUE, results="asis"}
svgR( wh=c(100,50),
  text( xy=c(20,10), "line 1", 
    tspan(dy=10, "line2 ")
  )
)
```

What is hidden in the above code listing is 
the
usual 3 backtics and the options *echo=FALSE* and *results="asis"* 
For completeness, we show these below in red.

```{r, echo=FALSE, results="asis"}
WH<-WH+c(0,40)
txt1<-'\\red \\`\\`\\`{r, echo=FALSE, results="asis"}
svgR( wh=c(100,50),
  text( xy=c(20,10), "line 1", 
    tspan(dy=10, "line2 ")
  )
)
\\red \\`\\`\\`
'
svgR( wh=WH, font.family='sans-serif', font.size=12, stroke.width=0.2, letter.spacing=2,
  codeBlock(WH),
  showCode(txt1)
)
```

Next we break apart the code into it's basic components, starting with the function call **svgR**

```{r, echo=FALSE, results="asis"}
txt1<-
'\\purple svgR( \\black wh=c(100,50),
  text( xy=c(20,10), "line 1", 
    tspan(dy=10, "line2 ")
  )
)
'
svgR( wh=WH, font.family='sans-serif', font.size=12, stroke.width=0.2, letter.spacing=2,
  codeBlock(WH),
  showCode(txt1)
)
```



It has two arguments, a *named parameter wh=c(100,50)* and  an *unnamed parameter*
which is the function called **text**. 

The named parameter *wh* (shown in orange below) is an attribute specifying the *width* and *height* of the viewport,
the region on the screen where the image is shown.

```{r, echo=FALSE, results="asis"}
txt1<-
'svgR( \\orange wh=c(100,50),
   text( \\black xy=c(20,10), "line 1", 
    tspan(dy=10, "line2 ")
  )
 )
'
svgR( wh=WH, font.family='sans-serif', font.size=12, stroke.width=0.2, letter.spacing=2,
  codeBlock(WH),
  showCode(txt1)
)
```

The **text** function (shown in blue) is one of 80 functions that can be called within the scope of 
svgR to render an SVG markup. These functions are called *svgR elements* (or simply elements)
and each svgR *element* creates and SVG node whose tag name is the same as the function name.

```{r, echo=FALSE, results="asis"}
txt1<-
'svgR(  wh=c(100,50),
  \\blue text( \\black xy=c(20,10), "line 1", 
    tspan(dy=10, "line2 ")
  )
 )
'
svgR( wh=WH, font.family='sans-serif', font.size=12, stroke.width=0.2, letter.spacing=2,
  codeBlock(WH),
  showCode(txt1)
)
```

This text element has 3 components: an xy *attribute* (in orange), text content (in green), 
and a child element named **tspan** (in blue).

```{r, echo=FALSE, results="asis"}
txt1<-
'svgR(  wh=c(100,50),
  text( \\orange xy=c(20,10), \\green "line 1", 
    \\blue tspan( \\black dy=10, "line2 " \\blue ) 
  \\black )
 )
'
svgR( wh=WH, font.family='sans-serif', font.size=12, stroke.width=0.2, letter.spacing=2,
  codeBlock(WH),
  showCode(txt1)
)
```

The **tspan** element contains an *dy attribute* (orange) and *text content* (green)

```{r, echo=FALSE, results="asis"}
txt1<-
'svgR(  wh=c(100,50),
  text( xy=c(20,10), "line 1", 
    tspan( \\orange dy=10,  \\green "line2 " \\black )
  )
 )
'
svgR( wh=WH, font.family='sans-serif', font.size=12, stroke.width=0.2, letter.spacing=2,
  codeBlock(WH),
  showCode(txt1)
)
```

The  generated **SVG** markup is 

```
<svg xmlns="http://www.w3.org/2000/svg" xmlns:xlink="http://www.w3.org/1999/xlink"
  width="100" height="50">
  <text x="20" y="10">
    line 1
    <tspan dy="10">tspan line 2</tspan>
  </text>
</svg>
```

**Note**: The  SVG root node was created by the svgR call.


## Taxonomy

In the previous section we walked though an example where we
encountered  

- elements (text, tspan), 
- attributes (xy, wh, dy ) and 
- text content (line1, line2).

However, this is only small fragment of the available elements and attributes.  

The elements (functions within an svgR call) can be broadly broken into 
several categories:

```{r, echo=FALSE, results="asis"}
nodeMapping<-data.frame(
parent=c( 1,1,1,1,1, 2, 2, 5, 5, 5, 5),
child= c( 2,3,4,5,6, 7, 8,9, 10, 11, 12)
)
WH<-c(800,500)
nodeLabels<-c('elements', 'graphical', 'containers', 'animators', 'modifiers', 'other', 
'shape related', 'text related', 'filters', 'gradients', 'masks', 'markers')
svgR(wh=WH,
  svgRTree(nodeMapping, nodeLabels)
)
```

### Graphical Elements
The Graphical Elements, *Text* and *Shapes*, are the building blocks of an **SVG** image. 
Hence it is fundamental to understand their usage. This means that one needs to understand
the attributes that can be applied to *Text* and *Shapes*. 
Some of the relevant prominent attributes are 

attribute | description
-----------|----------------
xy,wh,cxy, ...| intial positioning
transform | for repositioning 
fill  | to specify how to fill the interior
filter | modify the image
mask | to hide a portion
clip.path | to trim a portion


### Containers

To apply the same value of an attribute to 
a collection of shapes and text, it is customary to group them together using a
container element such as 
either a **g** element or the **svg** element. 
The distinction between  **g** and **svg**  lies in which attributes the can take. 

- **svg** can has *coordinate attributes xy, wh* but not a *transform attribute*
- **g** has the *transform attribute*, but no coordinates

Other containers elements are **defs**, **symbol**.  
However, these are not displayed directly, but are used by reference. 

### Animators

Animation can be either instaneous or transitional, i.e. occuring over period of time.

The Prominent Animation Elements are

element | description
-----------|----------------
Set | Instaneous change
Animate | Continous change
Animate Transform | Continous Change
Animate Motion | Continous Chagne

Animators are often given as a child of the element to be animated, i.e., 
an unnamed argument of the shape/text element.

### Modifiers: Gradients, Filters, Masks, Markers

These Modifiers modify the appearence graphical element. More specifically,

- Gradients: provide a continous flow of the fill color from one spectrum to another
- Masks: hides the portion of an image given by the mask
- Markers: decorates the endpoints of a line, allow us to create arrow heads, tails, etc.
- Filters: are containers for filter primitives, used to provide, lighting, dropshawdow, 
and mamy special effects.

These elements are specified by the value of a named parameter of  the 
element they are to influence. For example:

element | parameter usage
-----------|----------------
Gradients |  fill=linearGradient(...) or fill=radialGradient(...)
Filter | filter=filter(...)
Mask | mask=mask(...)
marker | marker.end=marker(...), marker.start=marker(...), marker.mid=marker(...)


## Structure of This Document

As stated before, the graphical elements are the building blocks of any **SVG** image. 
However, in order to use those building blocks, we need to know where to put them
and how to glue them together.  

The sections 

section | description
---------|-----------------------------
Coordinates, Viewports and Containers | placement of shapes
Transform Attribute | placement of shapes
Shapes,  Paths, Markers | Basic Shapes
Text | Text related
The Fill Attribute | Gradients, Patterns
The Filter Attribute | Gaussian Blur, Filter Coords
Color Filters | Filters that influence color
Combining Filter Primitives| feComposite, feBlend, 
Texture Filter | Textures and Marbling
Lighting Filters | shiny bevel
Shape Shifters | clip.path
Animating | Set, Animation, Animation Transform, Animate Motion

Each section contains multiple short examples of each topic.
Enjoy!

