---
title: "Path Element"
output:
  html_document:
    toc: true
    theme: united
---




```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
if(!exists("newId")){
  source('Util.R')
}
```




## The Path Data Attribute

The path element is used to specify a curve. The path may be rendered directly by the setting the stroke attribute or used by other elements such as textPath. 

The curve of a path may consist of any combination of lines, Bézier curves (cubic and quadratic), and ellisoid arcs. The exact combination specified in the **d** atttribue as a sequence of commands and points, summarized by the following table



Command  | Points |  Interpretation
----|---------- | -------------
M,m  | (x y)+ | Move the current point to a new location (without drawing) 
L,l |  (x, y)+ | Draw a line from the current point to x,y
H,h | x+ | Draw a vertical line from the current
V,v | y+ | Draw a horizontal line from the current point
C,c | (x1 y1 x2 y2 x y)+ | Draw a cubic Bézier curve to x,y using x1,y1; x2,y2 as control pts
S,s | (x2 y2 x y)+ | Draw a cubic Bézier to  x,y using  x2,y2 as the second control pt (inheriting the first control point from a preceding  Bézier)
Q,q | (x1 y1 x y)+ | Draw a quadratic Bézier curve  to x,y using  x1,y1 as control pts
T,t | (x y)+ | Draw a quadratic Bézier curve to x,y (inheriting the first control from a preceding  Bézier)
A,a | (rx ry x-axis-rotation large-arc-flag sweep-flag x y)+ | Draw an elliptical arc to x, y
Z   | | closes the path

Capital lettered commands interpret the points as absolute coordinates, lowercase interprets the points as relative point to the current (starting) point.

The following illustrates the basics  of each command.


### Lines

Using absolute coordinates (Note: we graphPaper background to illustrate the coordinates)

```{r, echo=T, results="asis"}
WH<-c(600,200)
svgR( wh=WH, 
      graphPaper(WH, dxy=c(20,20), labels=TRUE),
      path(d=c("M", 180,90, "L", 150,142,90,142,60,90,90,38,150,38) , stroke='red', fill='none') )
```

Using relative coordinates
```{r, echo=T, results="asis"}
WH<-c(600,200)
svgR( wh=WH, 
      graphPaper(WH, dxy=c(20,20), labels=TRUE),
      path(d=c("m", 180,90, "l", -30,52,-60,0,-30,-52,30,-52,60,0 ) , stroke='green', fill='none') )
```

Closing the path.
To close the path add Z at the end
```{r, echo=T, results="asis"}
WH<-c(600,200)
svgR( wh=WH, 
      graphPaper(WH, dxy=c(20,20), labels=TRUE),
      path(d=c("m", 180,90, "l", -30,52,-60,0,-30,-52,30,-52,60,0, "Z" ) , stroke='blue', fill='none') )
```


## The Elliptic Arc Curve

An elliptic Arc is an arc formed from fragment of an ellipse which has possibly been rotated. 

To specify that
arc, we specify the end points of the arc and 
then the ellipse from which the arc is formed, 
and a flag to indicate which portion of the ellipse contains the arc.

To specify a general ellipse, we need 5 numbers, 
for example a rotational angle $\theta$, (called *x-axis-rotation*), 
the major and minor radii *rx, ry* and something equivalent to the center of the ellipse. 
However, we note that rx, ry, rotation together with endpoints of the arc, 
determine the center to be of one of two locations. So we need only one additional 
parameter, called the *large-arc-flag*
 
```{r, echo=T, results="asis"}
theta=20
startPt=c(300,75)
endPt<-startPt+c(0,50)
rxy=c(100,50)
laf0=0
laf1=1
sf0=0
sf1=1
WH<-c(600, 200)
svgR( wh=WH, fill="none",
      graphPaper(WH),
      stroke.width=3,
      path(d=c("M", startPt, "A", rxy, theta, laf0, sf0, endPt ) , stroke='red') ,
      path(d=c("M", startPt, "A", rxy, theta, laf0, sf1, endPt ) , stroke='blue') ,
      path(d=c("M", startPt, "A", rxy, theta, laf1, sf0, endPt ) , stroke='green') ,
      path(d=c("M", startPt, "A", rxy, theta, laf1, sf1, endPt ) , stroke='orange'),
      line( xy1=startPt, xy2=startPt-c(0,40), stroke.width=1, stroke='black'),
      line( xy1=endPt, xy2=endPt+c(0,40), stroke.width=1, stroke='black'),
      text("start pt", cxy=startPt-c(0,50), font.size=12, stroke='black', stroke.width=1),
      text("end pt",   cxy=endPt+c(0,50), font.size=12, stroke='black', stroke.width=1),
      text("sf=0 on this side", xy=c(WH[1]-100,50), font.size=12, stroke='black', stroke.width=1),
      text("sf=1 on this side", xy=c(50,WH[2]-50),  font.size=12, stroke='black', stroke.width=1)

)
```
 
 
## Q: The Quadratic Bézier

A  Quadratic Bézier is a parametric curve whose coodinates are given by quadratic expressions in a parameter t. The most common formulation is given points $P_0$, $P_1$, $C_0$, a point on the curve is given by
$$ P= (1-t)^2 P_0 + 2(1-t)t C_0 +t^2 P_1$$ for $0 \leq t \leq1$




Thus Quadratic Bézier requires three points: $P_0$,  the starting point, $P_1$ the ending point, and $C_0$, a  control point. One can easliy verify that the control point lies on the intersection of the the lines tangent to the end points. 


This is illustrated in the following example

```{r, echo=T, results="asis"}
P0=c(200,50)
P1=c(300,150)
C0<-c(100,100)
WH<-c(600,200)
svgR( wh=WH, 
  graphPaper(wh=WH, dxy=c(20,20), labels=TRUE), 
  path(d=c("M",P0, "Q", C0,  P1) , stroke='red',
  stroke.width=3, fill='none') ,
  g( 
    stroke='blue',
    line(xy1=P0,xy2=C0),
    line(xy2=P1,xy1=C0)
  ),
  g(
    stroke="black",
    text("P0",cxy=P0 ),
    text("C0",cxy=C0),
    text("P1", cxy=P1)
  )
)
```


## T: Extending a  Bézier with a Quadratic Bézier

Using the T command, we can append to a given Bézier a Quadratic Bézier without specifying an additional control point. The control point that is used will be the reflection of the control point of the previous Bézier. That is, $$C_{i+1}= 2 \times P_i + C_i$$ where $C_i$ is the control point of the previous Bézier. Thus,we ensure continuity of the deriverative of the beginning endpoint of the appended Bézier.

We demonstarte this
in the following example

```{r, echo=T, results="asis"}
Pt0=c(200,50)
Pt1=c(300,150)
C0<-c(100,100)
C1<-2*Pt1 -C0
Pt2=c(400,50)
svgR( wh=c(600, 250), 
      graphPaper(c(600,250)),
      path(d=c("M",Pt0, "Q", C0,  Pt1 , "T", Pt2) , stroke='red', stroke.width=3, fill='none') ,
      #path(d=c("T", endPt2) , stroke='green', fill='none') ,
      g( 
        stroke='blue',
        line(xy1=Pt0,xy2=C0),
        line(xy2=Pt1,xy1=C0),
        line(xy1=Pt1,xy2=C1),
        line(xy2=Pt2,xy1=C1)
      ),
      g(
        stroke="black",
        text("P0",cxy=Pt0 ),
        text("C0",cxy=C0),
        text("C1",cxy=C1),
        text("P2",cxy=Pt2),
        text("P1", cxy=Pt1)
      )
    )
```

### C: The Cubic Bézier

The **Cubic Bézier** is similar to the Quadratic Bézier, but has a higher degree allowing for a S-shaped curve. The Cubic Bézier requires four points, $P_0$ the start point, $P_1$ the end point and 
two control points $C_0$, $C_1$ . 
$$ P(t)=(1-t)^3 P_0 + 3 (1-t)^2 t C_0 + 3 (1-t)t^2 C_1 + t^3 P_p $$ 
for $0 \leq t \leq 1$

The first control point,  $C_0$,  together with the starting point, $P_0$,  determines a vector, $C_0-P_0$ which is tangent to the  curve at $P_0$. Similarly, the second control point, $C_1$,  together with ending point, $P_1$, determine a vector, $P_1-C_1$ which is tangent to the curve at $P_1$. The magnititude of the vectors $C_0 - p_0$, and $P_1 -C_1$ determine the intermediate shape.

The current location is considerd to be the start point ($P_0$), so we  usually begin a path specification with an **M** command. This is followed by a **C** command. (or a lower case c if we are using relative coordinates) and the control points,  $C_0$, $C_1$, and the ending point $P_1$.

This is illustrated by the following:

```{r, echo=T, results="asis"}
P0=c(200,50)
P1=c(300,150)
C0<-c(100,100)
C1<-c(400,100)
svgR( wh=c(600, 200), 
      graphPaper(c(600,200)),
      path(d=c("M",P0, "C", C0, C1, P1) , stroke='red', stroke.width=3, fill='none') ,
      g( 
        stroke='blue',
        line(xy1=P0,xy2=C0),
        line(xy2=P1,xy1=C1)
      ),
      g(
        stroke="black",
        text("P0",cxy=P0 ),
        text("C0",cxy=C0),
        text("C1",cxy=C1),
        text("P1", cxy=P1)
      )
    )
```

Interchanging ctrlPt2 with the end point gives 

```{r, echo=T, results="asis"}
P0=c(200,50)
C1=c(300,150)
C0<-c(100,100)
P1<-c(400,100)
svgR( wh=c(600, 200), 
      graphPaper(c(600,200)),
      path(d=c("M",P0, "C", C0, C1, P1) , stroke='red', stroke.width=3, fill='none') ,
      g( 
        stroke='blue',
        line(xy1=P0,xy2=C0),
        line(xy2=P1,xy1=C1)
      ),
      g(
        stroke="black",
        text("P0",cxy=P0 ),
        text("C0",cxy=C0),
        text("C1",cxy=C1),
        text("P1", cxy=P1)
      )
    )
```


## S: Extending a  Bézier with a Cubic Bézier


Just as with the quadratic case, we can extend a Bézier with a Cubic Bézier while specifying one less control point, i.e. we take as the first control point,the last control point of the previous Bézier, thus ensuring continuity of the first derivative.

We illustrate this with the following example. 



```{r, echo=T, results="asis"}
P0<-c(350,30)
P1<-c(400,150)
P2<-c(200,250)
  
C0<-P0+c(-100,50)  
C1<-P1+c(100,-50) 
C2<-P2+c(-80,-100)

R2<-2*P1-C1 #used only to display the reflected control point

svgR( wh=c(600, 300), 
      graphPaper(c(600,300)),
      #The path of interest
      path(d=c("M",P0, "C", C0, C1, P1, "S", C2, P2) , stroke='red', stroke.width=3, fill='none') ,
      
      g( # lines to illustrate control points
        stroke='blue',
        line(xy1=P0,xy2=C0),
        line(xy2=P1,xy1=C1),
        line(xy1=P1,xy2=R2),
        line(xy2=P2,xy1=C2)
      ),
      g( # labeling of control points
        stroke="black",
        text("P0",cxy=P0 ),
        text("C0",cxy=C0),
        text("C1",cxy=C1),
        text("P1", cxy=P1),
        text("C2",cxy=C2),
        text("R2",cxy=R2),
        text("P2", cxy=P2)
      )
    )
```


**Note**, the control points for the $P_1-P_2$ portions
of the path are $R_2$, $C_2$, where $C_2$ is specified in the path, but $R_2$ is implicitly defined
as the reflection of $C_1$ insidet the path data description attribute *d*. That is "S" uses the $P_1$ and $C_1$ defined to it's left to compute $R_2=2 \times P_1 - C_1$.


## More  Bezier Examples

```{r, echo=T, results="asis"}
svgR( wh=c(800, 200), 
      graphPaper(c(800,200)),
      g( stroke.width=3, fill="transparent",
        path(d = c("M",100, 0, "C", 100, 100, 200, 0, 200, 100), stroke='red' ), 
        path(d = c("M",200, 0, "c", 0, 100, 100, 0, 100, 100),  stroke='pink'), 
        path(d = c("M",300, 0, "c", 0, 100, 100, 0, 100, 100),  stroke='purple'), 
        path(d = c("M",400, 0, "c", 0, 100, 100, 0, 100, 200),  stroke='green'), 
        path(d = c("M",500, 0, "c", 0, 100, 100, 0, 100, 200),  stroke='blue'), 
        path(d = c("M",600, 0, "c", 0, 300, 100, -100, 100, 200),  stroke='orange')
      )
 )
```
