---
title: "Set Up"
output:
  html_document:
    toc: true
    theme: united
---

```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
if(!exists("newId")){
  source('Util.R')
}
```

## To install from github

```
install.packages("devtools")
install_github("svgR")
```

## To render in an RPres or R Markdown document 
- Include prior to any calls to svgR a block with
```
'''{r, echo=FALSE}
library(svgR)
'''
```
- To render either 
  + a simple call to svgR inside a block with results='asis'
```
'''{r, echo=T, results="asis"}
svgR(  wh=c(600, 100), circle(cxy=c(200,50),r=50, fill='red') )
'''
```
```{r, echo=T, results="asis"}
svgR(  wh=c(600, 100), circle(cxy=c(200,50),r=50, fill='red') )
```

or

  + call to svgR to create an svg markup object and echo back within a block with results='asis'
```
'''{r, echo=F, results="asis"}
svgM<-svgR(  wh=c(600, 100), circle(cxy=c(200,50),r=50, fill='red') )
svgM
'''
```
```{r, echo=F, results="asis"}
svgM<-svgR(  wh=c(600, 100), circle(cxy=c(200,50),r=50, fill='red') )
svgM
```

**Note** Note the mustache should be three backticks 


## To Rendered in an RStudio **Shiny-server**

- Require svgR
- In the ui, set the svg output to be an htmlOutput
- In the server, 
  + convert the svg markup to text using *as.character*
  + set the corresponding output object to renderUI of an 
  *HTML* frame of the svg markp text.
  
  A short example is:

```
require(shiny)
svgMarkup<-svgR( circle(cxy=c(200,100),r=50, fill='red') )
app <- list(
  ui = bootstrapPage(  htmlOutput("svg") ) ,
  server = function(input, output) {
    output$svg<-renderUI( { HTML(
      as.character(svgMarkup)) } ) } 
) 
runApp(app) 
```
**Again Note**  we convert the svgMarkup to a character vector before rendering.

