---
title: "Set Up"
output:
  html_document:
    toc: true
    theme: united
---

```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```


```{r, echo=FALSE}
library(stringr)
showCode %<c-% function(txt){
  indentCount<-function(s){
    if(grepl(pattern="^ +", s)){
      nchar(str_match(pattern="^ +", s))   
    } else {
      0
    }
  }
  line2span<-function(s, default='black'){
    pattern='\\\\\\w+'
    #browser()
    if(grepl(pattern,s)){
      spns<-str_split(pattern=pattern,s)[[1]]
      cols<-as.character(str_match_all(pattern=pattern,s)[[1]])
      cols<-gsub("\\\\","",cols)
      if(length(cols)<length(spns)){
        cols<-c('black',cols)
      }
    } else {
      spns<-s
      cols<-'black'
    }
    lapply(1:length(spns), function(i){
      font.size=ifelse( cols[i]=="black",12,16 )
      stroke.width=ifelse( cols[i]=="black",.2,1 )
      tspan(spns[i], stroke=cols[i], fill=cols[i], font.size=font.size, stroke.width=stroke.width)
    })
  }
  
  lines<-str_split(pattern="\\n",txt1)[[1]]  
  lapply(1:length(lines), function(i){
    line<-lines[i]
    ic<-indentCount(line)
    xy<-c(ic*10+20, i*18+10)
    text(xy=xy, line2span(line))
  })
}

WH<-c(910,100)

codeBlock %<c-% function(wh){
  rect(xy=c(1,1), wh=wh-c(2,2), fill="#F4F4F4", stroke='#444', rxy=c(5,5), stroke.width=.3)
}
```

## Installing


To install svgR  we recommend using the **devtools** package.  That is,

A. Install R if R is not already installed.

B. Install the **devtools** package if the **devtools** package is not already installed.


This can be done by opening up an R session and then executing the following command:

 
```
install.packages("devtools")
```

C.  Use **devtools** to install **svgR** from **github**


This can be done by opening up an R session and then executing the following commands:

```
library(devtools)
install_github("mslegrand/svgR")
```


## To render in an RPres or R Markdown Document 

- Include prior to any calls to svgR include a block with

```{r, echo=FALSE, results="asis"}
txt1<-'\\`\\`\\`{r, echo=FALSE}
   \\purple library(svgR)
\\`\\`\\`
'
WH=c(800,80)
svgR( wh=WH, font.family='sans-serif', font.size=12, stroke.width=0.2, letter.spacing=2,
  codeBlock(WH),
  showCode(txt1)
)
```

### To render 

either 

- a simple call to svgR inside a block with results='asis'

```{r, echo=FALSE, results="asis"}
txt1<-' \\`\\`\\`{r, echo=TRUE, \\red results="asis"}
   \\purple svgR( \\black wh=c(600, 100), circle(cxy=c(200,50),r=50, fill="red") \\purple )
\\`\\`\\`
'
WH=c(800,80)
svgR( wh=WH, font.family='sans-serif', font.size=12, stroke.width=0.2, letter.spacing=2,
  codeBlock(WH),
  showCode(txt1)
)
```

```{r, echo=FALSE, results="asis"}
svgR(  wh=c(600, 100), circle(cxy=c(200,50),r=50, fill='red') )
```


or

- a call to svgR to create an svg markup object and echo back within a block with results='asis'

```{r, echo=FALSE, results="asis"}
txt1<-'\\`\\`\\`{r, echo=TRUE, \\red results="asis"}
   \\purple svgM<-svgR( \\black wh=c(600, 100), circle(cxy=c(200,50),r=50, fill="red") \\purple )
   \\purple svgM 
\\black \\`\\`\\`
'
WH=c(800,80)
svgR( wh=WH, font.family='sans-serif', font.size=12, stroke.width=0.2, letter.spacing=2,
  codeBlock(WH),
  showCode(txt1)
)
```

```{r, echo=F, results="asis"}
svgM<-svgR(  wh=c(600, 100), circle(cxy=c(200,50),r=50, fill='red') )
svgM
```

**Note** Note the mustache is  three backticks 


## To Rendered in an RStudio **Shiny-server**

- Require svgR
- In the ui, set the svg output to be an htmlOutput
- In the server, 
  + convert the svg markup to text using *as.character*
  + set the corresponding output object to renderUI of an 
  *HTML* frame of the svg markp text.
  
  A short example is:

```
require(shiny)
svgMarkup<-svgR( circle(cxy=c(200,100),r=50, fill='red') )
app <- list(
  ui = bootstrapPage(  htmlOutput("svg") ) ,
  server = function(input, output) {
    output$svg<-renderUI( { HTML(
      as.character(svgMarkup)) } ) } 
) 
runApp(app) 
```
**Again Note**  we convert the svgMarkup to a character vector before rendering.

