---
title: "Shapes"
output:
  html_document:
    toc: true
    theme: united
---



```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
if(!exists("newId")){
  source('Util.R')
}
```

In this section we discuss shapes.
Shapes are circles, ellipses, lines, paths, polygons, polylines, rectangles. 

##The Line
We begin with the most basic of shapes, a simple line segment
from a point xy1 to xy2

```{r, echo=T, results="asis"}
WH<-c(600, 100) # window rect
svgR( wh=WH,
      line( xy1=c(50,50), xy2=c(500,50), stroke='black')
)
```

Changing the color is easy to do, simply assign a color to the stroke parameter of line.


```{r, echo=T, results="asis"}
WH<-c(600, 100) # window rect
svgR( wh=WH,
      line( xy1=c(50,50), xy2=c(500,50), stroke='red')
)
```

This color can also be specified as #FF0000 or
rgb(255,0,0)

To change the line width simply set 
```{r, echo=T, results="asis"}
WH<-c(600, 100) # window rect
svgR( wh=WH,
      line( xy1=c(50,50), xy2=c(500,50), stroke='#00FF00', stroke.width=10)
)
```

To create a dashed line set the stroke.dasharray attribute
```{r, echo=T, results="asis"}
WH<-c(600, 100) # window rect
svgR( wh=WH,
      line( xy1=c(50,50), xy2=c(500,50), stroke='rgb(0,0,255)', stroke.dasharray=8 )
)
```

## The polyline

To produce a sequence of connected line segments use 
polyline. The points are specified with the points attribute.
Points can be specified in any of the following ways:

- a list of n vectors of length 2
- a n by 2 matrix
- a vector of the form $(x_1,y_1,x_2,y_2 ...x_n,y_n)$

Other attributes can be set as before.

```{r, echo=T, results="asis"}
WH<-c(600, 100) # window rect
svgR( wh=WH,
      polyline( 
        points=rbind(seq(from=20, to=500, by=50),c(20,80)), 
        stroke='#FF00AA', stroke.dasharray=8, fill='none' )
)
```


Note: we added an additional attribute fill='none'. If we don't, most browser will fill this with a black color. 

## The polygon

For closed figures, we use the polygon element. We have the same options to specify points as we did with the polyline.


```{r, echo=T, results="asis"}
WH<-c(600, 100) # window rect
points<-lapply((0:7)*(2*pi)/8, function(theta){
  c(100,50)+40*c(cos(theta),sin(theta))
} )
svgR( wh=WH,
      polygon( 
        points=points, stroke='red', fill='none' )
)
```

Additionally, we can supply a color to the fill variable:

```{r, echo=T, results="asis"}
WH<-c(600, 100) # window rect
points<-lapply(0:7, function(i){
  theta<-i*(2*pi)/8
  c(100,50)+20*(1+i%%2)*c(cos(theta),sin(theta))
} )
svgR( wh=WH,
      polygon( 
        points=points, stroke='blue', fill='pink' )
)
```

And remove the stroke

```{r, echo=T, results="asis"}
WH<-c(600, 100) # window rect
points<-lapply(0:15, function(i){
  theta<-i*2*pi/16
  c(100,50)+20*(1+i%%2)*c(cos(theta),sin(theta))
} )
svgR( wh=WH,
      polygon( points=points, fill='orange' )
)
```


## The rectangle, 
For a rectangle, (a four sided polygon) we use the rect element. We can specify the position of the upper-left corner of the rectangle in 2 ways:

- one at a time, for example: x=50, y=20 
- as a vector, for example: xy=c(50,20)

Alternative we can specify the center of the rectangle

- as a vector, for example cxy=c(100,40)

Also we can specify the dimensions in 2 ways:

- one at a time, for example: w=100, h=40 
- as a vector, for example:  wh=c(100,40)
```{r, echo=T, results="asis"}
WH<-c(600, 100) # window rect
svgR( wh=WH,
      rect( xy=c(50,20), wh=c(100,40), stroke='blue', fill='none' )
)
```

## The circle 
For a circle, we specify the center and the radius r. The center can be specified in 2 ways:

- one at a time, for example: cx=100, cy=60 
- as a vector, for example: cxy=c(100,60)


```{r, echo=T, results="asis"}
WH<-c(600, 150) # window rect
svgR( wh=WH,
  circle( cxy=c(120,60), r=25, stroke='blue', fill='none' )
)
```

Moreover, we can compose with rect using cxy to create concentric rectangles and circles:
```{r, echo=T, results="asis"}
WH<-c(600, 150) # window rect
cxy<-c(120,60)
dx<-10
svgR( wh=WH,
  circle( cxy=cxy, r=1*dx*sqrt(2), stroke='blue', fill='none' ),
  circle( cxy=cxy, r=2*dx*sqrt(2), stroke='blue', fill='none' ),
  circle( cxy=cxy, r=3*dx*sqrt(2), stroke='blue', fill='none' ),
  rect(cxy=cxy, wh=1*dx*c(2,2), stroke='red',fill='none'),
  rect(cxy=cxy, wh=2*dx*c(2,2), stroke='red',fill='none'),
  rect(cxy=cxy, wh=3*dx*c(2,2), stroke='red',fill='none')
)
```

However, a better way is to use lapply and iterate
```{r, echo=T, results="asis"}
WH<-c(600, 200) # window rect
cxy<-c(120,100)
dx<-10
n<-6
svgR( wh=WH,
  lapply(1:n, function(i)circle( cxy=cxy, r=i*dx*sqrt(2), stroke='blue', fill='none' )),
  lapply(1:n, function(i)rect(cxy=cxy, wh=i*dx*c(2,2), stroke='red',fill='none'))
)
```




## The ellipse

An ellipse is similar to a circle, but instead of r, we specify a pair rx and ry. This can be done in two ways:

- one at a time, for example: rx=50, ry=30 
- as a vector, for example:  rxy=c(50,30)

We demonstrate this with a filled ellipse inside a filled rect

```{r, echo=T, results="asis"}
WH<-c(600, 150) # window rect
cxy=c(200,75)
rxy<-c(50,30)
svgR( wh=WH,
  rect(cxy=cxy, wh=2*rxy, stroke='blue', fill='yellow'),
  ellipse(cxy=cxy, rxy=rxy, stroke='green', fill='orange' )
)
```


