---
title: "Text"
output:
  html_document:
  toc: true
theme: united
---
  


```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
if(!exists("newId")){
  source('Util.R')
}
```


```{r, echo=T, results="asis"}
WH=c(800, 400) # window rect
r=WH[2]/2
svgR( wh=WH,
  g(
    circle(cxy=WH/2, stroke.width=8, stroke="red", r=r),
    text('Data Rules', cxy=WH/2, font.size=80, stroke="red") 
    ) 
)

```

A good reference is:
  http://www.hongkiat.com/blog/scalable-vector-graphics-text/
  
  ### Simple Text
  ```{r, echo=FALSE}
tmp<-data.frame(
  Attribute=c(
    "font.family",
    "font.size",
    "font.style",
    "font.variant",
    "font.weight"),
  Values=c(
    "CSS font family",
    "pt, px, em, etc | inherit", 
    "(normal | italic | oblique | inherit)", 
    "(normal | small-caps | inherit)", 
    "(normal | bold | inherit | +others)"
  )
  ,
  Inital=c("Depends","medium","normal","normal","normal")
  ,
  Comment=c(
    "Name of font or font family.",
    "The font size.",
    "Style in which letters are rendered.",
    "'small.caps' makes all lower case letters render as small capitals.", 
    "The thickness of Attribute the font."
  )
)
```

### Font Styling

Attribute    | Values                              | Initial    | Comment
-------------|-------------------------------------|------------|----------------------------------------------------------
  font.family  | CSS font family                     | Depends    |Name of font or font family
font.size    | pt, px, em, %,  etc,   inherit      | medium     |The font size. Any SVG unit can be used.
font.style   |(normal\| italic\|oblique\| inherit) | normal     |Style in which letters are rendered.
font.variant |(normal\| small.caps \| inherit)     |normal      |'small.caps' sets lower case letters render to small caps.
font.weight  |(normal\|bold\|inherit\|+others)     |normal      | The thickness of the font.                               


#### Examining some font families using font.family

```{r, echo=T, results="asis"}
y<-40
fontName<-c("serif","sans-serif","cursive","fantasy","Arial","Times New Roman","Verdana", "Comic Sans")
WH=c(800,200)
svgR( wh=WH,
      graphPaper(WH, dxy=c(40,40), labels=TRUE),
      font.size=40, stroke='black',  stroke.width=0.5,
      lapply(1:8, function(i){
        text(fontName[i], xy=c(40+400*(i%%2), 40*floor((i+1)/2)), font.family=fontName[i], fill=rrgb())
      })
)
```


#### In addition, there we can specify a font style:

```{r, echo=T, results="asis"}
y<-40
WH=c(800,120)
fontStyle=c("normal" , "italic" , "oblique", "inherit")
svgR( wh=WH,
      font.style= "inherit",
      graphPaper(WH, dxy=c(40,40), labels=TRUE),
      font.size=40, stroke='black',  stroke.width=0.5,
      lapply(1:4, function(i){
        text(fontStyle[i], xy=c(40+400*(i%%2), 40*floor((i+1)/2)), font.style=fontStyle[i], fill=rrgb())
      })
)
```


### Text Positioning

```{r, echo=F}
refLines<-function(wh){
  x<-wh[1]/2
  y<-wh[2]/2
  list(
    line(xy1=c(0,y),xy2=c(wh[1],y), stroke='red', stroke.width=2),
    line(xy1=c(x,0),xy2=c(x,wh[2]), stroke='red', stroke.width=2),
    text("baseline" ,xy=c(5,y-5), stroke='red',  font.size=16),
    text("anchor" ,xy=c(x+5,wh[2]-5), stroke='red',  font.size=16)          )
}
baseLineAt<-function(y, wh){
  list(
    line(xy1=c(0,y),xy2=c(wh[1],y), stroke='red', stroke.width=2),
    text("baseline" ,xy=c(0,y-5), stroke='red',  font.size=16)
  )
}
anchorAt<-function(x, wh){
  list(
    line(xy1=c(x,0),xy2=c(x,wh[2]), stroke='red', stroke.width=2),
    text("anchor" ,xy=c(x,wh[2]-5), stroke='red',  font.size=16)
  )
}
```

#### text-anchor
By default, text begins at the x value, however the this can be modified by the text.anchor attribute (parameter). The text.anchor attribute can take one of four values: "start" , "middle" , "end" , "inherit". These are illustrated below.

```{r, echo=T, results="asis"}
WH=c(800,220)
textAnchor=c("start" , "middle" , "end" , "inherit")
svgR( wh=WH,
      graphPaper(WH, dxy=c(40,40), labels=TRUE),
      anchorAt(WH[1]/2,WH),
      lapply(1:4, function(i){
        g(
          baseLineAt(i*40,WH),
          text( textAnchor[i], xy=c(400, i*40), fill=rrgb(), text.anchor=textAnchor[i],
                font.size=40, font.weight="bold", stroke='black',  stroke.width=0.5)
        )
      }
      )
)
```


#### dominant.baseline
By default, text lies directly above the y coordinate (the baseline) This alignment can be adjusted by setting the dominant.baseline attribute (parameter).
The baseline is the y coordinate given to the text element. By default, the text is aligned
using the **alphabetic** option as the baseline. Other alignments are possible by setting the
alignment.baseline parameter, as illustrated in the following:
  #### alignment.baseline
  
  The baseline is the y coordinate given to the text element. By default, the text is aligned
using the **alphabetic** option as the baseline. Other alignments are possible by setting the
alignment.baseline parameter, as illustrated in the following:
  
```{r, echo=T, results="asis"}
WH<-c(800,80*9)
baseLine<-c( "alphabetic","ideographic","hanging","mathematical","central","middle",
             "text-before-edge","text-after-edge")
svgR( wh=WH,
      graphPaper(WH, dxy=c(40,40), labels=TRUE),
      anchorAt(WH[1]/2,WH),
      lapply(1:8, function(i){
        g(
          baseLineAt(i*80,WH),
          text( baseLine[i], xy=c(400, i*80), fill=rrgb(), dominant.baseline=baseLine[i],
                font.size=40, font.weight="bold", stroke='black',  stroke.width=0.5)
        )
      }
      )
)
```
#### tspan

Within a text element we can change the positioning, style, color of component words using the tspan element child element.

```{r, echo=T, results="asis"}
WH<-c(800,160)
svgR( wh=WH,
      graphPaper(WH, dxy=c(40,40), labels=TRUE),
      text( xy=c(40,120), lapply(1:14, function(i){
        tspan( LETTERS[i] , font.size=i*10, fill=rrgb() )}
      ))
)
```

```{r, echo=T, results="asis"}
WH<-c(800,160)
svgR( wh=WH,
      graphPaper(WH, dxy=c(40,40), labels=TRUE),
      text( xy=c(40,40), font.size=20, lapply(1:14, function(i){
        tspan( LETTERS[i] , dy=10*(i%%2), fill=rrgb() )}
      ))
)
```


#### baseline.shift for superscripts and subscriptss

Baseline.shift can be used to create super/subscripts.
Values for baseline.shift are

- baseline: not shift
- sub: subscript
- super: superscript
- percentage: a percentate
- length

  
```{r, echo=T, results="asis"}
WH<-c(800,120)
shift<-c( "auto","baseline","super","sub","60%", 6, "inherit")
bl<-80
svgR( wh=WH,
      graphPaper(WH, dxy=c(40,40), labels=TRUE),
      baseLineAt(bl,WH),
      lapply(1:7, function(i){
        text( "X",
              tspan( shift[i], baseline.shift=shift[i] , font.size=20),
              xy=c(i*120, bl), fill=rrgb(), 
              font.size=20, font.weight="bold", stroke='black',  stroke.width=0.5
        )
      }
      )
)
```


#### Math equations: Superscripts, subscripts



```{r, echo=T, results="asis"}
WH=c(800,240)
svgR( wh=WH,
      graphPaper(WH, dxy=c(40,40), labels=TRUE),
      refLines(WH),
      text( 
        tspan(mathSymbol("\\int"), font.size=120, dominant.baseline='mathematical'), 
        "e",
        tspan("-x", baseline.shift="60%", dominant.baseline="mathematical", font.size=40,
              tspan( "2", baseline.shift='super', font.size=20)),
        "dx",
        dominant.baseline='mathematical', xy=WH/2,  font.size=80, 
        stroke='black',  stroke.width=0.5, fill='lightblue'
      )
)
```

### Text Along a Path

Here we create text in a circle.
```{r, echo=T, results="asis"}
WH<-c(800, 600)
txt<-"The product of the circumferenc times the redius equals twice the area."
r=200
Pt1=WH/2+c(0,r)
Pt2=WH/2-c(0,r)
rxy=c(r,r)
svgR( wh=WH,
  defs( 
    path(id="C1",d=c("M", Pt1, "A", rxy, 0, 1, 0, Pt2, "A", rxy, 0, 1, 0, Pt1) , stroke='red')
  ), 
  graphPaper(WH, dxy=c(50,50), labels=TRUE),
  text( textPath(txt, font.size=40, stroke='black', stroke.width=0.5, fill='red', xlink.href="#C1")),
  circle(cxy=WH/2, r=90, stroke='red', fill='orange')
)
```

We can also spiral our text
```{r, echo=T, results="asis"}
WH<-c(800, 300)
txt<-"help, I'm getting dizzy!"
r=200
Pt1=WH/2+c(0,r)
Pt2=WH/2-c(0,r)
rxy=c(r,r)
svgR( wh=WH,
  defs( 
    path(id="spiralText",
         d = "M 50,50 A 20,20 0 0 1 90,50 A 40,40 0 0 1 10,50 A 60,60 0 0 1 130,50",
         stroke='blue')
  ), 
  graphPaper(WH, dxy=c(50,50), labels=TRUE),
  text( textPath(txt, font.size=30, stroke='black', stroke.width=0.5, fill='red', xlink.href="#spiralText"),
        transform=list(translate=c(100,100))
  )
)
```


### Reflection Transformation

```{r, echo=T, results="asis"}
WH=c(800,250)
svgR( wh=WH,
      graphPaper(WH, dxy=c(50,50), labels=TRUE),
      text( "Reflections", xy=c(0,0), 
            font.size=100, font.weight="bold", 
            stroke='black',  stroke.width=0.5, fill='red',
            transform=list(translate=c(100,100))
            ),
      text( "Reflections", xy=c(0,0), 
            font.size=100, font.weight="bold", 
            stroke='black',  stroke.width=0.5, fill='pink',
            transform = list(translate=c(100, 100), scale=c(1, -1) ) 
          )
)
```



#### Writing mode

The writing mode is useful to change text orientation to top-bottom
```{r, echo=T, results="asis"}
WH<-c(200,600)
svgR( wh=WH,
      graphPaper(WH, dxy=c(40,40), labels=TRUE),
      text( linearRegression ,  writing.mode="tb" , 
              font.size=100,
              xy=c(80, 80), fill='red',
              font.weight="bold", stroke='black',  stroke.width=0.5
      )
)
```

Text can be outlined or filled depending on what argument is supplied to the **fill** attribute.

```{r, echo=T, results="asis"}
WH=c(800,100)
svgR( wh=WH,
      graphPaper(WH, dxy=c(20,20), labels=TRUE),
      text( "Outlined Text", xy=c(60,80), 
            font.size=50, font.weight="bold", 
            stroke='black',  stroke.width=0.5, fill='none'),
      text( "Filled Text", xy=c(460,80), 
            font.size=50, font.weight="bold", 
            stroke='black',  stroke.width=0.5, fill='lightblue')
)
```




```{r, echo=T, results="asis"}
WH=c(800,120)
svgR( wh=WH,
      graphPaper(WH, dxy=c(50,50), labels=TRUE),
      text( "svgR has it All", xy=c(50,100), 
            font.size=100, font.weight="bold", 
            stroke='black',  stroke.width=0.5, fill='red')
)
```


