---
title: "ToDo-Bugs"
output: html_document
---


```{r, echo=FALSE}
library(svgR)
```
```{r, echo=FALSE}
  source('Util.R')
```

# Themes
cosmo looks nice
themes “default”, “cerulean”, 
“journal”, “flatly”, “readable”, 
“spacelab”, “united”, “cosmo”

http://appadvice.com/appnn/2015/10/nest-challenges-apples-homekit-makes-weave-open-to-devs

#Outline of doc

[Foo](#foo)

[Todo Topics:](#todo-topics)

[Todo examples:](#todo-examples)

[Todo svgR Core:](#todo-svgr-core)

[Bugs](#bugs)

[Other Todo:](#otherTodo)

[Anim](#Animation)

polaris slingshot
2.4 liter turbo
0 to 60 in 4 seconds
canam 


There was a young man from Trinity,
Who solved the square root of infinity.
While counting the digits,
He was seized by the fidgets,
Dropped science, and took up divinity.
~Author Unknown


An ounce of algebra is worth of a ton of verbal argument. -JBS Haldane

Data do not give up their secrets easily. They must be tortured to confess. -Jeff Hopper, Bell Labs


79.48% of all statistics are made up on the spot.
John Paulos, Prof of Math

All models are wrong, but some are useful. (George E. P. Box)


In God we trust. All others must bring data.
(W. Edwards Deming)


Statisticians, like artists, have the bad habit of falling in love with their models.
-- George Box

Conducting data analysis is like drinking a fine wine. It is important to swirl and sniff the wine, to unpack the complex bouquet and to appreciate the experience. Gulping the wine doesn’t work.
-Daniel B. Wright (2003), see PDF of Article.


#Todo Topics

- rework section on Shiny bevel, (is just before drop shadowns)
- Move filtering up??
- stories
- graphs
- animation timing
- zooming
- chaining transformations


#Todo examples:

- maps
- graphs
    + pie chart
    + bar chart
    + line graph
    + animated graphing
- timing demos
- animate scrolling texts (horizontal and vertical)
    + animate viewbox
- example pattern background?? see [Useful links](#Useful links)


#Todo svgR Core:


- add default for animateTransform: attributeName='transform'
    + Found 2 references that claim 'transform' is the only alternative
- all animateTransform to take transform lists.
    + additive='sum' is necessary if more than one transform at a time.
    +  Issues:
        + list(rotate=(), translate=(), scale=()), would do for each from, to
        + list( rotate=list(), translate=list(), scale=list() ), With each list
        (scale,tran,.. of the same length)
    

    
- whitespace-semicolon list for values for animate command

- should feConvolveMatrix need x, y, w, h, cxy? (It seems so)

-
#Useful links

- https://ledrug.wordpress.com/2010/09/30/learning-svg/
- http://franzheidl.github.io/keysplines/
- http://www.cheatography.com/kbarni/cheat-sheets/textadept7/
- http://www.nbcnews.com/health/health-news/getting-it-wrong-everyone-suffers-wrong-or-late-diagnosis-n431496

		<pattern id="checker" patternUnits="userSpaceOnUse"
				x="0" y="0" width="20" height="20">
			<rect x="0" y="0" height="20" width="20" fill="#aaaaaa"/>
			<rect x="0" y="0" height="10" width="10" fill="#999999"/>
			<rect x="10" y="10" height="10" width="10" fill="#999999"/>
		</pattern>

# ToDo: 

## General
- Break functons in Utils.R into seperate files
- Integrate
    +  **Anatomy of svg call** 
        + svgR call tree to SVG markup tree (with attributes)
        + list promotion
        + filters
    +  **Motivation of Grammar** (this seems to be included in anatomy)
        + SVG markup tree
        + function call
    +  **Structure**
        +  SVG markup tree
        +  indexing with ids
        +  tables with svgR  svg elements correspondence
        +  table with common svgR shortcuts

##Animation
- Lighting (spotlight, etc. doesn't look right) FIX THIS
- animate
- set
- animateMotion 
- animateColor
- animateTransform
- ToDo: Animate: Filters
- Todo: viewbox: (zooming)
- triggering
- transaltion, rotation, scaling
- along path
- color
- lighting
- Todo: viewbox: (zooming)
- events: ‘onbegin’, ‘onend’, ‘onrepeat’, ‘onload’
- example: anishapes.svg on my desktop
- example: http://franzheidl.github.io/keysplines/
- example: http://blog.pixelingene.com/2006/09/a-great-tool-for-visualizing-keysplines/
- example: https://docs.webplatform.org/wiki/svg/tutorials/smarter_svg_animation 


<a name="otherTodo"/>

## Other Todo: 
* Todo: KeyTimes, keyvalues, and keypoints?
* ToDo: Transforms
* ToDo: Section on coords and zooming
* Todo: feComponentTransfer - make and move it
* Todo: viewbox: (zooming)
* Todo: scrollbar (for zooming, panning)
* Todo: Lighting: Composing different lightins.
* Todo: Lighting at work
* Todo: Explain physics behind lighting.
* Todo: Use braces Util
* Todo: Javascript: scroll, pan, zoom, new object
* Todo: Javascript: keystoke responses
* Todo: Using jason data
* Todo: More Controls
    + video tape 
    + scrollbar
    + 

Animation
---

* Animate
* AnimateMotion
* AnimateTransform

#Bugs

* Bug: wsp-scl is being used for
    + path attribute d in animate,  to get 'm 0 1; c 1 2 3 4; m 0 1; c 1 2 3 4' <=ok
    + xyz attribute xyz in animate spotlight gives  x= '1 4 7', y= '2 5 8'
    + xyz needs x= '1; 4; 7', y= '2; 5; 8'
        + cxy has 2 numbers so we need an x and y
        + attrN='cxy' for to=list(c(1,2),c(3,4) ) want x=1;3 y=3;4
        + but the problem comes with values
            +  values may refer to an xy
            +  values may refer to an d
    
* Bug: feComponeTransfer
* Bug: viewBox text is not centered in Viewports and Viewboxes
* Bug: Keytimes keypoints do not respond to c(,,)
* But: keySplines need to be space seperated, not ";" seperated!
* Bug: Animate combo params
* Bug: Animate gradient
* Bug: cxy and $\\infty$ didn't work together
* need to put def-like elements to the front of sibling list.

* Bug: add xyChannelSelector for feDisplacementMap 
currently we have xChannelSelector
* Bug: feDiffuseLighting: check if we really want our fix


```{r, echo=T, results="asis"}
WH<-c(800,160)
svgR( wh=WH,
  graphPaper(WH, dxy=c(40,40), labels=TRUE),
  brace(c(100,100),c(200,50),c(200,150))
)
```

# What are the 5 Extinctions


#Todo:logo
```{r, echo=T, results="asis"}
library(svgR)
WH=c(800, 250) # window rect
svgR( wh=WH,  
      rect(xy=c(0,0), wh=WH, fill=radialGradient( colors=c("white","red","blue","black"))),
      g(
        text( "svgR", cxy=WH/2, fill="darkblue", font.size=100, 
              font.family="san serif", stroke.width=3),
        stroke='black',
        filter = filter(
          feMerge( 
            feMergeNode(in1="SourceGraphic"),
            feMergeNode(
              in1=feComposite( operator='in',
                               in1=feSpecularLighting( surfaceScale=6,
                                                       specularConstant=1,
                                                       specularExponent=30,
                                                       lighting.color="yellow",
                                                       in1=feGaussianBlur(
                                                         stdDeviation=5,
                                                         in1="SourceAlpha"),
                                                       fePointLight(xyz=c(40,-30,200))
                               ),
                               in2="SourceAlpha"
              )
            )
          )
        )
      )
)->logo
print(logo)
```

#Notes

# Bug: feMerge in dropShadow using width

- with width=200 feMerge delivers good sharp image
- with w=200 feMerge delivers a blurry image
- without width feMerge delivers a blurry image
- with wh=c(200,60) feMerge fails to deliver any image


# arrow poly: Util
A fat arrow (filled) to use with block diagrams, to explain
- Svg overview (compile, choices, ,,)
- filter processing

# labeled rect, circle, equalateral triangle (orient=ULRD), diamond,


For diagram

# Todo logo

```{r, echo=T, results="asis"}
library(svgR)
WH=c(800, 250) # window rect
svgR( wh=WH,     
      g(
        text( "svgR", cxy=WH*c(.25,.5), fill="darkblue", font.size=100, 
              font.family="san serif", stroke.width=3),
        text( "Users Guide", cxy=WH*c(.7,.25), fill="darkblue", font.size=60, 
              font.family="fantasy", stroke.width=3),
        circle(cxy=WH*c(.25,.5), r=99, stroke.width=10, fill='none'  ),
        stroke='black',
        filter = filter(
          feMerge( 
            feMergeNode(in1="SourceGraphic"),
            feMergeNode(
              in1=feComposite( operator='in',
                               in1=feSpecularLighting( surfaceScale=6,
                                                       specularConstant=1,
                                                       specularExponent=30,
                                                       lighting.color="yellow",
                                                       in1=feGaussianBlur(
                                                         stdDeviation=5,
                                                         in1="SourceAlpha"),
                                                       fePointLight(xyz=c(40,-30,200))
                               ),
                               in2="SourceAlpha"
              )
            )
          )
        )
      )
)->logo
print(logo)
```

# Todo: add classes to the construction

# Todo: path=path
How to wrap the path attribute? 
It will need to be inside either

- defs
- symbol

#ToDo:  Some Fun with feComposite
Let's return to our shadow example

<feComposite operator=”over” in=”SourceGraphic” in2=”dropBlur” result=”final”/>

We change the operator to 

#ToDo: breakup rmd and put in new directory

# Fix: feTile


```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
     symbol(id='svgRaves', viewbox=c(0,0,"100%","100%"),
            text('svgR', xy=c(0,30), stroke='red')),
     use(filter( xy=c(0,0),wh=c("100%","100%"),
            feTile( in1="svgRaves")
          ))
)
```

# ToDo Section on symbol, use

Discuss

- symbol
- use

```{r, echo=T, results="asis"}
WH=c(800, 100) # window rect
svgR( wh=WH,
      symbol(id='svgRocks',
               text('svgR', xy=c(0,30), stroke='red')
      ),
      use( xlink.href="#svgRocks", xy=c(400,50) ),
      use( xlink.href="#svgRocks", xy=c(440,50) )
)
```

# Todo: Utility

braces util


Shapes
===

The shapes are circles, ellipses, lines, paths, polygons, polylines, rectangles. 
Shapes are customizable by

- fill
- transform
- stoke, width, ...
- animate
- filter
- mask
- clippath


End of Doc
===

Elements to Fill
===
Both shapes and text can be filled.
- Shapes we have seen so far are:
 + rectangles, circles, ellipses
- Other shapes are
 + lines
 + polygons
 + polylines
 + paths


Defs
===
Defs are used as a location to define reuseable graphical entities such as
- gradients
- patterns
- markers
- shapes


Functions to add???
===
```
rotate<-function(v){paste0("rotate(",paste(v,collapse=","),")")}
usage: transform=rotate(...)
(instead fo transform=list(rotate=...))
```

```
polygon( x= ..., y=...) or list(x=, y=) (as inspired from data.frame)
```





## lines,

## polylines, 
## paths, 
## polygons, 
## rectangles
## circles,
## ellipses, 

# Add content

#### Rexamining the tree

- The nodes having tags **circle**,**line**,**rect**,**polygon**,**polyline**,**ellipse** are the shape wlements Shapese elements are used to render corresponding shape. Shape nodes are contained in the **container** elements
- The nodes **svg**, **g** and **defs** are the tags for the container elements. Container eements are nodes which can have as child elements both graphics elements and other container elements. The container elements are: **a**, **defs**, **glyph**, **g**, **marker**, **mask**, **missing-glyph**, **pattern**, **svg**, **switch** and **symbol**.
- The **defs** node contain definitions. Nodes contained in a **defs** node are not rendered directly, but when referenced to by other nodes can be used to provide additional information or instructions for rendering.
- The nodes **mask**, **linearGradient**, **filter**, **marker** are not rendered directly, but rather are used to modify the rendering of shape nodes. These are usually placed in a **defs** node and referenced by the shapes that use them.  However, in svgR we for convenience we provide an alternate approach. 
That is, svgR allows a shape to have the modifying node as a child. This feature is particulary convenient when those nodes are to be applied to a single shape. (In reality, behind the scenes, svgR moves the nodes and references them in order to be compliant with the WC3 svg standard.)

# Todo: import text font
```
defs<-getNodeSet(tmp,"//defs")
filter<-getNodeSet(tmp,"//defs/filter")
filter.attrs<-xmlAttrs(filter[[1]])
filter.attr.names<-names(filter.attrs)
if("id" %in$ filter.attr.names){
  rtv.id<-filter.attrs["id"]
}

<defs>
  <style type="text/css">@import url('http://fonts.googleapis.com/css?family=Lobster|Fontdiner+Swanky|Crafty+Girls|Pacifico|Satisfy|Gloria+Hallelujah|Bangers|Audiowide|Sacramento');</style>
  </defs>
  
  <link href='http://fonts.googleapis.com/css?family=Sancreek' rel='stylesheet' type='text/css'>
```  
testing 


```{r, echo=FALSE}
library(svgR)
```


testing 2

```{r, echo=T, results="asis"}
library(XML)
y<-40
WH=c(800,120)
fontInfo<-"
  @font-face {
    font-family: Lato;
    src: url('http://fonts.googleapis.com/css?family=Lato:400,700'); 
  }
"  
svgR( wh=WH,
  defs(
    style( type="text/css", xmlCDataNode(value=fontInfo))
  ),
  graphPaper(WH, dxy=c(40,40), labels=TRUE),
  text( "Text using CSS @font-face",
    font.size=40, stroke='black',  #font.style='Spice Rice', 
    font.family='Lato',
    #style="font-family: 'Lato'; font-weight:normal; font-style: normal",
    xy=c(40, 40),  fill='red')
)
```
testing 3

```{r, echo=T, results="asis"}
library(XML)
y<-40
WH=c(800,120)
fontInfo<-"
  @font-face {
    font-family: Sancreek;
    src: url('http://fonts.googleapis.com/css?family=Sancreek'); 
  }
"  
svgR( wh=WH,
  defs(
    style( type="text/css", xmlCDataNode(value=fontInfo))
  ),
  graphPaper(WH, dxy=c(40,40), labels=TRUE),
  text( "Text using CSS @font-face",
    font.size=40, stroke='black',  #font.style='Spice Rice', 
    font.family='Sancreek',
    #style="font-family: 'Lato'; font-weight:normal; font-style: normal",
    xy=c(40, 40),  fill='red')
)
```



feDisplacementMap 2

```{r, echo=T, results="asis"}
library(svgR)
WH=c(400, 400) # window rect
svgR( wh=WH, 
      symbol(
        rect( id="radDisp23", xy=c(0,0), wh=WH, 
          fill=radialGradient( 
            xy1=c(0,0), xy2=c(1,1), 
            colors=c("red","blue")
          )
        )
      ),
      use(xlink.href="#radDisp23"),
      g(  
        graphPaper(WH, dxy=c(40,40), labels=TRUE),
          filter=
           filter(  filterUnits='userSpaceOnUse',
              feMerge(
                in1=feMergeNode(
                   feDisplacementMap( xy=c(0,0), wh=WH/2,
                    in1="SourceGraphic",
                    in2=feImage(xlink.href='#radDisp23'),
                    scale=50,
                    xChannelSelector="R",
                    yChannelSelector="R"
                  )                   
                ),
                 in1=feMergeNode( 
                   feDisplacementMap( xy=WH/2, wh=WH/2,
                    in1="SourceGraphic",
                    in2=feImage(xlink.href='#radDisp23'),
                    scale=50,
                    xChannelSelector="B",
                    yChannelSelector="B"
                  )                   
                ),
                 in1=feMergeNode( 
                   feDisplacementMap( xy=c(1,0)*WH/2, wh=WH/2,
                    in1="SourceGraphic",
                    in2=feImage(xlink.href='#radDisp23'),
                    scale=50,
                    xChannelSelector="R",
                    yChannelSelector="B"
                  )                   
                ),
                 in1=feMergeNode( 
                   feDisplacementMap( xy=c(0,1)*WH/2, wh=WH/2,
                    in1="SourceGraphic",
                    in2=feImage(xlink.href='#radDisp23'),
                    scale=50,
                    xChannelSelector="B",
                    yChannelSelector="R"
                  )                   
                )
          )
        )
      )
)
```




```{r, echo=T, results="asis"}
WH<-c(600,100)

svgR( wh=WH,
    rect(xy=c(0,0),wh=WH, fill="rgb(192,224,255)"),
    playButton( xy=WH),
    circle( cxy=c(20,20), r=10, stroke="black", fill="yellow"),
    g(
      circle( cxy=c(20,40),r=10, stroke="black", fill="blue"),
      animateMotion(from=c(0,0), to=c(WH[1],0), 
                    begin=playButton.click, dur=10, fill="freeze") 
    )
)
```

```{r, echo=T, results="asis"}
WH<-c(600,200)
svgR( wh=WH,
    graphPaper(wh=WH),
    rect(xy=c(0,WH[2]-40),  wh=c(WH[1],40), 
         fill= #'lightblue'
            linearGradient(
              xy1=c(0,0), xy2=c(0,1),
              colors=c("lightblue","white","lightblue") 
            )
    ),
    playButton(xy=WH),
    rect(xy=c(0,0),wh=WH, fill='none', stroke='blue', stroke.width=3),

    circle( cxy=c(20,20), r=10, stroke="black", fill="red"),
    g(
      circle( cxy=c(20,40),r=10, stroke="black", fill="green"),
      animateMotion(from=c(0,0), to=c(WH[1],0), 
                    begin=playButton.click, dur=10, fill="freeze") 
    )
)
```


#### Text Span size
```{r, echo=T, results="asis"}
WH<-c(800,200)
txt<-"svgR"
txt=unlist(strsplit(txt,""))
N<-length(txt)
ids<-paste0("animateTspan",1:N)
ids.click<-c(playButton.click, ids)
svgR( wh=WH,
    playBar(WH),
    text( "", cxy=WH/2, stroke="black", fill="lightgreen",
          lapply(1:N, function(i){
            tspan( txt[i], font.size=45, stroke="black", fill="red",
                   set(id=ids[i], 
                      attributeName='fill', from='black', to='red', 
                      begin=ids.click[i], dur=.5*(1+N-i)
                   )
            )
          })
     )
)
```


#olde examples

```{r, echo=TRUE, results="asis"}
myNames<-strsplit("animate animateColor animateMotion animateTransform set desc metadata title circle ellipse line path polygon polyline rect defs g svg symbol use linearGradient radialGradient altGlyph textPath tref tspan feBlend feColorMatrix feComponentTransfer feComposite feConvolveMatrix feDiffuseLighting feDisplacementMap feFlood feGaussianBlur feImage feMerge feMorphology feOffset feSpecularLighting feTile feTurbulence mask filter text marker"," ")[[1]]

#myNames<-myNames[1:10]
id<-1+1:length(myNames)

tree<-as.data.frame(cbind(
  parent = rep(1, length(myNames)),
  child  = id
))
nodeNames<-c('root svg',myNames)

nodeColors<-autoNodeColor(nodeNames)
WH=c(600, 1800)
svgR( wh=WH,
  svgRTree( tree, nodeNames, nodeColors=nodeColors, nodeWH=c(160,20))
)

```


# ToDo: insert tree graphs
```{r, echo=TRUE, results="asis"}
# tree<-as.data.frame(cbind(
#   parent = c(1, 1, 2, 2, 2, 2, 3, 3, 6,  6,  3,  3,  1  ),
#   child  = c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14 )
# ))
tree<-as.data.frame(cbind(
  parent = c(1, 1,  2, 2, 2, 3, 3, 6,  6,  3,  3,  1  ),
  child  = c(2, 3,  5, 6, 7, 8, 9, 10, 11, 12, 13, 14 )
))

#Positions in Nodes names determines the id
# nodeNames<-c("svg (root)","defs","g","mask","linearGradient","filter","marker", "circle","line","rect","polygon","polyline","ellipse","circle")

nodeNames<-c("svg","defs","g","mask","linearGradient","filter","marker", "circle","line","feGaussianBlur","feMerge","polyline","ellipse","text")

# nodeColors<-c('#AFFFA0','#AFFFA0','#AFFFA0','#FFFF50','#FF80FF','pink',,
#               '#D0D0FF','#D0D0FF','#D0D0FF','#D0D0FF','#D0D0FF','#D0D0FF','#D0D0FF')

nodeColors<-autoNodeColor(nodeNames)
WH=c(800, 600)
svgR( wh=WH,
  svgRTree( tree, nodeNames, nodeColors=nodeColors, nodeWH=c(110,20))
)

```

# Foo

```{r, echo=T, results="asis"}
library(XML)
y<-40
WH=c(800,320)
myFontInfo<-"
  @font-face {
    font-family: Delicious;
    src: url('http://nimbupani.com/demo/svgfonts/delicious-roman.otf');
  }
"  
svgR( wh=WH,

  #graphPaper(WH, dxy=c(40,40), labels=TRUE),
  text( "@ ☢ ☣ ☤ CSS font-face",
    font.size=200, stroke='black',  #font.style='Spice Rice', 
#    style="font-family: 'Delicious'; font-weight:normal; font-style: normal",
    xy=c(40, 240),  fill='red')
)
```


