---
title: "Transforms"
output:
  html_document:
    toc: true
    theme: united
---


```{r set-options, echo=FALSE, cache=FALSE}
#options(width = SOME-REALLY-BIG-VALUE)
```

```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```

```{r, echo=FALSE}
if(!exists("newId")){
  source('Util.R')
}
```

## **Transform** is an Attribute 

The transform attribute allows for multiple linear operations
to be applied prior to rendering.

SVG supports the composition 5 types transforms:

- translate: Translates the orgin 
- rotate: Rotates through a given angle about a given point
- scale: 
- skew
- matrix

We discus translate, rotate, scale, skew in this section and leave matrix to the next section.

The transform attribute is assigned a value that consists
of a named list. The name list entry name provides the type of operation, and list entry value give the amount.

For example
``
transform=list(translate=c(20,30),scale=c(1,2))
``
translates and then scales.  The transorm is sensitive to the order of operations, that is, transform operations are seldom commute.


In the remainging of this section we give some examples, but first we build a user defined element to illustrating those transforms. 

**Remember** A *transform* is not an element, it is an attribute, like *xy*. 

## The Tale of Two Circles, a Square and some Lines

### The cirCirSqr element

To illustrate the effects of transform we first create a custom element, for our illustration. It will consist of an ensemble of

- a border cosnisting of dashed lines , colored red on the top, left, colored green on the bottom right
- blue dash lines intersecting at the center
- a semi-transparent blue rectangle with it's upper left corner in the center
- a semi-transparent red circle centered on the upper left corner of the blue rectangle
- a semi-transparent gree circle centered on the upper right corner of the blue rectangle
```{r}
# A function to tell the tale of two circles and a rectangle
cirCirSqr<-function(...){
  attrs<-list(...)
  stopifnot("cxy" %in% names(attrs))
  centerPt<-attrs[['cxy']]
  attrs$cxy<-NULL
  wh<-c(100,50) #sets the rectangle dimension
  pt1<-centerPt+c(.5,0)*wh
  pt2<-centerPt+c(-.5,0)*wh
  g(
    attrs,
    line(xy1=c(0,0), xy2=c(WH[1],0), stroke='red', stroke.dasharray=10, stroke.width=4),
    line(xy1=c(0,0), xy2=c(0,WH[2]), stroke='red', stroke.dasharray=10, stroke.width=4 ),
    line(xy1=WH, xy2=c(WH[1],0), stroke='green', stroke.dasharray=15, stroke.width=4),
    line(xy1=WH, xy2=c(0,WH[2]), stroke='green', stroke.dasharray=15, stroke.width=4 ),
    line(xy1=c(0,WH[2]/2), xy2=c(WH[1],WH[2]/2), stroke='blue', stroke.dasharray=10),
    line(xy1=c(WH[1]/2,0), xy2=c(WH[1]/2,WH[2]), stroke='blue', stroke.dasharray=10 ),
    rect(xy=pt2, wh=wh,  stroke ='black', fill='blue',  opacity=.3),
    circle(cxy=pt1, r=wh[2], stroke="black", fill="red",   opacity=.3),
    circle(cxy=pt2, r=wh[2], stroke="black", fill="green", opacity=.3),
    text(paste("(",centerPt[1],",",centerPt[2],")"), cxy=centerPt, font.size=12)
  )
}
```

And also for convenience we create a function for showing a background of graphPaper and and a title. 
```{r, echo=T}
# Utility to cover my back
myBack<-function(WH, txt){
  list(wh=WH, #sets the viewPort dimension
      text(txt, xy=c(20,25), font.size=18),
      graphPaper(wh=WH, dxy=c(50,50), labels=TRUE),
      line(xy1=c(0,WH[2]/2), xy2=c(WH[1],WH[2]/2), stroke='blue' ),
      line(xy1=c(WH[1]/2,0), xy2=c(WH[1]/2,WH[2]), stroke='blue' )
  )
} 
```

Ok, now let's see what this looks like

#### Without Any Transform

```{r, echo=T, results="asis"}
WH=c(800, 400) # window rect
svgR( 
  myBack(WH, "Without any Transform"),
  cirCirSqr(cxy=WH/2)
)
```


### Translating

We want: 
  
  - to transform the circles, square, lines in cirCirSqr
  - not to transform the background in myBack

This suggests grouping the circCirSqr with the transform.

#### translate=c(100,40)

```{r, echo=T, results="asis"}
WH=c(800, 400) # window rect
svgR(
  myBack( WH, "With Transform translate=c(100,40)" ),
  cirCirSqr(cxy=WH/2, transform=list(translate=c(100,40) ) ) 
)
```

#### translate=c(-100,-40)

```{r, echo=T, results="asis"}
WH=c(800, 400) # window rect
svgR( 
  myBack(WH, "With Transform translate=c(-100,-40)"),
  cirCirSqr(cxy=WH/2, transform=list(translate=c(-100,-40))) 
)
```


### Scaling

Scaling can be applied horizontally, vertically, or both. A factor of 1 leaves the figure unscaled. A facor greater than 1 magnifies and factor less that 1 reduces. In our examples, we use values less than 1, so that the entire image can be seen.


#### Scaling about the Origin


We begin by presenting two examples: 

- The first scales by a factor of $\frac{1}{2}$ horizontally

```{r, echo=T, results="asis"}
WH=c(800, 400) # window rect
svgR( 
  myBack(WH, "With Transform scale=.5"),
  cirCirSqr(cxy=WH/2, transform=list(scale=c(.5,1))) 
)
```

- The seconds scales in both directions by a factor of $\frac{1}{2}$ 

```{r, echo=T, results="asis"}
WH=c(800, 400) # window rect
svgR( 
  myBack(WH, "With Transform scale=.5"),
  cirCirSqr(cxy=WH/2, transform=list(scale=.5) ) 
)
```

**Note:** The dimension of our figure have been reduce by $\frac{1}{2}$ and the distance to origin has been halved.  This is because scaling is scales using the origin as it's fixed point. (Other points change)

#### Scaling about a Given Fixed Point

To scale about a given point pt, we move pt p to the center, scale and then move back.
More precicely, if p is the fixed point and $\alpha$ the scale factor, then 

1. translate by -p
2. scale by $\alpha$
3. translate back by  $\frac{2p}{\alpha} -p$

For example:

```{r, echo=T, results="asis"}
WH=c(800, 400) # window rect
alpha<-.3
svgR( 
  myBack(WH, "Transform: scale=0.3 with center fixed"),
  cirCirSqr(cxy=WH/2, transform=list(translate=-0.5*WH, scale=alpha, translate=WH*(1/alpha -1/2)))
)
```

Above we scale with a fixed center, below we start with an image centered at c(.25,.25)*WH and
scale with a value of .3, with a fixed point of c(.25,.25)*WH.

```{r, echo=T, results="asis"}
WH=c(800, 400) # window rect
alpha<-.3
pt<-c(.25,.25)
svgR( 
  myBack(WH, "With Transform scale=0.3 and fixed point c(.25,.25)"),
  cirCirSqr(cxy=pt*WH, transform=list(translate=-pt*WH, scale=alpha, 
                                  translate=WH*(2*pt/alpha - pt)))
)
```

#### Reflections

Reflections are a special case of scaling. Just use a negative as a scale factor. We give 2 examples:

##### Vertical Reflection though the center, 
```{r, echo=T, results="asis"}
WH=c(800, 400) # window rect
alpha<-c(1,-1)
pt<-c(.5,.5)
svgR( 
  myBack(WH, "Reflection scale=c(1,-1) and fixed point=c(0,0)"),
  cirCirSqr(cxy=pt*WH, transform=list(translate=-pt*WH, scale=alpha, 
                                  translate=WH*(2*pt/alpha - pt)))
)
```


### Rotations

Like scaling, rotations default to having the origin be the fixed point.

#### Rotation about the Origin: 

This is simple rotate by 10 degrees (For those not familiar with degrees, thats $\frac{pi}{18}$)

```{r, echo=T, results="asis"}
WH=c(800, 400) # window rect
degrees<-10
svgR(  
  myBack(WH, paste0("With Transform rotate=",degrees)),
  g(cirCirSqr(cxy=WH/2),
     transform=list(rotate=degrees)
  )
) 
```

**Observer:** Rotation is about the origin.

#### Rotation about a Fixed Point


##### transform=list(rotate=c(5,400,200))

Althought we can try an approach similar to what was done for scaling, we fortunately have
an easy alternative, simple supply the position of the fixed point as the second and third  coordinates:

Here we rotated about the center of our viewPort.
```{r, echo=T, results="asis"}
WH=c(800, 400) # window rect
svgR( 
  myBack(WH, "With Transform rotate=c(5, WH/2)"),
  g(cirCirSqr(cxy=WH/2),
     transform=list(rotate=c(5, WH/2))
  )
)
```


### Skewing
There are two skews available: skewX, skewY.
They are illustrated as following:

#### skewX=30

SkewX (a shear stress) by an angle of $\frac{\pi}{6}$

```{r, echo=T, results="asis"}
WH=c(800, 400) # window rect
svgR( 
  myBack(WH, "With Transform skewX=30"),
  g(cirCirSqr(cxy=WH/2),
     transform=list(skewX=30)
  )
)
```

Now skewY (a shear stress) by an angle of $\frac{\pi}{36}$


```{r, echo=T, results="asis"}
WH=c(800, 400) # window rect
svgR( 
  myBack(WH, "With Transform skewY=5"),
  g(cirCirSqr(cxy=WH/2),
     transform=list(skewY=5)
  )
)
```


#### Combining Transforations

We have already seen an example of combining transformations. However
we a few more just to demonstrate order matters!

```{r, echo=T, results="asis"}
WH=c(800, 400) # window rect
svgR( 
  myBack(WH, "Order Matters: rotate, translate"),
  g(cirCirSqr(cxy=WH/2),
     transform=list(rotate=c(45,WH/2), translate=c(100,0))
  )
)
```

```{r, echo=T, results="asis"}
WH=c(800, 400) # window rect
svgR( 
  myBack(WH, "Order Matters: translate, rotate"),
  g(cirCirSqr(cxy=WH/2),
     transform=list(translate=c(100,0), rotate=c(45,WH/2))
  )
)
```

### Tranform Attribute using Matrix as a Value


In the previous section, we discussed using the transform with the operation types: tranlate, scale, rotate, and skew. In this section we discuss using matrix opertion.  

The advantage to using matrix is two fold

- we can combine several operations into one
- we need only know the desired effect on 3 points to get our specification (less thinking involved) 

But first some background

#### Linear Algebra

A basic result in linear algebra is any function mapping between vector spaces that satisifies
$$T(\alpha \vec v_1 + \alpha \vec v_2)=\alpha T ( \vec v_1 + \vec v_2 ) $$
can be represented by a matrix. Considering points (x,y) on the screen as elements in a 2 dimensional vector space we can easlily see that (rotations, scaling,  fit the criteria. However translation does not. But all is not lost. We can envision that our 2-d space is just the plane in 3-space defined by z=1. With trick, translation can be created (on that plane) using a shear stress, which does satisfy the above requirement. 

Essentially the trick is sending (x,y) to (x,y,1). Then the effect on points using rotation, scaling, translation, and skewing can be constructed as simple matrix muliplication of the form:
\[
  \begin{pmatrix} x_{new} \\ y_{new} \\ 1 \end{pmatrix}
  =
  \begin{pmatrix} a & c & e \\ b & d & f \\ 0 & 0 & 1\end{pmatrix} 
  \begin{pmatrix} x_{old} \\ y_{old} \\ 1 \end{pmatrix}
\]

For example, translation is performed by
\[
  \begin{pmatrix} x_{new} \\ y_{new} \\ 1 \end{pmatrix}
  =
  \begin{pmatrix} 1 & 0 & h \\ 0 & 1 & k \\ 0 & 0 & 1\end{pmatrix} 
  \begin{pmatrix} x_{old} \\ y_{old} \\ 1 \end{pmatrix}
\]

and rotation (about the origin) is given by
\[
  \begin{pmatrix} x_{new} \\ y_{new} \\ 1 \end{pmatrix}
  =
  \begin{pmatrix} cos(\theta) & sin(\theta) & 0 \\ -sin(\theta) & cos(\theta) & 0 \\ 0 & 0 & 1\end{pmatrix} 
  \begin{pmatrix} x_{old} \\ y_{old} \\ 1 \end{pmatrix}
\]

and scaling by
\[
  \begin{pmatrix} x_{new} \\ y_{new} \\ 1 \end{pmatrix}
  =
  \begin{pmatrix} S_x & 0 & 0 \\ 0 & S_y & 0 \\ 0 & 0 & 1\end{pmatrix} 
  \begin{pmatrix} x_{old} \\ y_{old} \\ 1 \end{pmatrix}
\]


#### Scaling with a Given Fixed Point

Now the transformation is always of the form
\[
  \begin{pmatrix} x_{new} \\ y_{new} \\ 1 \end{pmatrix}
  =
  \begin{pmatrix} a & c & e \\ b & d & f \\ 0 & 0 & 1\end{pmatrix} 
  \begin{pmatrix} x_{old} \\ y_{old} \\ 1 \end{pmatrix}
\]
with the last column always being
\[
  \begin{pmatrix} 0\\ 0 \\ 1 \end{pmatrix}
\]

Thus in practice, only the $a,b,c,d,e,f$ need be supplied. This is done either 
by
```
transformation=list(matrix=matrix(c(a,b,c,d,e,f)2,3))
```
or
```
transformation=list(matrix=c(a,b,c,d,e,f))
```

Now the transformation matrix $A$ is determined by the values $Y_i=A[X_i]$ on any three collinear points $X_i$. In particular if X and Y the matrices whose columns are the $X_i$ and $Y_i$, then $X$ is non-singular and $Y= A X$. From this we solve for A.

We given two examples,  the first being scaling in about a fixed point.

Let $\alpha=(\alpha_x, \alpha_y)$ be scaling factors in the x and y direction. Let $p=(p_x,p_y)$ be the fixed point. Then we have
\[
  \begin{pmatrix} p_x+1 \\ p_y  \end{pmatrix}
  \mapsto
  \begin{pmatrix} p_x+\alpha_x \\ p_y  \end{pmatrix}, \qquad
  \begin{pmatrix} p_x \\ p_y+1  \end{pmatrix}
  \mapsto
  \begin{pmatrix} p_x \\ p_y+\alpha_y  \end{pmatrix}, \qquad
  \begin{pmatrix} p_x \\ p_y  \end{pmatrix}
  \mapsto
  \begin{pmatrix} p_x \\ p_y  \end{pmatrix}
\]
Appending 1's and placing into matrix form we have
\[
  \begin{pmatrix} 
  p_x+1 & p_x   & 0 \\
  p_y   & p_y+1 & 0 \\
  1   & 1   & 1 
  \end{pmatrix} 
  \mapsto
  \begin{pmatrix} 
  p_x+\alpha_x   & p_x        & 0 \\
  p_y          & p_y+\alpha_y & 0 \\
  1          & 1              & 1 
  \end{pmatrix}
\]
So
\[
  \begin{pmatrix} 
  p_x+\alpha_x   & p_x        & 0 \\
  p_y          & p_y+\alpha_y & 0 \\
  1          & 1              & 1 
  \end{pmatrix}
  =
  \begin{pmatrix} 
  a & c & e \\
  b & d & f \\
  0 & 0 & 1 
  \end{pmatrix} 
  \begin{pmatrix} 
  p_x+1 & p_x   & 0 \\
  p_y   & p_y+1 & 0 \\
  1   & 1   & 1 
  \end{pmatrix} 
\]
Taking transposes
\[
  \begin{pmatrix} 
  p_x+\alpha_x   & p_y        & 1  \\
  p_x          & p_y+\alpha_y & 1 \\
  0         & 0        & 1 
  \end{pmatrix}
  =
  \begin{pmatrix} 
  p_x+1 & p_y   & 1 \\
  p_x   & p_y+1 & 1  \\
  0    & 0  & 1 
  \end{pmatrix} 
  \begin{pmatrix} 
  a & b & 0 \\
  c & d & 0 \\
  e & f & 1 
  \end{pmatrix} 
\]

Setting up the augmented matrix
$$
  \left[\begin{array}{rrr|rrr}
  p_x+\alpha_x   & p_y        & 1  & p_x+1 & p_y   & 1 \\
  p_x          & p_y+\alpha_y & 1 & p_x   & p_y+1 & 1  \\
  0         & 0        & 1 & 0    & 0  & 1 
  \end{array}\right]
$$

This reduces to 
$$
  \left[\begin{array}{rrr|rrr}
  \alpha_x   & 0        & 0  & 1 & 0   & 0 \\
  0          & \alpha_y & 0  & 0  & 1  & 0  \\
  p_x(1-\alpha_x)   & P_y(1-\alpha_y)   & 1 & 0    & 0  & 1 
  \end{array}\right]
$$

So the transformation matrix value is
$$ 
  \begin{pmatrix} 
  a   &  c & e   \\
  b   &  d & f  
  \end{pmatrix} =
  \begin{pmatrix} 
  \alpha_x   &  0 & p_x(1-\alpha_x) \\
  0          &  \alpha_y & p_y(1-\alpha_y)
  \end{pmatrix}
$$
or

```
 c(alpha[1], 0 , 0, alpha[2], p[1]*(1-alpha[1]), p[2]*(1-alpha[2]) )
```








##### Scaling with the center as a fixed point using matrix, 
```{r, echo=T, results="asis"}
inPlaceTrans<-function(fixedPoint, alpha){
  c( alpha[1],0,0, alpha[2], 
    (1-alpha[1])*fixedPoint[1],
    (1-alpha[2])*fixedPoint[2] )
}

svgR( 
  myBack(WH, "Scaling with the center as fixed point using matrix"),
  cirCirSqr(cxy=WH/2, 
      transform=list( matrix=inPlaceTrans( WH/2, c(0.5,1.5) ) )
  )
)
``` 

#### Rotation with the Center as a Fixed Point

For our second example, we peform a rotation of $\frac{\pi}{4}$  around the point c(400,200). *The difference here is that this time we
use R to solve the matrix equation.*

We choose our points x_i to be equaly space around the fixed point with radius 100. The points and mapping are:
$$ 
\begin{pmatrix} 
400+100 cos(\frac{2 \pi k}{3}) \\ 
200+ 100 sin(\frac{2 \pi k}{3}) \\ 
1 
\end{pmatrix} 
\mapsto
\begin{pmatrix} 
400+100 cos(\frac{2 \pi k}{3} + \frac{\pi}{4}) \\ 
200+ 100 sin(\frac{2 \pi k}{3}+ \frac{\pi}{4}) \\
1 
\end{pmatrix}
$$ 



```{r, echo=TRUE}
WH<-c(800,400)
theta<-pi/4 #The angle we will rotate about
r<-100
eta<-(1:3)*(2*pi)/3
X<-rbind( r*cos(eta), r*sin(eta), rep(1,3) ) + c(WH/2,0)
eta<-(1:3)*(2*pi)/3 +theta 
Y<-rbind( r*cos(eta), r*sin(eta), rep(1,3) ) + c(WH/2,0)
solve(t(X),t(Y))->A
as.numeric(t(A[,1:2]))->A
```



####3 Rotation though the center using matrix, 
```{r, echo=T, results="asis"}
svgR( 
  myBack(WH, "Rotation though the center using matrix"),
  cirCirSqr(cxy=WH/2 , transform=list( matrix= A[1:6] ) )
)
```




### Alternatives to a list parameters

To keep a clear conscience, here are some alternatives to
using lists for transform values.

Instead of a list we can use a string as so:

```{r, echo=T, results="asis"}
WH=c(600, 200) # window rect
wh=c(80,30) # small rect
pt=c(c,0)
svgR( 
  wh=WH, 
  graphPaper(wh=WH, dxy=c(50,50), labels=TRUE),
  g(
      rect( xy=pt, wh=wh, stroke ='black', fill='blue', opacity=.3),
      circle(cxy=pt, r=wh[2], stroke="black", fill="red", opacity=.3),
      transform="translate(200,100) rotate(-30) scale(.5,3) skewX(80)"
  )
 )
```

Note the latter can become awkward for dynamic situations when the amount and sequence of translation, scaling, rotation
are determined at runtime.  But this is easily overcome with some additional helper functions
 
  
```{r, echo=T}
 translate<-function(...)paste0("translate(", paste(list(...),collapse=","), ")")   
 rotate<-function(...)paste0("rotate(", paste(list(...),collapse=","), ")")   
 skewX<-function(...)paste0("skewX(", paste(list(...),collapse=","), ")")   
 skewY<-function(...)paste0("skewX(", paste(list(...),collapse=","), ")")   
 scale<-function(...)paste0("scale(", paste(list(...),collapse=","), ")")   
```

Now with the helper functions
```{r, echo=T, results="asis"}
WH=c(600, 200) # window rect
wh=c(80,30) # small rect
svgR( 
  wh=WH, 
  graphPaper(wh=WH, dxy=c(50,50),TRUE),
  rect( xy=c(0,0), wh=wh, stroke ='blue', fill='green', opacity=.3,
        transform=c(translate(200,100), rotate(-30),  scale(.5,3), skewX(80))
        )
 )
```

