---
title: ""
author: "M. S. Legrand"
output:
  html_document:
    includes:
      in_header: svgLogo.svg
    number_sections: no
    theme: cerulean
    toc: true
    toc_depth: 2
  pdf_document:
    toc: true
    toc_depth: 2 
---


This document describes some of the basics of svgR.  This document has been created from an R Markdown document using svgR. 
For more details on using R Markdown see <http://rmarkdown.rstudio.com>. 
For a good starting place on using SVG see <https://github.com/willianjusten/awesome-svg>


<style>
h1 {
   color: darkblue;
   font-size:3.5em;
}
</style>




<style>
#TOC {
  position: fixed;
  left: 0;
  top: 0;
  width: 220px;
  height: 100%;
  overflow:auto;
}
body {
  max-width: 800px;
  margin: auto;
  margin-left:225px;
  line-height: 20px;
}
</style>







```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```
```{r, echo=FALSE}
if(!exists("newId")){
  source('Util.R')
}
```

<hr><br/><br/> 

One Line Teasers
===

```{r child = 'OneLineTeasers.Rmd'}
```

<hr><br/><br/> 

```
Setting Up
========================================================
```{r child = 'SetUp.Rmd'}
```


<hr><br/><br/> 

```

Overview
=====
```{r child = 'overView.Rmd'}
```


Coordinates, Viewports, and Containers
=======================================================
```{r child = 'Coordinates.Rmd'}
```


<hr><br/><br/> 

The Transform Attribute
=======================================================
```{r child = 'Transforms.Rmd'}
```



<hr><br/><br/> 


<!-- shapes and structures, groups symbols, use -->
Shapes,  Paths, Markers 
=======================================================

```{r child = 'Shapes.Rmd'}
```

```{r child = 'Path.Rmd'}
```

```{r child = 'Marker.Rmd'}
```


<hr><br/><br/> 

```
The Text Element
=======================================================
```{r child = 'text.Rmd'}
```


<hr><br/><br/> 


The Fill Attribute
=======================================================
```{r child = 'Fill.Rmd'}
```


<hr>
```{r, echo=F, results="asis"}
svgR(  logo2(wh=c(600,50)))
```


The Filter Attribute
=======================================================
```{r child = 'filter-Overview.Rmd'}
```

## The Gaussian Blur Filter
```{r child = 'filter-GaussianBlur.Rmd'}
```

## Filter Coordinates
```{r child = 'filter-Coordinates.Rmd'}
```



# Color Filters
```{r child = 'filter-Colors.Rmd'}
```

```{r child = 'filter-feComponentTransfer.Rmd'}
```

# Combining Filter Primitives
```{r child = 'filter-Combining.Rmd'}
```

## DropShadows
```{r child = 'filter-DropShadows.Rmd'}
```
<!-- Textures -->
# Texture Filters
```{r child = 'filter-Textures.Rmd'}
```

# Lighting Filters
```{r child = 'filter-Lighting.Rmd'}
```

```{r child = 'caseStudy_lighting.Rmd'}
```



## The Shiny Bevel
```{r child = 'filter-ShinyBevel.Rmd'}
```



<!-- shapes modification: mask and markers -->
Shape Shifters
=======================================================

## The clip.path attibute
```{r child = 'ClipPath.Rmd'}
```
## The mask attribute
```{r child = 'Mask.Rmd'}
```

## Morpolgy filtering
```{r child = 'filter-ChangingShapes.Rmd'}
```
## Changing Shapes feDisplacementMap (Warping)
```{r child = 'filter-Displacement.Rmd'}
```

Animating
=======================================================

```{r child = 'animate-main.Rmd'}
```

```{r child = 'animate-set.Rmd'}
```

```{r child = 'animate-animate.Rmd'}
```

```{r child = 'animate-animateTransform.Rmd'}
```

```{r child = 'animate-animateMotion.Rmd'}
```
