---
title: ""
author: "M. S. Legrand"
output:
  html_document:
    includes:
      in_header: svgLogo.svg
    number_sections: no
    theme: cerulean
    toc: true
    toc_depth: 2
  pdf_document:
    toc: true
    toc_depth: 2 
---

```{r, echo=FALSE}
library(svgR)
path<-"./Compounds/"
source(paste0(path,"svgR_logo.R"))
source(paste0(path,"svgRTree.R"))
```

<style>
h1 {
   background: linear-gradient(#c8da3d, #220066);
   border-style: groove;
   border-width: 3px;
   border-left-width: 10px;
   border-right-width: 10px;
   border-color: #440044;
   line-height: 1.1;
   text-align: left;
   color: white;
   padding: 10px;
   font-size:3.5em;
   border-radius: 10px;
   text-shadow: -2px -2px 2px #222;
}
</style>

<style>
.line-center{
    margin:0;padding:0 10px;
    background:#fff;
    display:inline-block;
}
h6{

    text-align:center;
    position:relative;
    z-index:2;

}
h6:after{
    content:"";
    position:absolute;
    top:50%;
    left:0;
    right:0;
    border-top:solid 1px #DDD;
    z-index:-1;
}
</style>


<style>
#TOC {
  position: fixed;
  left: 0;
  top: 0;
  width: 220px;
  height: 100%;
  overflow:auto;
}
body {
  max-width: 800px;
  margin: auto;
  margin-left:225px;
  line-height: 20px;
}
</style>



```{r, echo=FALSE}
stopifnot(require(svgR, quietly=TRUE))
```


This document describes some of the basics of svgR.  This document has been created from an R Markdown document using svgR. 
For more details on using R Markdown see <http://rmarkdown.rstudio.com>. 
For a good starting place on using SVG see <http://tutorials.jenkov.com/svg> and 
<https://github.com/willianjusten/awesome-svg>

###### ∞
<br/><br/><br/><br/> 
<!-- sp -->

One Line Teasers
===============================================

```{r child = 'OneLineTeasers.Rmd'}
```

###### ∞
<br/><br/><br/><br/> 
<!-- sp -->


Setting Up
========================================================

```{r child = 'SetUp.Rmd'}
```

<hr><br/><br/><br/><br/> 
<!-- sp -->

<h6 end section/>

Overview
===============================
```{r child = 'overView.Rmd'}
```

###### ∞
<br/><br/><br/><br/> 
<!-- sp -->

Coordinates, Viewports, and Containers
=======================================================
```{r child = 'Coordinates.Rmd'}
```

###### ∞
<br/><br/><br/><br/> 
<!-- sp -->

The Transform Attribute
=======================================================

```{r child = 'Transforms.Rmd'}
```


###### ∞
<br/><br/><br/><br/> 
<!-- sp -->


<!-- shapes and structures, groups symbols, use -->
Shapes,  Paths, Markers 
=======================================================

```{r child = 'Shapes.Rmd'}
```

```{r child = 'Path.Rmd'}
```

```{r child = 'Marker.Rmd'}
```

###### ∞
<br/><br/><br/><br/> 
<!-- sp -->


Text
=======================================================

```{r child = 'text.Rmd'}
```


###### ∞
<br/><br/><br/><br/> 
<!-- sp -->


The Fill Attribute
=======================================================

```{r child = 'Fill.Rmd'}
```

<hr><br/><br/> <br/>
<hr>
```{r, echo=F, results="asis"}
svgR(  logo2(wh=c(600,50)))
```

<hr><br/><br/><br/><br/> 
<!-- sp -->

The Filter Attribute
=======================================================

```{r child = 'filter-Overview.Rmd'}
```

## The Gaussian Blur Filter
```{r child = 'filter-GaussianBlur.Rmd'}
```

## Filter Coordinates
```{r child = 'filter-Coordinates.Rmd'}
```
###### ∞
<br/><br/><br/><br/> 
<!-- sp -->

Manipulating Colors
================================

```{r child = 'filter-Colors.Rmd'}
```

```{r child = 'filter-feComponentTransfer.Rmd'}
```

###### ∞
<br/><br/><br/><br/> 
<!-- sp -->

Combining Filter Primitives
==============
```{r child = 'filter-Combining.Rmd'}
```

###### ∞
<br/><br/><br/><br/> 
<!-- sp -->

DropShadows
=====
```{r child = 'filter-DropShadows.Rmd'}
```

###### ∞
<br/><br/><br/><br/> 
<!-- sp -->

Textures
==============
```{r child = 'filter-Textures.Rmd'}
```

###### ∞
<br/><br/><br/><br/> 
<!-- sp -->

Lighting
=============
```{r child = 'filter-Lighting.Rmd'}
```

```{r child = 'caseStudy_lighting.Rmd'}
```

###### ∞
<br/><br/><br/><br/> 
<!-- sp -->

Creating Shiny Bevels
===

```{r child = 'filter-ShinyBevel.Rmd'}
```

###### ∞
<br/><br/><br/><br/> 
<!-- sp -->

Shape Shifting
=======================================================

## Clipping
```{r child = 'ClipPath.Rmd'}
```
## Masking
```{r child = 'Mask.Rmd'}
```

## Morpolgy
```{r child = 'filter-ChangingShapes.Rmd'}
```

## Displacing (aka. Warping)
```{r child = 'filter-Displacement2.Rmd'}
```

###### ∞
<br/><br/><br/><br/> 
<!-- sp -->
>

Animating
=======================================================

```{r child = 'animate-main.Rmd'}
```

###### ∞
<br/><br/><br/><br/> 
<!-- sp -->

Discrete Animation
===
```{r child = 'animate-set.Rmd'}
```

###### ∞
<br/><br/><br/><br/> 
<!-- sp -->

Continous Animation 
===
```{r child = 'animate-animate.Rmd'}
```

###### ∞
<br/><br/><br/><br/> 
<!-- sp -->

Continous Movement via the Transform Attribute
===

```{r child = 'animate-animateTransform.Rmd'}
```

###### ∞
<br/><br/><br/><br/> 
<!-- sp -->

Continous Movement along a Path
===
```{r child = 'animate-animateMotion.Rmd'}
```

###### ∞
<br/><br/><br/><br/> 
<!-- sp -->
